### 2018-12-19 paper.tex
Add initial TeX files

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Paper Title*\\
}

\author{\IEEEauthorblockN{Tugrul Yatagan\IEEEauthorrefmark{1} and
Sema F. Oktug\IEEEauthorrefmark{2}}
\IEEEauthorblockA{Department of Computer Engineering,
Istanbul Technical University\\
Istanbul, Turkey\\
Email: \IEEEauthorrefmark{1}yatagan@itu.edu.tr,
\IEEEauthorrefmark{2}oktug@itu.edu.tr}}
\maketitle

\begin{abstract}
Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands. Focus on this work is one of the most promising LPWAN technologies: Lora. LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS). LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth. SF assignment of nodes has strong effect on collisions thus network performance. In this work, we implemented a simulation environment to study different SF assignment schemes and we proposed a machine learning technique to optimize SF assignment. Finally, we illustrate simulation results for proposed machine learning SF assignment technique.
\end{abstract}

\begin{IEEEkeywords}
LoRa, Spreading Factor, IoT, LPWAN, Machine Learning
\end{IEEEkeywords}

\section{Introduction}
\par In the last few years, number of Internet of Things applications increase exponentially. \cite{7721743} Recent developments on LPWAN technologies has great impact on IoT application number growth. LPWAN technologies offer solutions to some of the oldest wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications don't require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best. \cite{7815384} LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, SigFox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost. Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.

LoRa has open standard MAC protocol called LoRaWAN.\\
LoRaWAN and SigFox use unslotted ALOHA based MAC protocol.\\
Private LoRaWAN networks can be deployed however SigFox and NB-IoT are operator based.\\
Sigfox limits the number of messages an end device can send (140 UL, 4 DL per day).\\
Sigfox payload is very limited 12 bytes UL and 8 bytes DL. LoRa 243 bytes, NB-IoT 1600 bytes.\\
Sigfox maximum data rate is also very low 100 bps, LoRa 50 kbps, NB-IoT 200 kbps.\\
LoRa can change data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between sensitivity and data rate. LoRa has 7 different spreading factor option. Different spreading factor communications are orthogonal to each other.

\par TODO: SF Assignment Issue on LoRa

\par TODO: Proposal for SF Assignment Issue on LoRa

\par TODO: Organization of the paper

\section{Other Related Works}
\par TODO: Performance evaluation of lora networks in a smart city scenario\cite{7996384}

\par TODO: Scalability analysis of large-scale lorawan networks in ns-3 \cite{8090518}

\par TODO: Lora scalability: A simulation model based on interference measurements \cite{s17061193}

\par TODO: Impact of lora imperfect orthogonality: Analysis of link-level performance \cite{8267219}

\par TODO: Scalability analysis of a lora network under imperfect orthogonality \cite{8430542}

\section{LoRa}
\par LoRa is the name of the physical layer Radio/Chipset technology that provides wireless link for low power wide area networks. LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS). LoRa is registered trademark of Semtech Corporation. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently and incorrectly used for each other.

\subsection{LoRaWAN}
\par LoRaWAN is a media access control (MAC) layer protocol which designed for large scale LoRa networks. LoRaWAN is an open source standard developed and maintained by LoRa Alliance. LoRa Alliance is an open, non-profit organization dedicated to promoting the interoperability and standardization of LoRaWAN. LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions. LoRaWAN also provides interoperability between different networks. LoRaWAN uses unslotted ALOHA to access medium. LoRaWAN provides lightweight but powerful standard for IoT applications.

\par TODO: explain end node, gateway, class A

\subsection{LoRa Modulation}
\par TODO

\subsection{Spreading Factor}
\par LoRa CSS modulation can spread the symbols.\\
The spread factor is the ratio between symbol rate and chip rate. The ratio between symbol and chip rate is $2$\textsuperscript{spreadfactor}.\\
Six different spread factors are available (between 7 and 12); increasing the spread factor makes the signal more robust to noise, but decreases the data rate.\\

\par Lowest SF suitable area section is relatively higher then the others. Most of the end devices which close to GW will fall into this area. Figure~\ref{fig:problem1}

\par End devices near the GW will probably select lowest SF all the time. Which results a lot of collisions between lowest SF transmissions while end device number near the GW increases. Figure~\ref{fig:problem2}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora0-1}
\caption{Different SF ranges.}
\label{fig:problem1}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora1-1}
\caption{Collision between nodes near the GW.}
\label{fig:problem2}
\end{figure}

\section{Proposed Technique}
\par If GW force some of close end nodes to select higher SF even if they can able to communicate with lower SF, this will result lower collisions due to the orthogonality of different SF. Figure~\ref{fig:proposal_single_gw}

\par GW can avoid increasing SF of end nodes which is close to other GW using location information obtained by triangulating the signal strength of end nodes. Figure~\ref{fig:proposal_multi_gw}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora2-1}
\caption{Collision between nodes near the GW can be prevented by using higher SF.}
\label{fig:proposal_single_gw}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{lora3-1}
\caption{Collision prevention proposal for intersecting GWs.}
\label{fig:proposal_multi_gw}
\end{figure}

\section{Simulation Environment}
A discrete event simulator is implemented in Python to simulate LoRa network performance. 
https://github.com/tugrulyatagan/simlorafs

\subsection{Link Model}
TODO

\subsection{Interference Model}
TODO

\subsection{Simulation Assumptions}
TODO

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{sf_1000}
\caption{PDR of different SF when radius is 1000m.}
\label{fig:sf_1000}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{sf_3000}
\caption{PDR of different SF when radius is 3000m.}
\label{fig:sf_3000}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{sf_10000}
\caption{PDR of different SF when radius is 10000m.}
\label{fig:sf_10000}
\end{figure}

\section{Simulation Results}
TODO: Evaluation

\section{Conclusion}
TODO: Summary, future works
\cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}

\section*{Acknowledgment}
This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1. The authors would also like to thank .TODO. for their guidance in .TODO.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2018-12-23 paper.tex
Move other works
figures updated
related work added

@@
 
-\title{Paper Title*\\
+\title{Paper Title*
 }
@@
 
-\section{Other Related Works}
-\par TODO: Performance evaluation of lora networks in a smart city scenario\cite{7996384}
-
-\par TODO: Scalability analysis of large-scale lorawan networks in ns-3 \cite{8090518}
-
-\par TODO: Lora scalability: A simulation model based on interference measurements \cite{s17061193}
-
-\par TODO: Impact of lora imperfect orthogonality: Analysis of link-level performance \cite{8267219}
-
-\par TODO: Scalability analysis of a lora network under imperfect orthogonality \cite{8430542}
-
 \section{LoRa}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{lora0-1}
+\includegraphics[width=0.5\textwidth]{lora0}
 \caption{Different SF ranges.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{lora1-1}
+\includegraphics[width=0.5\textwidth]{lora1}
 \caption{Collision between nodes near the GW.}
@@
 \end{figure}
+
+\section{Other Related Works}
+The literature related to the work presented in this paper has started growing recently.
+LPWAN technologies and especially LoRa/LoRaWAN attract researchers attention lately.
+We listed some of these works which studying  LoRaWAN and LoRa spreading factor.
+
+\par In \cite{7996384}, the authors evaluated the performance of LoRa networks in a smart city scenario. The
+authors proposed a link measurement and a link performance model for LoRa.
+The authors also proposed a SINR threshold matrix for modeling LoRa interference between simultaneous but different spreading factor LoRa transmissions.
+They implement a LoRa simulator in ns-3 to study scalability and performance of LoRaWAN networks.
+Their results show that LoRaWAN networks scale well as the number of nodes and gateways increases.
+They also show that spreading factor assignment has great effect on network performance.
+
+\par In \cite{8090518}, another LoRaWAN ns-3 simulator is presented.
+Authors introduced an error model for determining range as well as interference between multiple simultaneous LoRa transmissions.
+Their simulator supports LoRaWAN class A end devices, multiple gateways, both upstream and downstream confirmed messages.
+Their results show that allocating network parameters to end devices is hugely important for the performance of LoRaWAN networks.
+
+\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements.
+They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from measurements.
+Their results show that LoRaWAN has better scalability than pure Aloha since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.
+
+\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa spreading factor transmissions.
+The authors state that a LoRa transmission can be interfered even between different spreading factor transmissions when power of the interfering signal significantly overcomes the reference signal.
+Their experimental results show that this power difference is around 16 dB and this power difference can be seen when  interferer is close to receiver or multiple interferer can create this power difference cumulatively.
+
+\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions using the same spreading factor as well as different spreading factors.
+They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage performance with respect to the distance from the gateway for modeling interference in multiple gateway scenarios.
+Their results show that transmission among different spreading factors can cause a significant impact in high-density LoRaWAN networks.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{lora2-1}
+\includegraphics[width=0.5\textwidth]{lora2}
 \caption{Collision between nodes near the GW can be prevented by using higher SF.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{lora3-1}
+\includegraphics[width=0.5\textwidth]{lora3}
 \caption{Collision prevention proposal for intersecting GWs.}
@@
 
-\section{Simulation Environment}
+\section{Simulation Environment and Results}
 A discrete event simulator is implemented in Python to simulate LoRa network performance. 
@@
 
-\subsection{Link Model}
+\subsection{Simulation Environment}
+
+\subsubsection{Link Model}
 TODO
 
-\subsection{Interference Model}
+\subsubsection{Interference Model}
 TODO
 
-\subsection{Simulation Assumptions}
+\subsubsection{Simulation Assumptions}
 TODO
@@
 
-\section{Simulation Results}
+\subsection{Simulation Results}
 TODO: Evaluation
@@
 \section*{Acknowledgment}
-This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No.
-ITU-AYP-2017-1.
-The authors would also like to thank .TODO.
-for their guidance in .TODO.
+This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1.
 

### 2018-12-24 paper.tex
intro intro

@@
 Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
-Cellular networks can provide long range and high data rate but they are complex and consume too much power.
+(CITE) Cellular networks can provide long range and high data rate but they are complex and consume too much power.
 Besides, most of the IoT applications don't require high data rate.
@@
 LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost.
-Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
-
-LoRa has open standard MAC protocol called LoRaWAN.\\
-LoRaWAN and SigFox use unslotted ALOHA based MAC protocol.\\
-Private LoRaWAN networks can be deployed however SigFox and NB-IoT are operator based.\\
-Sigfox limits the number of messages an end device can send (140 UL, 4 DL per day).\\
-Sigfox payload is very limited 12 bytes UL and 8 bytes DL. LoRa 243 bytes, NB-IoT 1600 bytes.\\
+(CITE) Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
+LoRa has open standard MAC protocol called LoRaWAN.
+(CITE) LoRaWAN and SigFox MAC protocols are based on pure ALOHA medium access.
+(CITE) LoRaWAN networks can be deployed as private network.
+However, SigFox and NB-IoT supports only operator contracted deployments.
+(CITE) Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
+Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink.
+(CITE) However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
+(CITE)
 Sigfox maximum data rate is also very low 100 bps, LoRa 50 kbps, NB-IoT 200 kbps.\\
@@
 LoRaWAN also provides interoperability between different networks.
-LoRaWAN uses unslotted ALOHA to access medium.
+LoRaWAN uses pure ALOHA to access medium.
 LoRaWAN provides lightweight but powerful standard for IoT applications.
@@
 They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from measurements.
-Their results show that LoRaWAN has better scalability than pure Aloha since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.
+Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.
 

### 2018-12-25 paper.tex
lorawan definition

@@
 LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth.
-SF assignment of nodes has strong effect on collisions thus network performance.
+Spreading factor (SF) assignment of nodes has strong effect on collisions thus network performance.
 In this work, we implemented a simulation environment to study different SF assignment schemes and we proposed a machine learning technique to optimize SF assignment.
@@
 (CITE) However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
+(CITE) Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively.
 (CITE)
-Sigfox maximum data rate is also very low 100 bps, LoRa 50 kbps, NB-IoT 200 kbps.\\
-LoRa can change data rate by spreading symbols within a fixed channel bandwidth.
-This enables tradeoff between sensitivity and data rate.
-LoRa has 7 different spreading factor option.
-Different spreading factor communications are orthogonal to each other.
 
-\par TODO: SF Assignment Issue on LoRa
+\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth.
+This enables tradeoff between receive sensitivity and transmission air time.
+(CITE) LoRa supports 6 different spreading factor option.
+Simultaneous same spreading factor transmissions are prone to collision however different spreading factor transmissions in same channel are orthogonal to each other, thus spreading factor assignment is crucial for overall network performance.
+(CITE) In this work, we introduce a custom made LoRa discrete event simulator to study LoRa spreading factor assignment issue.
+We also proposed an spreading factor assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.
 
-\par TODO: Proposal for SF Assignment Issue on LoRa
+\par The following section gives an overview about LoRa and LoRaWAN.
+Section \ref{Other Related Works} describes other related works.
+Section \ref{Proposed Technique} describes proposed technique for spreading factor assignment.
+Simulation environments are described and simulation results are shown in Section \ref{Simulation Environment and Results}.
+Finally, Section \ref{Conclusion} concludes the paper.
 
-\par TODO: Organization of the paper
-
-\section{LoRa}
-\par LoRa is the name of the physical layer Radio/Chipset technology that provides wireless link for low power wide area networks.
+\section{LoRa} \label{LoRa}
+\par LoRa is the name of the physical layer Radio/Chipset technology that provides wireless link solution for low power wide area networks.
 LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS).
@@
 \subsection{LoRaWAN}
-\par LoRaWAN is a media access control (MAC) layer protocol which designed for large scale LoRa networks.
+\par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
 LoRaWAN is an open source standard developed and maintained by LoRa Alliance.
-LoRa Alliance is an open, non-profit organization dedicated to promoting the interoperability and standardization of LoRaWAN.
+LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
 LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions.
-LoRaWAN also provides interoperability between different networks.
-LoRaWAN uses pure ALOHA to access medium.
-LoRaWAN provides lightweight but powerful standard for IoT applications.
+LoRaWAN also provides inter-operability between different LoRa networks.
+LoRaWAN is based on pure ALOHA medium access.
+LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.
 
-\par TODO: explain end node, gateway, class A
+\par TODO: Add LoRaWAN network architecture figure here.
+
+\par A typical LoRaWAN network consist of three network entities.
+\subsubsection{End node}
+Low power embedded device that communicates to gateways.
+LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios.
+Class A end nodes generates uplink transmission at any time and only receives a period of time after uplink transmission.
+Class B end nodes extend Class A by adding scheduled receive windows for downlink transmission.
+Receive window is synchronized using beacon packet transmitted by gateway.
+Class C end nodes extend Class A by keeping receive window open unless they are transmitting.
+This provides low latency donwlink communication but requires more power consumption.
+(CITE) In this paper, we focused on class A end devices since Class A provides lowest power consumption.
+
+\subsubsection{Gateway}
+Devices that receive/transmit packets coming from/to end nodes.
+A typical gateway can receive from multiple channel at the same time.
+Gateways are typically connected to power grid so power consumption of a gateway is insignificant most of the deployments.
+
+\subsubsection{Network server}
+Server that provides MAC layer processing. NS routes messages from application to end nodes, and vice versa.
 
@@
 \par TODO
+LoRa CSS modulation can spread the symbols.\\
+TODO frequency and band
 
-\subsection{Spreading Factor}
-\par LoRa CSS modulation can spread the symbols.\\
-The spread factor is the ratio between symbol rate and chip rate. The ratio between symbol and chip rate is $2$\textsuperscript{spreadfactor}.\\
-Six different spread factors are available (between 7 and 12); increasing the spread factor makes the signal more robust to noise, but decreases the data rate.\\
+We invite readers to refer to \cite{AN1200.22} for further details.
+
+\par Spreading factor is the ratio between symbol rate and chip rate. (CITE) The ratio between symbol and chip rate is equal to $2$\textsuperscript{SF}. SF can take values between 7 to 12.
+
+and determines the LoRa symbol rate as: Rs = (BW/2SF).
+As the SF increases the symbol rate is lowered, thereby trading reduced data rate for increased range.
+For a bandwidth of 125 kHz, the PHY data rates range from 6835 to 365 b/s for SF7–SF12.
+The different SFs are orthogonal.
+This means that a LoRa gateway can receive multiple transmissions on different SFs simultaneously.
+Note that SF bits are mapped per LoRa symbol.
+
+increasing the spread factor makes the signal more robust to noise, but decreases the data rate.
 
@@
 \includegraphics[width=0.5\textwidth]{lora0}
-\caption{Different SF ranges.}
+\caption{Different SF transmission ranges.}
 \label{fig:problem1}
@@
 \includegraphics[width=0.5\textwidth]{lora1}
-\caption{Collision between nodes near the GW.}
+\caption{Collision between nodes close to the GW.}
 \label{fig:problem2}
@@
 
-\section{Other Related Works}
+\section{Other Related Works} \label{Other Related Works}
 The literature related to the work presented in this paper has started growing recently.
@@
 
-\section{Proposed Technique}
+\section{Proposed Technique} \label{Proposed Technique}
 \par If GW force some of close end nodes to select higher SF even if they can able to communicate with lower SF, this will result lower collisions due to the orthogonality of different SF.
@@
 \includegraphics[width=0.5\textwidth]{lora2}
-\caption{Collision between nodes near the GW can be prevented by using higher SF.}
+\caption{Collision between nodes close to the GW can be prevented by using higher SF.}
 \label{fig:proposal_single_gw}
@@
 
-\section{Simulation Environment and Results}
-A discrete event simulator is implemented in Python to simulate LoRa network performance. 
-https://github.com/tugrulyatagan/simlorafs
+\section{Simulation Environment and Results} \label{Simulation Environment and Results}
+A discrete event simulator is implemented in Python to simulate LoRa network performance. https://github.com/tugrulyatagan/simlorafs
 
@@
 
-\section{Conclusion}
+\section{Conclusion} \label{Conclusion}
 TODO: Summary, future works

### 2018-12-25 paper.tex
lorawan definition

@@
 
-\par Lowest SF suitable area section is relatively higher then the others. Most of the end devices which close to GW will fall into this area. Figure~\ref{fig:problem1}
+\par Lowest SF suitable area section is relatively higher then the others. Most of the end devices which close to GW will fall into this area.
 
@@
 Which results a lot of collisions between lowest SF transmissions while end device number near the GW increases.
-Figure~\ref{fig:problem2}
+Figure~\ref{fig:collision}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{lora0}
-\caption{Different SF transmission ranges.}
-\label{fig:problem1}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{lora1}
-\caption{Collision between nodes close to the GW.}
-\label{fig:problem2}
+\includegraphics[width=0.5\textwidth]{collision}
+\caption{Collision between nodes near the GW.}
+\label{fig:collision}
 \end{figure}
@@
 \par If GW force some of close end nodes to select higher SF even if they can able to communicate with lower SF, this will result lower collisions due to the orthogonality of different SF.
-Figure~\ref{fig:proposal_single_gw}
+Figure~\ref{fig:collision_solution_single_gw}
 
-\par GW can avoid increasing SF of end nodes which is close to other GW using location information obtained by triangulating the signal strength of end nodes. Figure~\ref{fig:proposal_multi_gw}
+\par GW can avoid increasing SF of end nodes which is close to other GW using location information obtained by triangulating the signal strength of end nodes.
+Figure~\ref{fig:collision_solution_multi_gw}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{lora2}
+\includegraphics[width=0.5\textwidth]{collision_solution_single_gw}
 \caption{Collision between nodes close to the GW can be prevented by using higher SF.}
-\label{fig:proposal_single_gw}
+\label{fig:collision_solution_single_gw}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{lora3}
+\includegraphics[width=0.5\textwidth]{collision_solution_multi_gw}
 \caption{Collision prevention proposal for intersecting GWs.}
-\label{fig:proposal_multi_gw}
+\label{fig:collision_solution_multi_gw}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{sf_1000}
-\caption{PDR of different SF when radius is 1000m.}
+\includegraphics[width=0.5\textwidth]{sf_pdr}
+\caption{PDR of different spreading factors.}
 \label{fig:sf_1000}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{sf_3000}
-\caption{PDR of different SF when radius is 3000m.}
+\includegraphics[width=0.5\textwidth]{gw_pdr}
+\caption{PDR of different gateway numbers.}
 \label{fig:sf_3000}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{sf_10000}
-\caption{PDR of different SF when radius is 10000m.}
+\includegraphics[width=0.5\textwidth]{r_pdr}
+\caption{PDR of different topology radius.}
+\label{fig:sf_10000}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{pr_pdr}
+\caption{PDR of different packet rates.}
 \label{fig:sf_10000}

### 2018-12-26 paper.tex
gateway, packet rate, radius, spreading factor figures
spreading factor and data rate

@@
 Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
-(CITE) Cellular networks can provide long range and high data rate but they are complex and consume too much power.
+[CITE] Cellular networks can provide long range and high data rate but they are complex and consume too much power.
 Besides, most of the IoT applications don't require high data rate.
@@
 LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost.
-(CITE) Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
+[CITE] Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
 LoRa has open standard MAC protocol called LoRaWAN.
-(CITE) LoRaWAN and SigFox MAC protocols are based on pure ALOHA medium access.
-(CITE) LoRaWAN networks can be deployed as private network.
+[CITE] LoRaWAN and SigFox MAC protocols are based on pure ALOHA medium access.
+[CITE] LoRaWAN networks can be deployed as private network.
 However, SigFox and NB-IoT supports only operator contracted deployments.
-(CITE) Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
+[CITE] Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
 Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink.
-(CITE) However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
-(CITE) Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively.
-(CITE)
+[CITE] However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
+[CITE] Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively.
+[CITE]
 
@@
 This enables tradeoff between receive sensitivity and transmission air time.
-(CITE) LoRa supports 6 different spreading factor option.
-Simultaneous same spreading factor transmissions are prone to collision however different spreading factor transmissions in same channel are orthogonal to each other, thus spreading factor assignment is crucial for overall network performance.
-(CITE) In this work, we introduce a custom made LoRa discrete event simulator to study LoRa spreading factor assignment issue.
-We also proposed an spreading factor assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.
+[CITE] LoRa supports 6 different SF option.
+Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
+[CITE] In this work, we introduce a custom made LoRa discrete event simulator to study LoRa SF assignment issue.
+We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.
 
-\par The following section gives an overview about LoRa and LoRaWAN.
+\par The following two sections provides an background information about LoRaWAN and LoRa.
 Section \ref{Other Related Works} describes other related works.
-Section \ref{Proposed Technique} describes proposed technique for spreading factor assignment.
-Simulation environments are described and simulation results are shown in Section \ref{Simulation Environment and Results}.
+Section \ref{Proposed Technique} describes proposed technique for SF assignment.
+Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}.
 Finally, Section \ref{Conclusion} concludes the paper.
 
-\section{LoRa} \label{LoRa}
-\par LoRa is the name of the physical layer Radio/Chipset technology that provides wireless link solution for low power wide area networks.
-LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS).
-LoRa is registered trademark of Semtech Corporation.
-LoRa has an open standard MAC protocol called LoRaWAN.
-LoRa and LoRaWAN terms are frequently and incorrectly used for each other.
+\section{LoRaWAN} \label{LoRaWAN}
+\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently and incorrectly used for each other.
 
-\subsection{LoRaWAN}
 \par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
@@
 
-\par TODO: Add LoRaWAN network architecture figure here.
+\par [TODO] Add LoRaWAN network architecture figure here.
 
 \par A typical LoRaWAN network consist of three network entities.
-\subsubsection{End node}
-Low power embedded device that communicates to gateways.
+
+\subsection{End node}
+\par LoRaWAN end node is low power embedded device that communicates to gateways.
 LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios.
@@
 This provides low latency donwlink communication but requires more power consumption.
-(CITE) In this paper, we focused on class A end devices since Class A provides lowest power consumption.
+[CITE] In this paper, we focused on class A end devices since Class A provides lowest power consumption.
 
-\subsubsection{Gateway}
-Devices that receive/transmit packets coming from/to end nodes.
+\subsection{Gateway}
+\par LoRaWAN gateway is device that receive/transmit packets coming from/to end nodes.
 A typical gateway can receive from multiple channel at the same time.
@@
 
-\subsubsection{Network server}
-Server that provides MAC layer processing. NS routes messages from application to end nodes, and vice versa.
+\subsection{Network server}
+\par LoRaWAN network server is server that provides MAC layer processing.
+Network server routes messages from application to end nodes, and vice versa.
+Network server can be used for tweaking end node parameters like channel, transmit power, SF to increase network performance.
 
-\subsection{LoRa Modulation}
-\par TODO
-LoRa CSS modulation can spread the symbols.\\
-TODO frequency and band
+\section{LoRa}
+\par LoRa is the name of the proprietary physical layer radio and chipset technology that provides wireless link solution for low power wide area networks.
+LoRa is designed and patented by Semtech Corporation.
+LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS).
+Chirp is a sinusoidal signal which its  frequency increases over time.
+The chirp frequency increases linearly and sweeps entire bandwidth.
+[CITE] We invite readers to refer to \cite{AN1200.22} for further details.
 
-We invite readers to refer to \cite{AN1200.22} for further details.
+\subsection{Spreading Factor}
+\par Spreading factor (SF) is the ratio between symbol rate and chip rate.
+The ratio between symbol and chip rate is equal to $2$\textsuperscript{SF}.
+SF can take values between 7 to 12.
+[CITE] SF also determines data rate of a LoRa transmission.
 
-\par Spreading factor is the ratio between symbol rate and chip rate. (CITE) The ratio between symbol and chip rate is equal to $2$\textsuperscript{SF}. SF can take values between 7 to 12.
+\begin{equation} \label{eq:bit_rate_sf}
+R_{b} = SF * \dfrac{\left[ \dfrac{4}{4+CR} \right] }{ \left[ \dfrac{2^{SF}}{BW} \right]} \ bits/sec
+\end{equation}
 
-and determines the LoRa symbol rate as: Rs = (BW/2SF).
-As the SF increases the symbol rate is lowered, thereby trading reduced data rate for increased range.
-For a bandwidth of 125 kHz, the PHY data rates range from 6835 to 365 b/s for SF7–SF12.
-The different SFs are orthogonal.
-This means that a LoRa gateway can receive multiple transmissions on different SFs simultaneously.
-Note that SF bits are mapped per LoRa symbol.
+Where SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and BW is bandwidth in Hertz. [CITE:modulation basics]
 
-increasing the spread factor makes the signal more robust to noise, but decreases the data rate.
+\par While BW and CR are constant, as the SF increases the data rate decreases.
+Increasing SF makes the signal more resilient to noise thus increases the transmission range.
+Increasing SF also increases transmission duration which increases power consumption.
+Therefore it is possible to trade between data rate, range and power consumption by  changing SF.
 
-\par Lowest SF suitable area section is relatively higher then the others. Most of the end devices which close to GW will fall into this area.
+\subsection{Spreading Factor Assignment Issue}
+\par Simultaneous different SF transmissions are orthogonal to each other up to some extent.
+Which means, a LoRa gateway can simultaneously receive multiple transmissions with different SFs.
+However, simultaneous transmissions with same SF may not be received by gateway due to collision.
+For this reason, spreading factor assignment of nodes is crucial for network performance.
 
-\par End devices near the GW will probably select lowest SF all the time.
-Which results a lot of collisions between lowest SF transmissions while end device number near the GW increases.
+\par [TODO] Lowest SF suitable area section is relatively higher then the others. Most of the end devices which close to GW will fall into this area.
+
+\par [TODO] End devices which close to the GW will probably select lowest SF all the time.
+Which results a lot of collisions between lowest SF transmissions while end device number close to the GW increases.
 Figure~\ref{fig:collision}
@@
 \includegraphics[width=0.5\textwidth]{collision}
-\caption{Collision between nodes near the GW.}
+\caption{Collision between nodes close to the GW.}
 \label{fig:collision}
@@
 LPWAN technologies and especially LoRa/LoRaWAN attract researchers attention lately.
-We listed some of these works which studying  LoRaWAN and LoRa spreading factor.
+We listed some of these works which studying LoRaWAN and LoRa SF.
 
@@
 authors proposed a link measurement and a link performance model for LoRa.
-The authors also proposed a SINR threshold matrix for modeling LoRa interference between simultaneous but different spreading factor LoRa transmissions.
+The authors also proposed a SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
 They implement a LoRa simulator in ns-3 to study scalability and performance of LoRaWAN networks.
 Their results show that LoRaWAN networks scale well as the number of nodes and gateways increases.
-They also show that spreading factor assignment has great effect on network performance.
+They also show that SF assignment has great effect on network performance.
 
@@
 
-\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa spreading factor transmissions.
-The authors state that a LoRa transmission can be interfered even between different spreading factor transmissions when power of the interfering signal significantly overcomes the reference signal.
+\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa SF transmissions.
+The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal.
 Their experimental results show that this power difference is around 16 dB and this power difference can be seen when  interferer is close to receiver or multiple interferer can create this power difference cumulatively.
 
-\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions using the same spreading factor as well as different spreading factors.
+\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions using the same SF as well as different SFs.
 They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage performance with respect to the distance from the gateway for modeling interference in multiple gateway scenarios.
-Their results show that transmission among different spreading factors can cause a significant impact in high-density LoRaWAN networks.
+Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.
 
 \section{Proposed Technique} \label{Proposed Technique}
+\par [TODO]
+
 \par If GW force some of close end nodes to select higher SF even if they can able to communicate with lower SF, this will result lower collisions due to the orthogonality of different SF.
@@
 
-\section{Simulation Environment and Results} \label{Simulation Environment and Results}
-A discrete event simulator is implemented in Python to simulate LoRa network performance. https://github.com/tugrulyatagan/simlorafs
+\section{Simulation Environment} \label{Simulation Environment}
+[TODO] A discrete event simulator is implemented in Python to simulate LoRa network performance. https://github.com/tugrulyatagan/simlorafs
 
-\subsection{Simulation Environment}
+\begin{table}
+\centering
+\caption{Gateway Sensitivity for different SFs}
+\label{table:gw_sf_sensitivity}
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\hline
+            &         & \multicolumn{6}{c|}{\textbf{SF}}        \\ \cline{3-8} 
+            &         &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
+            & 125 kHz & -123 & -126 & -129 & -132 & -133 & -136 \\ 
+\textbf{BW} & 250 kHz & -120 & -123 & -125 & -128 & -130 & -133 \\
+            & 500 kHz & -116 & -119 & -122 & -125 & -128 & -130 \\
+\hline
+\end{tabular}
+\end{table}
+[CITE:datasheet]
 
-\subsubsection{Link Model}
-TODO
+\begin{equation} \label{eq:rx_sensitivity}
+P^{dBm}
+\end{equation}
+[CITE:modulation basics]
 
-\subsubsection{Interference Model}
-TODO
+\subsection{Link Model}
+[TODO]
 
-\subsubsection{Simulation Assumptions}
-TODO
+\subsection{Interference Model}
+[TODO]
+
+\subsection{Simulation Assumptions}
+[TODO]
 
@@
 \includegraphics[width=0.5\textwidth]{sf_pdr}
-\caption{PDR of different spreading factors.}
-\label{fig:sf_1000}
+\caption{PDR of different SFs.}
+\label{fig:sf_pdr}
 \end{figure}
@@
 \caption{PDR of different gateway numbers.}
-\label{fig:sf_3000}
+\label{fig:gw_pdr}
 \end{figure}
@@
 \caption{PDR of different topology radius.}
-\label{fig:sf_10000}
+\label{fig:r_pdr}
 \end{figure}
@@
 \caption{PDR of different packet rates.}
-\label{fig:sf_10000}
+\label{fig:pr_pdr}
 \end{figure}
 
-\subsection{Simulation Results}
-TODO: Evaluation
+\section{Simulation Results} \label{Simulation Results}
+[TODO] Evaluation
 
 \section{Conclusion} \label{Conclusion}
-TODO: Summary, future works
+[TODO] Summary, future works
 \cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}

### 2018-12-26 paper.tex
add datasheet reference
link budget

@@
 SF can take values between 7 to 12.
-[CITE] SF also determines data rate of a LoRa transmission.
+SF also determines data rate of a LoRa transmission \cite{AN1200.22}.
 
@@
 
-Where SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and BW is bandwidth in Hertz. [CITE:modulation basics]
+Where, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.
 
@@
 \section{Simulation Environment} \label{Simulation Environment}
-[TODO] A discrete event simulator is implemented in Python to simulate LoRa network performance. https://github.com/tugrulyatagan/simlorafs
+\par [TODO] A discrete event simulator is implemented in Python to simulate LoRa network performance. https://github.com/tugrulyatagan/simlorafs
+
+\subsection{Link Model}
+\par [TODO]
+
+The link budget of a wireless system or network is a measure of all the gains and losses from the
+transmitter, through the propagation channel, to the target receiver. These gains and losses include
+system gains and losses associated with the antenna, matching networks, etc. as well as losses
+associated propagation channel itself (either though modelling or measured data).
+Typically randomly varying channel mechanisms such as multipath and Doppler fading are taken into
+account by factoring additional margin depending on the anticipated severity.
+The link budget of a network wireless link can be expressed as:
+
+\begin{equation} \label{eq:rx_sensitivity}
+P^{dBm}_{RX} = P^{dBm}_{TX} + G^{dB}_{SYS} - L^{dB}_{SYS} - L^{dB}_{PATH}
+\end{equation}
+
+Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver.
+$P^{dBm}_{TX}$ is the transmit power.
+$G^{dB}_{SYS}$ is the system gains such as transmitter and receiver antenna gain.
+$L^{dB}_{SYS}$ is the system losses such as transmitter and receiver line, circuit, antenna losses.
+$L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas.
+
+\par Receive sensitivity of a LoRa gateway for different SFs and BWs can be found in Table \ref{table:gw_sf_sensitivity} \cite{SX1276}.
 
@@
 \end{table}
-[CITE:datasheet]
-
-\begin{equation} \label{eq:rx_sensitivity}
-P^{dBm}
-\end{equation}
-[CITE:modulation basics]
-
-\subsection{Link Model}
-[TODO]
 

### 2018-12-27 paper.tex
propagation

@@
 
-The link budget of a wireless system or network is a measure of all the gains and losses from the
-transmitter, through the propagation channel, to the target receiver. These gains and losses include
-system gains and losses associated with the antenna, matching networks, etc. as well as losses
-associated propagation channel itself (either though modelling or measured data).
-Typically randomly varying channel mechanisms such as multipath and Doppler fading are taken into
-account by factoring additional margin depending on the anticipated severity.
-The link budget of a network wireless link can be expressed as:
+The link budget of a wireless system or network is a measure of all the gains and losses from the transmitter, through the propagation channel, to the target receiver.
+Link budget of a wireless link can be calculated as:
 
-\begin{equation} \label{eq:rx_sensitivity}
+\begin{equation} \label{eq:expected_rx_power}
 P^{dBm}_{RX} = P^{dBm}_{TX} + G^{dB}_{SYS} - L^{dB}_{SYS} - L^{dB}_{PATH}
@@
 
-Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver.
+\par Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver.
 $P^{dBm}_{TX}$ is the transmit power.
-$G^{dB}_{SYS}$ is the system gains such as transmitter and receiver antenna gain.
+$G^{dB}_{SYS}$ is the system gains such as transmitter and receiver antenna gains.
 $L^{dB}_{SYS}$ is the system losses such as transmitter and receiver line, circuit, antenna losses.
 $L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas.
+
+\par Propagation loss can be calculated as:
+
+\begin{equation} \label{eq:propagation_loss}
+\begin{split}
+P^{dB}_{PATH} = 40(1 - 4 \times 10^{-3} \times h){\log_{10} R|_{km}} \\
+- 18 {\log_{10} h|_{m}} + 21 {\log_{10} f|_{MHz}} + 80
+\end{split}
+\end{equation}
+
+\par Where $h$ is the gateway altitude and $f$ is the frequency of the signal. We assume that $h$ = 15 m and $f$ = 868 MHz, Propagation loss can be calculated as:
+
+\begin{equation} \label{eq:propagation_loss}
+P^{dB}_{PATH} = 120.5 + 37.6 {\log_{10} R|_{km}}
+\end{equation}
 

### 2018-12-30 paper.tex
interference model

@@
 
-\title{Paper Title*
+\title{Smart Spreading Factor Assignment for LoRaWAN
 }
@@
 
-\par [TODO] Add LoRaWAN network architecture figure here.
-
 \par A typical LoRaWAN network consist of three network entities.
@@
 
-\par [TODO] Lowest SF suitable area section is relatively higher then the others. Most of the end devices which close to GW will fall into this area.
+\par Initially, a node is not aware of how far it is from a GW.
+However, a node can guess how far away from the gateway by observing received signal power of a downlink transmission.
+If a downlink transmission received signal power is too high then node can decrease its next transmission SF or transmit power to decrease power consumption.
+Also GW can request from a node to decrease node's SF or transmit power. This SD assignment is called lowest possible SF assignemnt scheme for the rest of the paper.
 
-\par [TODO] End devices which close to the GW will probably select lowest SF all the time.
-Which results a lot of collisions between lowest SF transmissions while end device number close to the GW increases.
-Figure~\ref{fig:collision}
+\par In Figure~\ref{fig:collision}, a LoRaWAN network distributed around a single GW is illustrated.
+Different color rings represent achievable range of different SF from GW and different color circles represent nodes with selected SF.
+Most of the end devices which close to the GW will fall into lowest SF (SF7) area section.
+End devices which close to the GW will probably select lowest SF all the time.
+Which results a lot of collisions between same SF transmissions.
+Collision number increases while end device number close to the GW increases.
 
@@
 \par Receive sensitivity of a LoRa gateway for different SFs and BWs can be found in Table \ref{table:gw_sf_sensitivity} \cite{SX1276}.
+Maximum TX power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
 
@@
 
+\begin{table}
+\centering
+\caption{EU863-870 ISM Band LoRa Default Channels}
+\label{table:max_tx_power}
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Channel} & \textbf{Frequency} & \textbf{Bandwidth} & \textbf{Max EIRP} \\ \hline
+      1 &     868.1 &   125 kHz &   14 dBm \\
+      2 &     868.3 &   125 kHz &   14 dBm \\
+      3 &     868.5 &   125 kHz &   14 dBm \\
+
+\hline
+\end{tabular}
+\end{table}
+
 \subsection{Interference Model}
-[TODO]
+\par [TODO]
+
+\begin{equation} \label{eq:sinr}
+T = \begin{bmatrix}
+       6 & -16 & -18 & -19 & -19 & -20 \\
+     -24 &   6 & -20 & -22 & -22 & -22 \\
+     -27 & -27 &   6 & -23 & -25 & -25 \\
+     -30 & -30 & -30 &   6 & -26 & -28 \\
+     -33 & -33 & -33 & -33 &   6 & -29 \\
+     -36 & -36 & -36 & -36 & -36 &   6
+     \end{bmatrix}
+\end{equation}
 
 \subsection{Simulation Assumptions}
-[TODO]
+\par [TODO]
 
@@
 \section{Simulation Results} \label{Simulation Results}
-[TODO] Evaluation
+\par [TODO] Evaluation
 
 \section{Conclusion} \label{Conclusion}
-[TODO] Summary, future works
+\par [TODO] Summary, future works
 \cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}
@@
 \section*{Acknowledgment}
-This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1.
+\par This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1.
 

### 2018-12-31 paper.tex
proposed teq. conlcusion

@@
 \usepackage{xcolor}
+\usepackage{multirow}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
-\title{Smart Spreading Factor Assignment for LoRaWAN
-}
+\title{Smart Spreading Factor Assignment for LoRaWANs}
 
-\author{\IEEEauthorblockN{Tugrul Yatagan\IEEEauthorrefmark{1} and
-Sema F. Oktug\IEEEauthorrefmark{2}}
+\author{\IEEEauthorblockN{Tugrul Yatagan\IEEEauthorrefmark{1} and Sema F. Oktug\IEEEauthorrefmark{2}}
 \IEEEauthorblockA{Department of Computer Engineering,
@@
 Istanbul, Turkey\\
-Email: \IEEEauthorrefmark{1}yatagan@itu.edu.tr,
-\IEEEauthorrefmark{2}oktug@itu.edu.tr}}
+Email: \IEEEauthorrefmark{1}yatagan@itu.edu.tr, \IEEEauthorrefmark{2}oktug@itu.edu.tr}}
 \maketitle
@@
 Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands.
-Focus on this work is one of the most promising LPWAN technologies: Lora.
-LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS).
+Focus on this work is one of the most promising LPWAN technologies: LoRa.
+LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum.
 LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth.
@@
 \par There are several emerging LPWAN technologies.
-LoRa, SigFox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies.
-LoRa and SigFox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost.
-[CITE] Both LoRa and SigFox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
+LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies.
+LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost.
+[CITE] Both LoRa and Sigfox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
 LoRa has open standard MAC protocol called LoRaWAN.
-[CITE] LoRaWAN and SigFox MAC protocols are based on pure ALOHA medium access.
+[CITE] LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access.
 [CITE] LoRaWAN networks can be deployed as private network.
-However, SigFox and NB-IoT supports only operator contracted deployments.
+However, Sigfox and NB-IoT supports only operator contracted deployments.
 [CITE] Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
@@
 Section \ref{Proposed Technique} describes proposed technique for SF assignment.
-Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}.
+Simulation environments and results are shown in Section \ref{Simulation Environment and Results}.
 Finally, Section \ref{Conclusion} concludes the paper.
@@
 \subsection{End node}
-\par LoRaWAN end node is low power embedded device that communicates to gateways.
+\par LoRaWAN end node (EN) is low power embedded device that communicates to gateways.
 LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios.
@@
 This provides low latency donwlink communication but requires more power consumption.
-[CITE] In this paper, we focused on class A end devices since Class A provides lowest power consumption.
+[CITE] In this paper, we focused on Class A end devices since Class A provides lowest power consumption.
 
 \subsection{Gateway}
-\par LoRaWAN gateway is device that receive/transmit packets coming from/to end nodes.
+\par LoRaWAN gateway (GW) is device that receive/transmit packets coming from/to end nodes.
 A typical gateway can receive from multiple channel at the same time.
@@
 \subsection{Network server}
-\par LoRaWAN network server is server that provides MAC layer processing.
+\par LoRaWAN network server (NS) is server that provides MAC layer processing.
 Network server routes messages from application to end nodes, and vice versa.
@@
 If a downlink transmission received signal power is too high then node can decrease its next transmission SF or transmit power to decrease power consumption.
-Also GW can request from a node to decrease node's SF or transmit power. This SD assignment is called lowest possible SF assignemnt scheme for the rest of the paper.
+Also GW can request from a node to decrease node's SF or transmit power.
+This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper.
 
@@
 Authors introduced an error model for determining range as well as interference between multiple simultaneous LoRa transmissions.
-Their simulator supports LoRaWAN class A end devices, multiple gateways, both upstream and downstream confirmed messages.
+Their simulator supports LoRaWAN Class A end devices, multiple gateways, both upstream and downstream confirmed messages.
 Their results show that allocating network parameters to end devices is hugely important for the performance of LoRaWAN networks.
@@
 \section{Proposed Technique} \label{Proposed Technique}
-\par [TODO]
+\par The collision issue illustrated in Figure~\ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even if they can able to communicate with lower SF.
+This may result lower collisions due to the orthogonality of different SF as shown in Figure~\ref{fig:collision_solution_single_gw}.
+Higher SF assigned nodes are drawn with bold circle border in Figure~\ref{fig:collision_solution_single_gw}.
+However, how GW assign SF to which node becomes an another issue.
+Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions.
+In multiple GW scenarios, this approach may increase the collision number with nodes in other GW's range.
+Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure~\ref{fig:collision_solution_multi_gw}.
 
-\par If GW force some of close end nodes to select higher SF even if they can able to communicate with lower SF, this will result lower collisions due to the orthogonality of different SF.
-Figure~\ref{fig:collision_solution_single_gw}
+\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes SFs requires different rules.
+For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collision number.
+This technique learns nodes behavior of a network.
+A NS can keep track of successful uplink transmission from nodes and their SFs.
+NS can also keep track of some of the collided transmission if header part of the packet does not interfered at GW.
+However, NS cannot keep track of transmissions with receive power lower than GW sensitivity.
+Using these obtained information NS can train a classifier for predicting transmission result for specific node and specific SF input.
+Using this predictor GWs can assign the lowest possible SF to nodes considering the collision probability.
 
-\par GW can avoid increasing SF of end nodes which is close to other GW using location information obtained by triangulating the signal strength of end nodes.
-Figure~\ref{fig:collision_solution_multi_gw}
+\par In this work, we used decision tree classifier to predict transmission results.
+We utilized Gini index criteria to measure quality of decision tree splits and used balanced class weights.
+As most of the machine learning applications, increasing data set size increases prediction accuracy.
+For our case, data set size is directly proportional to simulation duration.
+Thus increasing simulation duration, increases the prediction accuracy.
+For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GW can request from nodes to use this new SF.
 
@@
 
-\section{Simulation Environment} \label{Simulation Environment}
-\par [TODO] A discrete event simulator is implemented in Python to simulate LoRa network performance. https://github.com/tugrulyatagan/simlorafs
+\section{Simulation Environment and Results} \label{Simulation Environment and Results}
+\par A discrete event simulator is implemented in Python to study LoRa network performance. The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorafs.
+
+Simulation Assumptions
 
@@
 
-\begin{equation} \label{eq:propagation_loss}
+\begin{equation} \label{eq:propagation_loss_simplified}
 P^{dB}_{PATH} = 120.5 + 37.6 {\log_{10} R|_{km}}
@@
 \hline
-            &         & \multicolumn{6}{c|}{\textbf{SF}}        \\ \cline{3-8} 
-            &         &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
-            & 125 kHz & -123 & -126 & -129 & -132 & -133 & -136 \\ 
-\textbf{BW} & 250 kHz & -120 & -123 & -125 & -128 & -130 & -133 \\
-            & 500 kHz & -116 & -119 & -122 & -125 & -128 & -130 \\
-\hline
+\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
+\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
+\multirow{3}{*}{\textbf{BW}}  & 125 kHz & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
+                              & 250 kHz & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
+                              & 500 kHz & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
 \end{tabular}
@@
 \textbf{Channel} & \textbf{Frequency} & \textbf{Bandwidth} & \textbf{Max EIRP} \\ \hline
-      1 &     868.1 &   125 kHz &   14 dBm \\
-      2 &     868.3 &   125 kHz &   14 dBm \\
+      1 &     868.1 &   125 kHz &   14 dBm \\ \hline
+      2 &     868.3 &   125 kHz &   14 dBm \\ \hline
       3 &     868.5 &   125 kHz &   14 dBm \\
@@
 \end{equation}
-
-\subsection{Simulation Assumptions}
-\par [TODO]
 
@@
 
-\section{Simulation Results} \label{Simulation Results}
+\subsection{Simulation Results}
 \par [TODO] Evaluation
@@
 \section{Conclusion} \label{Conclusion}
-\par [TODO] Summary, future works
-\cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}
+\par In this paper, after brief introduction about LPWAN technologies, we provide background information about one of the most promising LPWAN technology, LoRa.
+Then, we discuss about LoRa modulation basics and spreading factor assignment issue.
+We have implemented a discrete event simulator from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
+We have shown how same SF transmission collisions can be avoided and we proposed a machine learning based solution for collision with same SF transmissions.
+We illustrate simulation results for proposed machine learning SF assignment technique.
+Our simulation results show that, proposed technique can increase network performance up to [TODO] \%.
+
+\par As for future works, transmit power optimization can be included to proposed machine learning scheme.
+In this paper, we assume nodes always uses maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
+This will make transmission more vulnerable to interference thus requires extra care.
+Optimizing another node parameter alongside SF can be interesting research.
+Also other machine learning methods can be investigated for SF optimization.
+Reinforcement learning can be a good candidate for this network optimization issue.
+
+\par \cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}
 

### 2018-12-31 paper.tex
interference

@@
 
+\begin{equation} \label{eq:sinr_db}
+SINR_{i,j} = \dfrac{P_{rc,0}}{\sum_{l \in I_j} P_{rc,l}}
+\end{equation}
+
+\begin{equation} \label{eq:sinr_t}
+SINR_{i,j}^{dB} > T_{i,j}
+\end{equation}
+
+\begin{equation} \label{eq:p_interference}
+P_{rc,y}^{interf} = \dfrac{P_{rc,y}(T_{x}-t_{1})}{T_{x}}
+\end{equation}
+
 \begin{figure}
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{predict_pdr}
+\caption{PDR of different prediction SF assignment scheme.}
+\label{fig:pr_pdr}
+\end{figure}
+
 \subsection{Simulation Results}
@@
 We illustrate simulation results for proposed machine learning SF assignment technique.
-Our simulation results show that, proposed technique can increase network performance up to [TODO] \%.
+Our simulation results show that, proposed technique can increase network performance up to [TBD] \%.
 

### 2019-01-01 paper.tex
simulation env

@@
 \begin{equation} \label{eq:bit_rate_sf}
-R_{b} = SF * \dfrac{\left[ \dfrac{4}{4+CR} \right] }{ \left[ \dfrac{2^{SF}}{BW} \right]} \ bits/sec
+R_{b}|_{bps} = SF * \dfrac{\left[ \dfrac{4}{4+CR} \right] }{ \left[ \dfrac{2^{SF}}{BW|_{Hz}} \right]}
 \end{equation}
@@
 
-\par In Figure~\ref{fig:collision}, a LoRaWAN network distributed around a single GW is illustrated.
+\par In Figure \ref{fig:collision}, a LoRaWAN network distributed around a single GW is illustrated.
 Different color rings represent achievable range of different SF from GW and different color circles represent nodes with selected SF.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{collision}
+\includegraphics[width=\linewidth]{collision}
 \caption{Collision between nodes close to the GW.}
@@
 \section{Proposed Technique} \label{Proposed Technique}
-\par The collision issue illustrated in Figure~\ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even if they can able to communicate with lower SF.
-This may result lower collisions due to the orthogonality of different SF as shown in Figure~\ref{fig:collision_solution_single_gw}.
-Higher SF assigned nodes are drawn with bold circle border in Figure~\ref{fig:collision_solution_single_gw}.
+\par The collision issue illustrated in Figure \ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even if they can able to communicate with lower SF.
+This may result lower collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}.
+Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}.
 However, how GW assign SF to which node becomes an another issue.
@@
 In multiple GW scenarios, this approach may increase the collision number with nodes in other GW's range.
-Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure~\ref{fig:collision_solution_multi_gw}.
+Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{collision_solution_single_gw}
-\caption{Collision between nodes close to the GW can be prevented by using higher SF.}
+\includegraphics[width=\linewidth]{collision_solution_single_gw}
+\caption{Collision avoidance proposal between nodes close to the GW by using higher SF.}
 \label{fig:collision_solution_single_gw}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{collision_solution_multi_gw}
-\caption{Collision prevention proposal for intersecting GWs.}
+\includegraphics[width=\linewidth]{collision_solution_multi_gw}
+\caption{Collision avoidance proposal for intersecting GWs.}
 \label{fig:collision_solution_multi_gw}
@@
 \section{Simulation Environment and Results} \label{Simulation Environment and Results}
-\par A discrete event simulator is implemented in Python to study LoRa network performance. The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorafs.
-
-Simulation Assumptions
+\par A discrete event simulator is implemented in Python to study LoRa network performance considering SF orthogonality.
+The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorafs.
+Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
+Simulator can generate uniformly distributed random topology with node number and gateway number input.
+Global simulation parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
+With these input, tool provide outputs of total generated packet number, successfully received packet number, interfered packet number, under sensitivity packet number, network packet delivery ratio (PDR), network throughput and total transmit energy consumption.
 
 \subsection{Link Model}
-\par [TODO]
-
-The link budget of a wireless system or network is a measure of all the gains and losses from the transmitter, through the propagation channel, to the target receiver.
-Link budget of a wireless link can be calculated as:
+\par Link quality of a wireless system can be expressed by the metric of link budget.
+Link budget is a measure of all gains and losses from transmitter to receiver.
+Link budget of a wireless link can be calculated as \cite{AN1200.22}:
 
@@
 \par Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver.
-$P^{dBm}_{TX}$ is the transmit power.
+$P^{dBm}_{TX}$ is the transmit power of the transmitter.
 $G^{dB}_{SYS}$ is the system gains such as transmitter and receiver antenna gains.
 $L^{dB}_{SYS}$ is the system losses such as transmitter and receiver line, circuit, antenna losses.
-$L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas.
+$L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas in open space.
+In our simulation we assume that total of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dBm.
 
-\par Propagation loss can be calculated as:
+\begin{table}
+\centering
+\caption{EU863-870 ISM Band LoRa Default Channels \cite{lorawan.regional.parameters}}
+\label{table:max_tx_power}
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Channel} & \textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
+      1 &     868.1 &   125 &   14 \\ \hline
+      2 &     868.3 &   125 &   14 \\ \hline
+      3 &     868.5 &   125 &   14 \\
+
+\hline
+\end{tabular}
+\end{table}
+
+\par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
+In our simulator we assume that nodes are always uses maximum allowed transmit power which is 14 dBm.
+Different channel transmissions are independent from each other.
+But, in this work we focuse on SF orthogonality.
+Thus, we only utilized single channel transmissions in our simulator.
+
+\begin{table}
+\centering
+\caption{Gateway Sensitivity for different SFs \cite{SX1276}}
+\label{table:gw_sf_sensitivity}
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\hline
+\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
+\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
+\multirow{3}{*}{\textbf{BW (kHz)}}  & 125 & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
+                                    & 250 & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
+                                    & 500 & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
+\end{tabular}
+\end{table}
+
+\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm can be found in Table \ref{table:gw_sf_sensitivity} \cite{SX1276}.
+In our simulation, we use 125 kHz bandwidth which is also the only allowed bandwidth in EU863-870 regulation.
+
+\par Free space propagation loss can be calculated as \cite{TR136.942}:
 
@@
 
-\par Where $h$ is the gateway altitude and $f$ is the frequency of the signal. We assume that $h$ = 15 m and $f$ = 868 MHz, Propagation loss can be calculated as:
+\par Where $h$ is the gateway altitude and $f$ is the frequency of the signal. We assume that $h$ = 15 m and $f$ = 868 MHz. With these assumptions, propagation loss calculation become \cite{7996384}:
 
@@
 
-\par Receive sensitivity of a LoRa gateway for different SFs and BWs can be found in Table \ref{table:gw_sf_sensitivity} \cite{SX1276}.
-Maximum TX power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
-
-\begin{table}
-\centering
-\caption{Gateway Sensitivity for different SFs}
-\label{table:gw_sf_sensitivity}
-\begin{tabular}{|c|c|c|c|c|c|c|c|}
-\hline
-\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
-\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
-\multirow{3}{*}{\textbf{BW}}  & 125 kHz & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
-                              & 250 kHz & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
-                              & 500 kHz & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
-\end{tabular}
-\end{table}
-
-\begin{table}
-\centering
-\caption{EU863-870 ISM Band LoRa Default Channels}
-\label{table:max_tx_power}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Channel} & \textbf{Frequency} & \textbf{Bandwidth} & \textbf{Max EIRP} \\ \hline
-      1 &     868.1 &   125 kHz &   14 dBm \\ \hline
-      2 &     868.3 &   125 kHz &   14 dBm \\ \hline
-      3 &     868.5 &   125 kHz &   14 dBm \\
-
-\hline
-\end{tabular}
-\end{table}
+\par If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission at the transmission period.
 
 \subsection{Interference Model}
-\par [TODO]
+\par We invite readers to refer to \cite{7996384}
 
@@
 
+\par Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$.
+Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.
+Duration of a LoRa transmission:
+
+\begin{equation} \label{eq:transmission_duration}
+t_{x}|_{s} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}
+\end{equation}
+
 \begin{equation} \label{eq:p_interference}
-P_{rc,y}^{interf} = \dfrac{P_{rc,y}(T_{x}-t_{1})}{T_{x}}
+P_{rc,y}^{interf} = \dfrac{P_{rc,y}(t_{interf})}{t_{x}}
 \end{equation}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{sf_pdr}
+\includegraphics[width=\linewidth]{sf_pdr_r3000_g1_p001_s3600}
 \caption{PDR of different SFs.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{gw_pdr}
+\includegraphics[width=\linewidth]{gw_pdr_r3000_p001_s3600}
 \caption{PDR of different gateway numbers.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{r_pdr}
+\includegraphics[width=\linewidth]{r_pdr_g1_p001_s3600}
 \caption{PDR of different topology radius.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{pr_pdr}
+\includegraphics[width=\linewidth]{pr_pdr_r3000_g1_s3600}
 \caption{PDR of different packet rates.}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{predict_pdr}
+\includegraphics[width=\linewidth]{predict_pdr_r5000_g3_p001_s3600}
 \caption{PDR of different prediction SF assignment scheme.}
-\label{fig:pr_pdr}
+\label{fig:predict_pdr}
 \end{figure}
@@
 We illustrate simulation results for proposed machine learning SF assignment technique.
-Our simulation results show that, proposed technique can increase network performance up to [TBD] \%.
+Our simulation results show that, proposed technique can increase network performance for dense LoRaWAN network deployed close to GWs.
 
 \par As for future works, transmit power optimization can be included to proposed machine learning scheme.
-In this paper, we assume nodes always uses maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
+In this paper, we assume nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
 This will make transmission more vulnerable to interference thus requires extra care.
@@
 
-\par \cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}
+%% \par \cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}
 

### 2019-01-01 paper.tex
pdr figures updated

@@
 Which results a lot of collisions between same SF transmissions.
-Collision number increases while end device number close to the GW increases.
+Collision number increases while number of end devices close to the GW increases.
 
@@
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
-Simulator can generate uniformly distributed random topology with node number and gateway number input.
+Simulator can generate uniformly distributed random topology with number of nodes and number of gateways input.
 Global simulation parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
-With these input, tool provide outputs of total generated packet number, successfully received packet number, interfered packet number, under sensitivity packet number, network packet delivery ratio (PDR), network throughput and total transmit energy consumption.
+With these input, the simulator provide outputs of total number of generated packets, number of successfully received packet, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
 
@@
 $L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas in open space.
-In our simulation we assume that total of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dBm.
+In our simulation we assume that total of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dB.
 
@@
 \subsection{Interference Model}
-\par We invite readers to refer to \cite{7996384}
+\par We have adopted the interference model described in \cite{7996384} to our simulator.
+They introduced SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
+
+\par We assume that there is no other technology interference is present in the network except LoRa interference.
+To exploit imperfect orthogonality of different SF transmission, we need to calculate the effect of different SF transmissions to each other.
+We used signal to interference plus noise ratio (SINR) threshold matrix \cite{goursaud:hal-01231221}:
 
@@
 
+\par To calculate if a referenced signal is interfered at receiver by a interferer signal, we use SINR threshold matrix.
+$T_{i,j}$ is the needed SINR margin in dB between referenced signal with SF = i and interferer signal with SF = j to correctly decode referenced signal.
+If there are more than one interferer, referenced signal must satisfy the margin for cumulative sum of all received power values for each SF \cite{7996384}.
+
 \begin{equation} \label{eq:sinr_db}
@@
 \end{equation}
+
+\par
 
@@
 
+\par TODOOOOOOOOO
+
 \begin{equation} \label{eq:p_interference}
@@
 
+\subsection{Simulation Results}
+\par We have been able to evaluate different 
+For simulation results in this paper, we set packet size as 60 Bytes including header and payload.
+
 \begin{figure}
 \centering
-\includegraphics[width=\linewidth]{sf_pdr_r3000_g1_p001_s3600}
+\includegraphics[width=\linewidth]{{sf_pdr_r3000_g1_p0.01_s3600}.png}
 \caption{PDR of different SFs.}
@@
 \centering
-\includegraphics[width=\linewidth]{gw_pdr_r3000_p001_s3600}
-\caption{PDR of different gateway numbers.}
+\includegraphics[width=\linewidth]{{gw_pdr_r3000_p0.01_s3600}.png}
+\caption{PDR of different number of gateways.}
 \label{fig:gw_pdr}
@@
 \centering
-\includegraphics[width=\linewidth]{r_pdr_g1_p001_s3600}
+\includegraphics[width=\linewidth]{{r_pdr_g1_p0.01_s3600}.png}
 \caption{PDR of different topology radius.}
@@
 \centering
-\includegraphics[width=\linewidth]{pr_pdr_r3000_g1_s3600}
+\includegraphics[width=\linewidth]{{pr_pdr_r3000_g1_s3600}.png}
 \caption{PDR of different packet rates.}
@@
 \centering
-\includegraphics[width=\linewidth]{predict_pdr_r5000_g3_p001_s3600}
+\includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
 \caption{PDR of different prediction SF assignment scheme.}
@@
 \end{figure}
-
-\subsection{Simulation Results}
-\par [TODO] Evaluation
 

### 2019-01-01 paper.tex
ml

@@
 
+\par In our simulator we only cover LoRaWAN Class A devices.
+Transmissions are always initiated by end nodes in pure ALOHA manner.
+Nodes generate a new packet according to Poisson interval for given packet rate input.
+In our simulator, downlink transmissions are not considered.
+Downlink transmissions are very rare in real world deployments since ISM band regulations dictates duty cycle transmission limit for all devices including GW.
+This makes downlink transmissions to very rare.
+
 \subsection{Link Model}
@@
 
-\par To calculate if a referenced signal is interfered at receiver by a interferer signal, we use SINR threshold matrix.
-$T_{i,j}$ is the needed SINR margin in dB between referenced signal with SF = i and interferer signal with SF = j to correctly decode referenced signal.
-If there are more than one interferer, referenced signal must satisfy the margin for cumulative sum of all received power values for each SF \cite{7996384}.
+\par To decide if a referenced signal is interfered at receiver by a interferer signal, we use SINR threshold matrix.
+$T_{i,j}$ is the needed SINR margin in dB between referenced signal with SF = i and interferer signal with SF = j to correctly decode the referenced signal.
+If there are more than one interferer, referenced signal must satisfy the margin for cumulative sum of all interfering signal received power values for each SF \cite{7996384}.
 
@@
 
-\par
+\par Where $P_{rc,0}$ is received signal power of referenced signal and $P_{rc,l}$ is received signal power of interfering signal at SF = j.
+If a packet with SF = i satisfies the following condition for every SF = j, then packet is survived from all interference.
 
@@
 
-\par Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$.
+\par To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power value \cite{7996384}:
+
+\begin{equation} \label{eq:p_interference}
+P_{rc,y}^{interf} = \dfrac{P_{rc,y}(t_{interf})}{t_{x}}
+\end{equation}
+
+\par Where $t_{x}$ is transmission duration of referenced signal.
+$t_{interf}$ is intersection duration between referenced signal and interfering signal.
+Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$.
 Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.
@@
 
-\par TODOOOOOOOOO
+\subsection{Machine Learning Model}
+\par We have been able to generate mass amount of LoRaWAN transmission data for different topologies using our  tool.
+To come up with a solution for SF assignment issue, we have integrated a machine learning scheme to our simulator.
+This will allow us train a model from generated transmission data and we use this trained model to select best possible SF for nodes using prediction of the model.
+Tool first run a simulation with assigning random SF to nodes.
+Simulation results are combined into four columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
+Transmission result can be successful, interfered or under sensitivity.
+These data is fed to Python scikit-learn decision tree classifier.
+After the model is created, a second simulation is run with a predictor.
+First, SF of nodes are assigned according to lowest possible scheme.
+Then, for every nodes if predictors predicts a interference
 
-\begin{equation} \label{eq:p_interference}
-P_{rc,y}^{interf} = \dfrac{P_{rc,y}(t_{interf})}{t_{x}}
-\end{equation}
+For machine learning model, we utilized at least three GWs for a LoRaWAN network knowing the fact that three GWs is sufficient for triangulation in real world deployments.
 
 \subsection{Simulation Results}
-\par We have been able to evaluate different 
-For simulation results in this paper, we set packet size as 60 Bytes including header and payload.
+\par For simulation results in this paper, we set global simulation parameters as follows. Packet size is 60 Bytes including header and payload. Simulation duration is 3600 seconds.
 

### 2019-01-01 paper.tex
figure evaluations

@@
 Section \ref{Proposed Technique} describes proposed technique for SF assignment.
-Simulation environments and results are shown in Section \ref{Simulation Environment and Results}.
+Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}.
 Finally, Section \ref{Conclusion} concludes the paper.
@@
 
-\section{Simulation Environment and Results} \label{Simulation Environment and Results}
+\section{Simulation Environment} \label{Simulation Environment}
 \par A discrete event simulator is implemented in Python to study LoRa network performance considering SF orthogonality.
@@
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
-Simulator can generate uniformly distributed random topology with number of nodes and number of gateways input.
+Simulator can generate uniformly distributed random circle shape network topology with radius, number of nodes and number of gateways inputs.
 Global simulation parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
@@
 After the model is created, a second simulation is run with a predictor.
-First, SF of nodes are assigned according to lowest possible scheme.
-Then, for every nodes if predictors predicts a interference
+Model tries to find the lowest possible SF with prediction of successful transmission.
 
-For machine learning model, we utilized at least three GWs for a LoRaWAN network knowing the fact that three GWs is sufficient for triangulation in real world deployments.
+\section{Simulation Results}  \label{Simulation Results}
+\par For simulation results in this paper, we set global simulation parameters as follows. Packet size is 60 bytes including header and payload. Simulation duration is 3600 seconds.
 
-\subsection{Simulation Results}
-\par For simulation results in this paper, we set global simulation parameters as follows. Packet size is 60 Bytes including header and payload. Simulation duration is 3600 seconds.
+\par In Figure \ref{fig:sf_pdr}, PDR plot of different SF assignment is shown.
+Randomly generated network topology radius is set to 3000 meters, number of GW is set to 1 and packet generation rate is set to 0.01 packet per second.
+Increasing SFs, increases time on air.
+This increases the number of collisions thus decreases the PDR of the network.
+PDR of higher SFs are decreased to almost 0 while number of node is increasing.
+Since network topology radius is quite small almost all SFs can reach to the GW.
 
@@
 
+\par In Figure \ref{fig:gw_pdr}, PDR plot of different number of GW is shown.
+Randomly generated topology radius is set to 3000 meters, packet generation rate is set to 0.01 packet per second and lowest possible SF assignment scheme is used.
+Increasing number of GWs, decreases the SFs of nodes and decreases time on air.
+This decreases number of collisions thus increases the PDR of the network when network topology radius is constant.
+
 \begin{figure}
@@
 \end{figure}
+
+\par In Figure \ref{fig:r_pdr}, PDR plot of different network topology radius is shown.
+Number of GW is set to 1, packet generation rate is set to 0.01 packet per second and lowest possible SF assignment scheme is used.
+Increasing network topology radius, increases the number of under sensitivity transmissions thus decreases the PDR of the network.
 
@@
 
+\par In Figure \ref{fig:pr_pdr}, PDR plot of different packet generation rate is shown.
+Randomly generated topology radius is set to 3000 meters, number of GW is set to 1 and lowest possible SF assignment scheme is used.
+Increasing packet generation rate, increases the number of collisions thus decreases the PDR of the network.
+
 \begin{figure}
@@
 \end{figure}
+
+\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning predictor SF scheme is shown.
+Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second.
+Our predictor model needs nodes location so we simulate predictor SF scheme with at three GWs since three GWs is enough to roughly locate nodes positions by triangulation.
+Predictor scheme gives better PDR when number of nodes increases since number of interference increases when number of nodes increases.
+Predictor scheme can improve performance when LoRa interference is present.
+Also predictor scheme gives better results when nodes are deployed close to the gateway, since nodes have a margin to increase their SFs.
+If a node is far away from the GW, then predictor scheme cannot increase the SF to avoid interference since the SF is already high.
 

### 2019-01-01 paper.tex
cite

@@
 \section{Introduction}
-\par In the last few years, number of Internet of Things applications increase exponentially.
-\cite{7721743} Recent developments on LPWAN technologies has great impact on IoT application number growth.
+\par In the last few years, number of Internet of Things applications increase exponentially \cite{7721743}.
+Recent developments on LPWAN technologies has great impact on IoT application number growth.
 LPWAN technologies offer solutions to some of the oldest wireless communication challenges.
 Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
-[CITE] Cellular networks can provide long range and high data rate but they are complex and consume too much power.
+Cellular networks can provide long range and high data rate but they are complex and consume too much power.
 Besides, most of the IoT applications don't require high data rate.
-Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best.
-\cite{7815384} LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication.
+Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}.
+LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication.
 LPWAN technologies basically sacrifice data rate to provide low power consumption.
@@
 LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies.
-LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost.
-[CITE] Both LoRa and Sigfox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
+LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}.
+Both LoRa and Sigfox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
 LoRa has open standard MAC protocol called LoRaWAN.
-[CITE] LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access.
-[CITE] LoRaWAN networks can be deployed as private network.
-However, Sigfox and NB-IoT supports only operator contracted deployments.
-[CITE] Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
+LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access.
+LoRaWAN networks can be deployed as private network.
+However, Sigfox and NB-IoT supports only operator contracted deployments \cite{7815384}.
+Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
 Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink.
-[CITE] However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
-[CITE] Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively.
-[CITE]
+However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
+Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.
 
 \par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth.
-This enables tradeoff between receive sensitivity and transmission air time.
-[CITE] LoRa supports 6 different SF option.
+This enables tradeoff between receive sensitivity and transmission air time \cite{7803607}.
+LoRa supports 6 different SF option.
 Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
-[CITE] In this work, we introduce a custom made LoRa discrete event simulator to study LoRa SF assignment issue.
+In this work, we introduce a custom made LoRa discrete event simulator to study LoRa SF assignment issue.
 We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.
@@
 This provides low latency donwlink communication but requires more power consumption.
-[CITE] In this paper, we focused on Class A end devices since Class A provides lowest power consumption.
+In this paper, we focused on Class A end devices since Class A provides lowest power consumption.
 
@@
 The chirp frequency increases linearly and sweeps entire bandwidth.
-[CITE] We invite readers to refer to \cite{AN1200.22} for further details.
+We invite readers to refer to \cite{AN1200.22} for further details.
 
@@
 \centering
-\caption{EU863-870 ISM Band LoRa Default Channels \cite{lorawan.regional.parameters}}
+\caption{EU863-870 ISM Band LoRa Default Channels \cite{lorawan.regional.parameters}.}
 \label{table:max_tx_power}
@@
 \centering
-\caption{Gateway Sensitivity for different SFs \cite{SX1276}}
+\caption{Gateway Sensitivity for different SFs \cite{SX1276}.}
 \label{table:gw_sf_sensitivity}
@@
 
-%% \par \cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}
-
 \section*{Acknowledgment}

### 2019-01-04 paper.tex
grammer

@@
 Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands.
-Focus on this work is one of the most promising LPWAN technologies: LoRa.
+Focus of this work is one of the most promising LPWAN technologies: LoRa.
 LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum.
 LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth.
-Spreading factor (SF) assignment of nodes has strong effect on collisions thus network performance.
-In this work, we implemented a simulation environment to study different SF assignment schemes and we proposed a machine learning technique to optimize SF assignment.
-Finally, we illustrate simulation results for proposed machine learning SF assignment technique.
+Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance.
+In this work, a simulation environment to study different SF assignment schemes is implemented.
+Furthermore, novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization.
+Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
 \end{abstract}
@@
 \section{Introduction}
-\par In the last few years, number of Internet of Things applications increase exponentially \cite{7721743}.
-Recent developments on LPWAN technologies has great impact on IoT application number growth.
-LPWAN technologies offer solutions to some of the oldest wireless communication challenges.
-Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
+\par In the last few years, number of Internet of Things (IoT) applications increased exponentially \cite{7721743}.
+Recent developments on LPWAN technologies has great impact on growth of number of IoT applications.
+LPWAN technologies addresses some of the well-known wireless communication challenges.
+Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
 Cellular networks can provide long range and high data rate but they are complex and consume too much power.
-Besides, most of the IoT applications don't require high data rate.
+Besides, most of the IoT applications do not require high data rate.
 Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}.
@@
 \par There are several emerging LPWAN technologies.
-LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies.
+LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies.
 LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}.
-Both LoRa and Sigfox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
+Both LoRa and Sigfox are known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
 LoRa has open standard MAC protocol called LoRaWAN.
 LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access.
-LoRaWAN networks can be deployed as private network.
-However, Sigfox and NB-IoT supports only operator contracted deployments \cite{7815384}.
+LoRaWAN networks can be deployed as a private network.
+However, Sigfox and NB-IoT are only available with operator contracted deployments \cite{7815384}.
 Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
-Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink.
+Also, Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink.
 However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
-Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.
+Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.
 
 \par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth.
-This enables tradeoff between receive sensitivity and transmission air time \cite{7803607}.
+This enables tradeoff between receive sensitivity and transmission time on air \cite{7803607}.
 LoRa supports 6 different SF option.
-Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
-In this work, we introduce a custom made LoRa discrete event simulator to study LoRa SF assignment issue.
-We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.
+Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other up to some extent, thus SF assignment is crucial for overall network performance.
+In this work, a custom made LoRa discrete event simulator is introduced to study different LoRa SF assignment strategies.
+We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using the simulator.
 

### 2019-01-05 paper.tex
grammer reprhrese

@@
 \begin{abstract}
-Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas using license-free frequency bands.
+Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas.
 Focus of this work is one of the most promising LPWAN technologies: LoRa.
-LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum.
-LoRa trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth.
+LoRa offers long range communication and strong resilience to interference by proprietary modulation technique based on Chirp Spread Spectrum (CSS).
+LoRa modulation trades data rate for sensitivity and communication range by spreading symbols within a fixed channel bandwidth.
 Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance.
@@
 LPWAN technologies addresses some of the well-known wireless communication challenges.
-Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
+Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication technologies (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
 Cellular networks can provide long range and high data rate but they are complex and consume too much power.
@@
 Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}.
-LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication.
+LPWAN technologies fill the technology gap between short range and cellular technologies by providing low power and long range communication.
 LPWAN technologies basically sacrifice data rate to provide low power consumption.
@@
 LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}.
-Both LoRa and Sigfox are known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate.
+Both LoRa and Sigfox are known for ultra low power consumption and resilience to interference.
+While NB-IoT and LTE-M are promoted for higher data rate.
 LoRa has open standard MAC protocol called LoRaWAN.
 LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access.
-LoRaWAN networks can be deployed as a private network.
-However, Sigfox and NB-IoT are only available with operator contracted deployments \cite{7815384}.
+LoRaWAN networks can be deployed as a private network like WiFi.
+However, Sigfox and NB-IoT are only available with operator contract \cite{7815384}.
 Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
-Also, Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink.
-However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload.
+Also, Sigfox packet payload size is limited to 12 bytes for uplink and 8 bytes for downlink.
+However LoRa supports up to 243 bytes payload size and NB-IoT supports up to 1600 bytes payload size.
 Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.
@@
 \section{LoRaWAN} \label{LoRaWAN}
-\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently and incorrectly used for each other.
+\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently yet incorrectly used for each other.
 
 \par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
-LoRaWAN is an open source standard developed and maintained by LoRa Alliance.
+LoRaWAN is an open source standard developed and it is maintained by LoRa Alliance.
 LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
-LoRa can be used without LoRaWAN as a wireless link technology, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions.
+LoRa can be used as a wireless link technology without complying LoRaWAN, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions.
 LoRaWAN also provides inter-operability between different LoRa networks.
-LoRaWAN is based on pure ALOHA medium access.
 LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.
+LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmitting, accepting the possibility of the collision.
 
@@
 \subsection{End node}
-\par LoRaWAN end node (EN) is low power embedded device that communicates to gateways.
-LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios.
-Class A end nodes generates uplink transmission at any time and only receives a period of time after uplink transmission.
-Class B end nodes extend Class A by adding scheduled receive windows for downlink transmission.
-Receive window is synchronized using beacon packet transmitted by gateway.
-Class C end nodes extend Class A by keeping receive window open unless they are transmitting.
-This provides low latency donwlink communication but requires more power consumption.
-In this paper, we focused on Class A end devices since Class A provides lowest power consumption.
+\par LoRaWAN end node (EN) is a low power embedded device that only communicates between gateways.
+LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
+Different classes provides LPWAN solutions to different deployments and topologies.
+Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission.
+Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission.
+Receive window is synchronized using beacon packet transmitted by gateways.
+Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
+This provides Class C end nodes to low latency donwlink communication but it requires more power consumption.
+In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.
 
 \subsection{Gateway}
-\par LoRaWAN gateway (GW) is device that receive/transmit packets coming from/to end nodes.
-A typical gateway can receive from multiple channel at the same time.
-Gateways are typically connected to power grid so power consumption of a gateway is insignificant most of the deployments.
+\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes.
+A typical gateway can receive from multiple channels at the same time.
+Gateways are typically connected to power grid so power consumption of a gateway is insignificant in most of the deployments.
 
 \subsection{Network server}
-\par LoRaWAN network server (NS) is server that provides MAC layer processing.
-Network server routes messages from application to end nodes, and vice versa.
-Network server can be used for tweaking end node parameters like channel, transmit power, SF to increase network performance.
+\par LoRaWAN network server (NS) is a server that provides MAC layer processing.
+Network server routes messages from application to end nodes and vice versa.
+Network server can be used for tweaking end node parameters like channel, transmit power and SF to increase network performance.
 
 \section{LoRa}
-\par LoRa is the name of the proprietary physical layer radio and chipset technology that provides wireless link solution for low power wide area networks.
+\par LoRa is the name of the proprietary physical layer radio/chipset technology that provides wireless link solution for low power wide area networks.
 LoRa is designed and patented by Semtech Corporation.
 LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS).
-Chirp is a sinusoidal signal which its  frequency increases over time.
-The chirp frequency increases linearly and sweeps entire bandwidth.
+A chirp is a sinusoidal signal which its frequency increases over time.
+Chirp frequency increases linearly and sweeps the entire bandwidth.
 We invite readers to refer to \cite{AN1200.22} for further details.
@@
 \subsection{Spreading Factor}
-\par Spreading factor (SF) is the ratio between symbol rate and chip rate.
-The ratio between symbol and chip rate is equal to $2$\textsuperscript{SF}.
+\par Spreading factor (SF) is the ratio between symbol rate and chirp rate.
+The ratio between symbol and chirp rate is equal to $2$\textsuperscript{SF}.
 SF can take values between 7 to 12.
@@
 \begin{equation} \label{eq:bit_rate_sf}
-R_{b}|_{bps} = SF * \dfrac{\left[ \dfrac{4}{4+CR} \right] }{ \left[ \dfrac{2^{SF}}{BW|_{Hz}} \right]}
+R_{b} = SF * \dfrac{\left[ \dfrac{4}{4+CR} \right] }{ \left[ \dfrac{2^{SF}}{BW|_{Hz}} \right]} \ bps
 \end{equation}
 
-Where, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.
+Where, $R_{b}$ is data rate in bps, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.
 
-\par While BW and CR are constant, as the SF increases the data rate decreases.
-Increasing SF makes the signal more resilient to noise thus increases the transmission range.
-Increasing SF also increases transmission duration which increases power consumption.
+\par When BW and CR are constant, as the SF increases, the data rate decreases.
+Increasing the SF makes the signal more resilient to noise thus increases the transmission range.
+Increasing the SF also increases the transmission duration which increases the power consumption.
 Therefore it is possible to trade between data rate, range and power consumption by changing SF.
@@
 
-\par Initially, a node is not aware of how far it is from a GW.
-However, a node can guess how far away from the gateway by observing received signal power of a downlink transmission.
+\par In a LoRaWAN network, initially, a node is not aware of how far it is from a GW.
+However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission.
 If a downlink transmission received signal power is too high then node can decrease its next transmission SF or transmit power to decrease power consumption.
-Also GW can request from a node to decrease node's SF or transmit power.
 This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper.
+Also a GW can request from a node to decrease node's SF or transmit power.
 
-\par In Figure \ref{fig:collision}, a LoRaWAN network distributed around a single GW is illustrated.
-Different color rings represent achievable range of different SF from GW and different color circles represent nodes with selected SF.
+\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated.
+Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes.
 Most of the end devices which close to the GW will fall into lowest SF (SF7) area section.
@@
 Which results a lot of collisions between same SF transmissions.
-Collision number increases while number of end devices close to the GW increases.
+Number of collisions increases while number of end devices close to the GW increases.
 
@@
 The literature related to the work presented in this paper has started growing recently.
-LPWAN technologies and especially LoRa/LoRaWAN attract researchers attention lately.
-We listed some of these works which studying LoRaWAN and LoRa SF.
+LPWAN technologies and especially LoRa  attracted researchers attentions lately.
+We listed some of these works which studies LoRaWAN and LoRa SF.
 
-\par In \cite{7996384}, the authors evaluated the performance of LoRa networks in a smart city scenario. The
-authors proposed a link measurement and a link performance model for LoRa.
+\par In \cite{7996384}, the authors evaluated the performance of LoRaWAN networks in a smart city scenario.
+The authors proposed a link measurement and a link performance model for LoRa.
 The authors also proposed a SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
-They implement a LoRa simulator in ns-3 to study scalability and performance of LoRaWAN networks.
+They implemented a LoRa simulator in ns-3 to study scalability and performance of LoRaWAN networks.
 Their results show that LoRaWAN networks scale well as the number of nodes and gateways increases.
-They also show that SF assignment has great effect on network performance.
+They also show that SF assignment has great effect on LoRaWAN network performance.
 
@@
 \par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements.
-They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from measurements.
+They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from their measurements.
 Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.
@@
 The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal.
-Their experimental results show that this power difference is around 16 dB and this power difference can be seen when  interferer is close to receiver or multiple interferer can create this power difference cumulatively.
+Their experimental results show that this power difference is around 16 dB and this power difference can be seen when a interferer is close to receiver or multiple interferer can create this power difference cumulatively.
 
-\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions using the same SF as well as different SFs.
-They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage performance with respect to the distance from the gateway for modeling interference in multiple gateway scenarios.
+\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions with same and different SFs.
+They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage  distance from the gateway for modeling interference in multiple gateway scenarios.
 Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.
@@
 \section{Proposed Technique} \label{Proposed Technique}
-\par The collision issue illustrated in Figure \ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even if they can able to communicate with lower SF.
+\par The collision issue illustrated in Figure \ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even they can able to communicate with lower SF.
 This may result lower collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}.
 Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}.
-However, how GW assign SF to which node becomes an another issue.
+However, how GW should assign SF to which node becomes an another issue.
 Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions.
-In multiple GW scenarios, this approach may increase the collision number with nodes in other GW's range.
+In multiple GW scenarios, this approach may increase the collisions with nodes in other GW's range.
 Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
@@
 \par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes SFs requires different rules.
-For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collision number.
+For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collisions.
 This technique learns nodes behavior of a network.
@@
 Using these obtained information NS can train a classifier for predicting transmission result for specific node and specific SF input.
-Using this predictor GWs can assign the lowest possible SF to nodes considering the collision probability.
+Using this prediction model GWs can assign the lowest possible SF to nodes considering the collision probability.
 
@@
 \par A discrete event simulator is implemented in Python to study LoRa network performance considering SF orthogonality.
-The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorafs.
+The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorasf.
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
@@
 Different channel transmissions are independent from each other.
-But, in this work we focuse on SF orthogonality.
+But, in this work we focus on SF orthogonality.
 Thus, we only utilized single channel transmissions in our simulator.
@@
 \begin{equation} \label{eq:transmission_duration}
-t_{x}|_{s} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}
+t_{x} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}  \ s
 \end{equation}
@@
 These data is fed to Python scikit-learn decision tree classifier.
-After the model is created, a second simulation is run with a predictor.
+After the model is created, a second simulation is run with a prediction scheme.
 Model tries to find the lowest possible SF with prediction of successful transmission.
@@
 
-\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning predictor SF scheme is shown.
+\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning prediction scheme is shown.
 Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second.
-Our predictor model needs nodes location so we simulate predictor SF scheme with at three GWs since three GWs is enough to roughly locate nodes positions by triangulation.
+Our prediction model needs nodes location so we simulate predictor SF scheme with at three GWs since three GWs is enough to roughly locate nodes positions by triangulation.
 Predictor scheme gives better PDR when number of nodes increases since number of interference increases when number of nodes increases.
@@
 \includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
-\caption{PDR of different prediction SF assignment scheme.}
+\caption{PDR of prediction SF assignment scheme.}
 \label{fig:predict_pdr}

### 2019-01-06 paper.tex
passive grammer

@@
 SF also determines data rate of a LoRa transmission \cite{AN1200.22}.
+Data rate of a LoRa transmission can be calculated as:
 
@@
 \section{Proposed Technique} \label{Proposed Technique}
-\par The collision issue illustrated in Figure \ref{fig:collision} can be solved if the GW force some of the close end nodes to select higher SF even they can able to communicate with lower SF.
-This may result lower collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}.
+\par The collision issue illustrated in Figure \ref{fig:collision} can be solved by forcing some of the close nodes to select higher SF even they can able to communicate with lower SF.
+This may prevent collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}.
 Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}.
-However, how GW should assign SF to which node becomes an another issue.
+However, how network should assign which SF to which node becomes an another issue.
 Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions.
@@
 For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collisions.
-This technique learns nodes behavior of a network.
-A NS can keep track of successful uplink transmission from nodes and their SFs.
-NS can also keep track of some of the collided transmission if header part of the packet does not interfered at GW.
-However, NS cannot keep track of transmissions with receive power lower than GW sensitivity.
-Using these obtained information NS can train a classifier for predicting transmission result for specific node and specific SF input.
-Using this prediction model GWs can assign the lowest possible SF to nodes considering the collision probability.
+This technique learns behavior of the nodes in a network.
+A NS can keep track of successful uplink transmissions and their SFs.
+NS can also keep track of some of the collided transmission if header part of the packet does not interfered at location of the GW.
+However, NS cannot keep track of transmissions with receive power lower than the GW sensitivity.
+Using these obtained information NS can train a classifier for predicting future transmission result for specific node and specific SF.
+Using this prediction model NS can assign the lowest possible SF to nodes considering the collision probability.
 
-\par In this work, we used decision tree classifier to predict transmission results.
-We utilized Gini index criteria to measure quality of decision tree splits and used balanced class weights.
-As most of the machine learning applications, increasing data set size increases prediction accuracy.
-For our case, data set size is directly proportional to simulation duration.
+\par In this work, a decision tree classifier is used to predict transmission results.
+Gini index criteria is used for measuring quality of decision tree splits.
+As in most of the machine learning applications, increasing the data set size increases the prediction accuracy.
+For in this case, data set size is directly proportional to simulation duration.
 Thus increasing simulation duration, increases the prediction accuracy.
-For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GW can request from nodes to use this new SF.
+For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use this new SF.
 
@@
 \includegraphics[width=\linewidth]{collision_solution_single_gw}
-\caption{Collision avoidance proposal between nodes close to the GW by using higher SF.}
+\caption{Collision avoidance by using higher SF for nodes close to the GW.}
 \label{fig:collision_solution_single_gw}
@@
 \includegraphics[width=\linewidth]{collision_solution_multi_gw}
-\caption{Collision avoidance proposal for intersecting GWs.}
+\caption{Collision avoidance for intersecting GWs.}
 \label{fig:collision_solution_multi_gw}
@@
 \section{Simulation Environment} \label{Simulation Environment}
-\par A discrete event simulator is implemented in Python to study LoRa network performance considering SF orthogonality.
+\par A discrete event simulator is implemented in Python to study effects of different SF strategies in LoRaWANs.
 The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorasf.
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
-Simulator can generate uniformly distributed random circle shape network topology with radius, number of nodes and number of gateways inputs.
-Global simulation parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
-With these input, the simulator provide outputs of total number of generated packets, number of successfully received packet, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
+Simulator can generate uniformly distributed circular shape network topology with radius (m), number of nodes and number of gateways input parameters.
+Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
+With these inputs, the simulator provide outputs of total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
 
-\par In our simulator we only cover LoRaWAN Class A devices.
+\par In the simulator, only LoRaWAN Class A devices are covered.
 Transmissions are always initiated by end nodes in pure ALOHA manner.
-Nodes generate a new packet according to Poisson interval for given packet rate input.
-In our simulator, downlink transmissions are not considered.
-Downlink transmissions are very rare in real world deployments since ISM band regulations dictates duty cycle transmission limit for all devices including GW.
-This makes downlink transmissions to very rare.
+Nodes generate a new packet according to Poisson interval for given packet rate input parameter.
+In the simulator, downlink transmissions are not considered.
+Downlink transmissions are rare in real world deployments since ISM band regulations dictates duty cycle transmission limit for all devices including GWs.
+This makes downlink transmissions to rare.
 
@@
 \par Link quality of a wireless system can be expressed by the metric of link budget.
-Link budget is a measure of all gains and losses from transmitter to receiver.
+Link budget is a measure of all gains and losses from transmitter device to receiver device.
 Link budget of a wireless link can be calculated as \cite{AN1200.22}:
@@
 $L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas in open space.
-In our simulation we assume that total of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dB.
+In the simulator, it is assumed that total of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dB.
 
@@
 \par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
-In our simulator we assume that nodes are always uses maximum allowed transmit power which is 14 dBm.
+In the simulator, it is assumed that nodes always select maximum allowed transmit power which is 14 dBm.
 Different channel transmissions are independent from each other.
 But, in this work we focus on SF orthogonality.
-Thus, we only utilized single channel transmissions in our simulator.
+Thus, we only utilized single channel transmissions in the simulator.
 
@@
 \centering
-\caption{Gateway Sensitivity for different SFs \cite{SX1276}.}
+\caption{Gateway Sensitivity for Different SFs \cite{SX1276}.}
 \label{table:gw_sf_sensitivity}
@@
 
-\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm can be found in Table \ref{table:gw_sf_sensitivity} \cite{SX1276}.
-In our simulation, we use 125 kHz bandwidth which is also the only allowed bandwidth in EU863-870 regulation.
+\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in Table \ref{table:gw_sf_sensitivity} \cite{SX1276}.
+In the simulator, 125 kHz bandwidth is used which is allowed bandwidth in EU863-870 regulation.
 
@@
 
-\par Where $h$ is the gateway altitude and $f$ is the frequency of the signal. We assume that $h$ = 15 m and $f$ = 868 MHz. With these assumptions, propagation loss calculation become \cite{7996384}:
+\par Where, $h$ is the gateway altitude and $f$ is the frequency of the signal.
+In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz.
+With these assumptions, propagation loss calculation become \cite{7996384}:
 
@@
 
-\par If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission at the transmission period.
+\par If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission at the duration of transmission.
 
 \subsection{Interference Model}
-\par We have adopted the interference model described in \cite{7996384} to our simulator.
+\par In the simulator, interference model described in \cite{7996384} is adopted.
 They introduced SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
 
-\par We assume that there is no other technology interference is present in the network except LoRa interference.
-To exploit imperfect orthogonality of different SF transmission, we need to calculate the effect of different SF transmissions to each other.
-We used signal to interference plus noise ratio (SINR) threshold matrix \cite{goursaud:hal-01231221}:
+\par In our simulator, it is assumed that there is no other technology interference is present in the network except LoRa interference.
+To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other.
+In simulator, following signal to interference plus noise ratio (SINR) threshold matrix from \cite{goursaud:hal-01231221} is used:
 
@@
 
-\par To decide if a referenced signal is interfered at receiver by a interferer signal, we use SINR threshold matrix.
-$T_{i,j}$ is the needed SINR margin in dB between referenced signal with SF = i and interferer signal with SF = j to correctly decode the referenced signal.
-If there are more than one interferer, referenced signal must satisfy the margin for cumulative sum of all interfering signal received power values for each SF \cite{7996384}.
+\par To decide if a referenced signal is interfered at receiver by a interferer signal, SINR threshold matrix is used.
+$T_{i,j}$ is the needed SINR margin in dB unit between referenced signal with SF = i and interferer signal with SF = j to correctly decode the referenced signal.
+If there are more than one interferer, referenced signal must satisfy the margin for cumulative sum of all interfering signal received powers for each SF \cite{7996384}.
+SINR threshold can be calculated as:
 
@@
 
-\par To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power value \cite{7996384}:
+\par To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power \cite{7996384}:
 
@@
 Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.
-Duration of a LoRa transmission:
+Duration of a LoRa transmission can be calculated as:
 
@@
 \subsection{Machine Learning Model}
-\par We have been able to generate mass amount of LoRaWAN transmission data for different topologies using our  tool.
-To come up with a solution for SF assignment issue, we have integrated a machine learning scheme to our simulator.
-This will allow us train a model from generated transmission data and we use this trained model to select best possible SF for nodes using prediction of the model.
-Tool first run a simulation with assigning random SF to nodes.
-Simulation results are combined into four columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
+\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator.
+To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator.
+A classifier is trained from transmission data and this classifier is used for selecting best possible SF for the nodes.
+Simulator first runs a simulation with assigning random SF to nodes.
+Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
 Transmission result can be successful, interfered or under sensitivity.
 These data is fed to Python scikit-learn decision tree classifier.
-After the model is created, a second simulation is run with a prediction scheme.
-Model tries to find the lowest possible SF with prediction of successful transmission.
+After the classifier model is generated, a second simulation is run with classifier.
+Simulator selects the lowest possible SF considering  prediction of transmission.
+If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF.
+If no SF is predicted as successful, then simulator again assigns lowest possible SF.
 
 \section{Simulation Results}  \label{Simulation Results}
-\par For simulation results in this paper, we set global simulation parameters as follows. Packet size is 60 bytes including header and payload. Simulation duration is 3600 seconds.
+\par For simulation results in this paper, global simulation parameters are set as follows. Packet size is 60 bytes including header and payload. Simulation duration is 3600 seconds.
 
@@
 Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second.
-Our prediction model needs nodes location so we simulate predictor SF scheme with at three GWs since three GWs is enough to roughly locate nodes positions by triangulation.
-Predictor scheme gives better PDR when number of nodes increases since number of interference increases when number of nodes increases.
-Predictor scheme can improve performance when LoRa interference is present.
-Also predictor scheme gives better results when nodes are deployed close to the gateway, since nodes have a margin to increase their SFs.
-If a node is far away from the GW, then predictor scheme cannot increase the SF to avoid interference since the SF is already high.
+Prediction model needs nodes location so three GWs is used for prediction SF scheme since three GWs is enough to roughly locate nodes positions by triangulation.
+Prediction scheme gives better PDR when number of nodes increases since number of interference is also increases.
+Prediction scheme can only improve network performance when LoRa interference is present.
+Also prediction scheme gives better results when nodes are deployed to close to the gateway, since nodes have margin to increase their SFs.
+If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.
 
@@
 \section{Conclusion} \label{Conclusion}
-\par In this paper, after brief introduction about LPWAN technologies, we provide background information about one of the most promising LPWAN technology, LoRa.
+\par In this paper, after a brief introduction about LPWAN technologies, we provide background information about one of the most promising LPWAN technology, LoRa.
 Then, we discuss about LoRa modulation basics and spreading factor assignment issue.
@@
 We illustrate simulation results for proposed machine learning SF assignment technique.
-Our simulation results show that, proposed technique can increase network performance for dense LoRaWAN network deployed close to GWs.
+Our simulation results show that, proposed technique can increase network performance for LoRaWAN networks when nodes are deployed close to the GWs.
 
 \par As for future works, transmit power optimization can be included to proposed machine learning scheme.
-In this paper, we assume nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
-This will make transmission more vulnerable to interference thus requires extra care.
+In this paper, we assume that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
+This will make transmissions to more vulnerable to interference thus requires extra care.
 Optimizing another node parameter alongside SF can be interesting research.
 Also other machine learning methods can be investigated for SF optimization.
-Reinforcement learning can be a good candidate for this network optimization issue.
+Reinforcement learning is a good candidate for this network optimization issue.
 

### 2019-01-09 paper.tex
grammer

@@
 
-\author{\IEEEauthorblockN{Tugrul Yatagan\IEEEauthorrefmark{1} and Sema F. Oktug\IEEEauthorrefmark{2}}
+\author{\IEEEauthorblockN{Tugrul Yatagan and Sema F. Oktug}
 \IEEEauthorblockA{Department of Computer Engineering,
@@
 Istanbul, Turkey\\
-Email: \IEEEauthorrefmark{1}yatagan@itu.edu.tr, \IEEEauthorrefmark{2}oktug@itu.edu.tr}}
+Email: \{yatagan, oktug\}@itu.edu.tr}}
 \maketitle
@@
 In this work, a simulation environment to study different SF assignment schemes is implemented.
-Furthermore, novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization.
+Furthermore, a novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization of SF assignment.
 Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
@@
 Recent developments on LPWAN technologies has great impact on growth of number of IoT applications.
-LPWAN technologies addresses some of the well-known wireless communication challenges.
+LPWAN technologies address some of the well-known wireless communication challenges.
 Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication technologies (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time.
-Cellular networks can provide long range and high data rate but they are complex and consume too much power.
+Cellular networks can provide long range and high data rate, but they are complex and consume too much power.
 Besides, most of the IoT applications do not require high data rate.
-Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}.
-LPWAN technologies fill the technology gap between short range and cellular technologies by providing low power and long range communication.
+Short-range communication methods can provide relatively low power consumption, but their range is limited to a few hundred meters at best \cite{7815384}.
+LPWAN technologies fill the technology gap between short range and cellular technologies by providing low power and long-range communication.
 LPWAN technologies basically sacrifice data rate to provide low power consumption.
@@
 LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}.
-Both LoRa and Sigfox are known for ultra low power consumption and resilience to interference.
+Both LoRa and Sigfox are known for ultra-low power consumption and resilience to interference.
 While NB-IoT and LTE-M are promoted for higher data rate.
@@
 Also, Sigfox packet payload size is limited to 12 bytes for uplink and 8 bytes for downlink.
-However LoRa supports up to 243 bytes payload size and NB-IoT supports up to 1600 bytes payload size.
+However, LoRa supports up to 243 bytes payload size and NB-IoT supports up to 1600 bytes payload size.
 Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.
@@
 LoRa supports 6 different SF option.
-Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other up to some extent, thus SF assignment is crucial for overall network performance.
+Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other up to some extent, thus SF assignment is crucial for overall network performance.
 In this work, a custom made LoRa discrete event simulator is introduced to study different LoRa SF assignment strategies.
@@
 
-\par The following two sections provides an background information about LoRaWAN and LoRa.
+\par The following two sections provides a background information about LoRaWAN and LoRa.
 Section \ref{Other Related Works} describes other related works.
@@
 \subsection{End node}
-\par LoRaWAN end node (EN) is a low power embedded device that only communicates between gateways.
+\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways.
 LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
@@
 Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission.
-Receive window is synchronized using beacon packet transmitted by gateways.
+Receive window is synchronized using a beacon packet transmitted by gateways.
 Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
-This provides Class C end nodes to low latency donwlink communication but it requires more power consumption.
+This provides Class C end nodes to low latency downlink communication, but it requires more power consumption.
 In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.
@@
 A typical gateway can receive from multiple channels at the same time.
-Gateways are typically connected to power grid so power consumption of a gateway is insignificant in most of the deployments.
+Gateways are typically connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.
 
@@
 Increasing the SF also increases the transmission duration which increases the power consumption.
-Therefore it is possible to trade between data rate, range and power consumption by changing SF.
+Therefore, it is possible to trade between range and power consumption by changing SF.
 
@@
 However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission.
-If a downlink transmission received signal power is too high then node can decrease its next transmission SF or transmit power to decrease power consumption.
+If a downlink transmission received signal power is too high, then node can decrease its next transmission SF to decrease power consumption.
 This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper.
-Also a GW can request from a node to decrease node's SF or transmit power.
+Also, a GW can request from a node to decrease node's SF or transmit power.
 
@@
 End devices which close to the GW will probably select lowest SF all the time.
-Which results a lot of collisions between same SF transmissions.
+This results a lot of collisions between same SF transmissions.
 Number of collisions increases while number of end devices close to the GW increases.
@@
 \section{Other Related Works} \label{Other Related Works}
-The literature related to the work presented in this paper has started growing recently.
-LPWAN technologies and especially LoRa  attracted researchers attentions lately.
+\par The literature related to the work presented in this paper has started growing recently.
+LPWAN technologies and especially LoRa attracted researchers’ attentions lately.
 We listed some of these works which studies LoRaWAN and LoRa SF.
@@
 They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from their measurements.
-Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the the last six symbols of preamble and header of the packet does not collide.
+Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the last six symbols of preamble and header of the packet does not collide.
 
@@
 The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal.
-Their experimental results show that this power difference is around 16 dB and this power difference can be seen when a interferer is close to receiver or multiple interferer can create this power difference cumulatively.
+Their experimental results show that this power difference is around 16 dB and this power difference can be seen when an interferer is close to a receiver or sum of multiple interferer energy can create this power difference.
 
@@
 Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}.
-However, how network should assign which SF to which node becomes an another issue.
+However, how network should assign which SF to which node becomes another issue.
 Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions.
 In multiple GW scenarios, this approach may increase the collisions with nodes in other GW's range.
-Thus, extra care must be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
+Thus, extra care should be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
 
@@
 A NS can keep track of successful uplink transmissions and their SFs.
-NS can also keep track of some of the collided transmission if header part of the packet does not interfered at location of the GW.
+NS can also keep track of some of the collided transmission if header part of the packet is not interfered at location of the GW.
 However, NS cannot keep track of transmissions with receive power lower than the GW sensitivity.
-Using these obtained information NS can train a classifier for predicting future transmission result for specific node and specific SF.
+Using these obtained information, NS can train a classifier for predicting future transmission result for specific node and specific SF.
 Using this prediction model NS can assign the lowest possible SF to nodes considering the collision probability.
@@
 For in this case, data set size is directly proportional to simulation duration.
-Thus increasing simulation duration, increases the prediction accuracy.
+Thus, increasing simulation duration, increases the prediction accuracy.
 For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use this new SF.
@@
 \par A discrete event simulator is implemented in Python to study effects of different SF strategies in LoRaWANs.
-The LoRaWAN SF simulation tool is publicly available at https://github.com/tugrulyatagan/simlorasf.
+The LoRaWAN SF simulation tool source code is available at \cite{simlorasf}.
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
@@
 Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
-With these inputs, the simulator provide outputs of total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
+With these inputs, the simulator provides outputs of total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
 
@@
 To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator.
-A classifier is trained from transmission data and this classifier is used for selecting best possible SF for the nodes.
+A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes.
 Simulator first runs a simulation with assigning random SF to nodes.
@@
 Transmission result can be successful, interfered or under sensitivity.
-These data is fed to Python scikit-learn decision tree classifier.
+These data are fed to Python scikit-learn decision tree classifier.
 After the classifier model is generated, a second simulation is run with classifier.
@@
 Randomly generated network topology radius is set to 3000 meters, number of GW is set to 1 and packet generation rate is set to 0.01 packet per second.
-Increasing SFs, increases time on air.
+Increasing SFs increases time on air.
 This increases the number of collisions thus decreases the PDR of the network.
-PDR of higher SFs are decreased to almost 0 while number of node is increasing.
+PDR of higher SFs are decreased to almost 0 while number of nodes is increasing.
 Since network topology radius is quite small almost all SFs can reach to the GW.
@@
 Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second.
-Prediction model needs nodes location so three GWs is used for prediction SF scheme since three GWs is enough to roughly locate nodes positions by triangulation.
-Prediction scheme gives better PDR when number of nodes increases since number of interference is also increases.
+Prediction model needs nodes location so three GWs are used for prediction SF scheme since three GWs are enough to roughly locate nodes positions by triangulation.
+Prediction scheme gives better PDR when number of nodes increases since number of interferences also increases.
 Prediction scheme can only improve network performance when LoRa interference is present.
-Also prediction scheme gives better results when nodes are deployed to close to the gateway, since nodes have margin to increase their SFs.
+Also, prediction scheme gives better results when nodes are deployed close to the gateway, since nodes have margin to increase their SFs.
 If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.
@@
 \includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
-\caption{PDR of prediction SF assignment scheme.}
+\caption{PDR of SF prediction scheme.}
 \label{fig:predict_pdr}
@@
 Optimizing another node parameter alongside SF can be interesting research.
-Also other machine learning methods can be investigated for SF optimization.
+Also, other machine learning methods can be investigated for SF optimization.
 Reinforcement learning is a good candidate for this network optimization issue.

### 2019-01-13 paper.tex
SVM pdr
update figures
classifier accuracy

@@
 
-\par The following two sections provides a background information about LoRaWAN and LoRa.
+\par The following two sections provides a background information about LoRa and LoRaWAN.
 Section \ref{Other Related Works} describes other related works.
@@
 Finally, Section \ref{Conclusion} concludes the paper.
-
-\section{LoRaWAN} \label{LoRaWAN}
-\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently yet incorrectly used for each other.
-
-\par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
-LoRaWAN is an open source standard developed and it is maintained by LoRa Alliance.
-LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
-LoRa can be used as a wireless link technology without complying LoRaWAN, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions.
-LoRaWAN also provides inter-operability between different LoRa networks.
-LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.
-LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmitting, accepting the possibility of the collision.
-
-\par A typical LoRaWAN network consist of three network entities.
-
-\subsection{End node}
-\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways.
-LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
-Different classes provides LPWAN solutions to different deployments and topologies.
-Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission.
-Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission.
-Receive window is synchronized using a beacon packet transmitted by gateways.
-Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
-This provides Class C end nodes to low latency downlink communication, but it requires more power consumption.
-In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.
-
-\subsection{Gateway}
-\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes.
-A typical gateway can receive from multiple channels at the same time.
-Gateways are typically connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.
-
-\subsection{Network server}
-\par LoRaWAN network server (NS) is a server that provides MAC layer processing.
-Network server routes messages from application to end nodes and vice versa.
-Network server can be used for tweaking end node parameters like channel, transmit power and SF to increase network performance.
 
@@
 \end{figure}
+
+\section{LoRaWAN} \label{LoRaWAN}
+\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently yet incorrectly used for each other.
+
+\par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
+LoRaWAN is an open source standard developed and it is maintained by LoRa Alliance.
+LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
+LoRa can be used as a wireless link technology without complying LoRaWAN, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions.
+LoRaWAN also provides inter-operability between different LoRa networks.
+LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.
+LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmitting, accepting the possibility of the collision.
+
+\par A typical LoRaWAN network consist of three network entities.
+
+\subsection{End node}
+\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways.
+LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
+Different classes provides LPWAN solutions to different deployments and topologies.
+Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission.
+Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission.
+Receive window is synchronized using a beacon packet transmitted by gateways.
+Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
+This provides Class C end nodes to low latency downlink communication, but it requires more power consumption.
+In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.
+
+\subsection{Gateway}
+\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes.
+A typical gateway can receive from multiple channels at the same time.
+Gateways are typically connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.
+
+\subsection{Network server}
+\par LoRaWAN network server (NS) is a server that provides MAC layer processing.
+Network server routes messages from application to end nodes and vice versa.
+Network server can be used for tweaking end node parameters like channel, transmit power and SF to increase network performance.
 
@@
 For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use this new SF.
+
+\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator.
+To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator.
+A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes.
+Simulator first runs a simulation with assigning random SF to nodes.
+Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
+Transmission result can be successful, interfered or under sensitivity.
+These data are fed to Python scikit-learn decision tree classifier.
+After the classifier model is generated, a second simulation is run with classifier.
+Simulator selects the lowest possible SF considering prediction of transmission.
+If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF.
+If no SF is predicted as successful, then simulator again assigns lowest possible SF.
 
@@
 
-\subsection{Machine Learning Model}
-\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator.
-To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator.
-A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes.
-Simulator first runs a simulation with assigning random SF to nodes.
-Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
-Transmission result can be successful, interfered or under sensitivity.
-These data are fed to Python scikit-learn decision tree classifier.
-After the classifier model is generated, a second simulation is run with classifier.
-Simulator selects the lowest possible SF considering prediction of transmission.
-If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF.
-If no SF is predicted as successful, then simulator again assigns lowest possible SF.
-
 \section{Simulation Results}  \label{Simulation Results}
@@
 
-\par In Figure \ref{fig:predict_pdr}, PDR plot of machine learning prediction scheme is shown.
+\par In Figure \ref{fig:prediction_pdr}, PDR plot of machine learning prediction scheme is shown.
 Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second.
@@
 \centering
-\includegraphics[width=\linewidth]{{predict_pdr_r5000_g3_p0.01_s3600}.png}
+\includegraphics[width=\linewidth]{{prediction_pdr_r5000_g3_p0.01_s3600}.png}
 \caption{PDR of SF prediction scheme.}
-\label{fig:predict_pdr}
+\label{fig:prediction_pdr}
 \end{figure}
+
+\begin{table}
+\centering
+\caption{PDR of lowest SF selection and different SF classifier schemes.}
+\label{table:prediction_pdr}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+\multicolumn{3}{|c|}{\multirow{2}{*}{}}                            & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
+\multicolumn{3}{|c|}{}                                             & 100           & 500           & 1000          \\ \hline
+\multirow{12}{*}{\textbf{R (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
+                                 &                        & SVM    & 98.0          & 88.2          & 75.2          \\ \cline{3-6}
+                                 &                        & DTC    & 97.8          & 89.8          & 78.7          \\ \cline{2-6}
+
+                                 & \multirow{3}{*}{5000}  & Lowest & 96.8          & 85.5          & 71.2          \\ \cline{3-6}
+                                 &                        & SVM    & 98.0          & 87.8          & 74.8          \\ \cline{3-6}
+                                 &                        & DTC    & 97.7          & 90.2          & 79.8          \\ \cline{2-6}
+
+                                 & \multirow{3}{*}{7000}  & Lowest & 97.2          & 87.5          & 76.8          \\ \cline{3-6}
+                                 &                        & SVM    & 98.2          & 88.8          & 78.6          \\ \cline{3-6}
+                                 &                        & DTC    & 97.8          & 90.7          & 81.6          \\ \cline{2-6}
+
+                                 & \multirow{3}{*}{10000} & Lowest & 98.2          & 90.3          & 81.5          \\ \cline{3-6}
+                                 &                        & SVM    & 98.3          & 90.3          & 81.9          \\ \cline{3-6}
+                                 &                        & DTC    & 98.3          & 90.6          & 81.9          \\ \hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+\centering
+\caption{Prediction accuracy (\%) of different classifiers.}
+\label{table:prediction_accuracy}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+\multicolumn{3}{|c|}{\multirow{2}{*}{}}                        & \multicolumn{3}{c|}{\textbf{Number of Nodes}} \\ \cline{4-6}
+\multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
+\multirow{8}{*}{\textbf{R (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
+                                &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}
+
+                                & \multirow{2}{*}{5000}  & SVM & 79.5          & 69.0          & 71.1          \\ \cline{3-6}
+                                &                        & DTC & 84.5          & 67.3          & 69.5          \\ \cline{2-6}
+
+                                & \multirow{2}{*}{7000}  & SVM & 79.5          & 70.6          & 71.2          \\ \cline{3-6}
+                                &                        & DTC & 84.5          & 67.7          & 69.2          \\ \cline{2-6}
+
+                                & \multirow{2}{*}{10000} & SVM & 79.2          & 74.4          & 76.1          \\ \cline{3-6}
+                                &                        & DTC & 83.8          & 70.7          & 74.3          \\ \hline
+\end{tabular}
+\end{table}
 
@@
 This will make transmissions to more vulnerable to interference thus requires extra care.
-Optimizing another node parameter alongside SF can be interesting research.
 Also, other machine learning methods can be investigated for SF optimization.
 Reinforcement learning is a good candidate for this network optimization issue.
+TODO, node id without location TODO
 

### 2019-01-13 paper.tex
omit lora details

@@
 However, Sigfox and NB-IoT are only available with operator contract \cite{7815384}.
-Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink.
-Also, Sigfox packet payload size is limited to 12 bytes for uplink and 8 bytes for downlink.
-However, LoRa supports up to 243 bytes payload size and NB-IoT supports up to 1600 bytes payload size.
-Sigfox maximum data rate is 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.
 
@@
 This enables tradeoff between receive sensitivity and transmission time on air \cite{7803607}.
-LoRa supports 6 different SF option.
-Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other up to some extent, thus SF assignment is crucial for overall network performance.
-In this work, a custom made LoRa discrete event simulator is introduced to study different LoRa SF assignment strategies.
+Simultaneous same spreading factor (SF) transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
+In this work, a LoRa discrete event simulator is developed to study different LoRa SF assignment strategies.
 We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using the simulator.
@@
 \section{LoRaWAN} \label{LoRaWAN}
-\par LoRa is a wireless link technology for LPWAN applications. LoRa has an open standard MAC protocol called LoRaWAN. LoRa and LoRaWAN terms are frequently yet incorrectly used for each other.
-
-\par LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
-LoRaWAN is an open source standard developed and it is maintained by LoRa Alliance.
+\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN.
+LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
+LoRaWAN is developed and maintained by LoRa Alliance.
 LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
-LoRa can be used as a wireless link technology without complying LoRaWAN, however LoRaWAN is developed considering well known LPWAN challenges and their best practice solutions.
-LoRaWAN also provides inter-operability between different LoRa networks.
-LoRaWAN provides lightweight but powerful standard for wide range of LoRa IoT applications.
+LoRaWAN  provides inter-operability between different LoRa networks.
 LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmitting, accepting the possibility of the collision.
@@
 
-\par In \cite{s17061193}, the authors investigated single gateway LoRaWAN network scalability in terms of the number of end nodes using a simulation model based on real measurements.
-They measure the impact of two concurrent LoRa transmissions on each other by using physical LoRaWAN end devices and a gateway then they created a simulation model from their measurements.
-Their results show that LoRaWAN has better scalability than pure ALOHA since a LoRa packet may still go through under collision if the last six symbols of preamble and header of the packet does not collide.
-
 \par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa SF transmissions.
@@
 
-\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes SFs requires different rules.
+\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules.
 For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collisions.
-This technique learns behavior of the nodes in a network.
+This technique works by learning the behavior of the nodes in a network.
 A NS can keep track of successful uplink transmissions and their SFs.
@@
 Using these obtained information, NS can train a classifier for predicting future transmission result for specific node and specific SF.
-Using this prediction model NS can assign the lowest possible SF to nodes considering the collision probability.
+Using this prediction model NS can assign the SF to nodes considering the collision probability.
 
-\par In this work, a decision tree classifier is used to predict transmission results.
-Gini index criteria is used for measuring quality of decision tree splits.
-As in most of the machine learning applications, increasing the data set size increases the prediction accuracy.
-For in this case, data set size is directly proportional to simulation duration.
-Thus, increasing simulation duration, increases the prediction accuracy.
-For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use this new SF.
+\par In this work, decision tree classifier (DTC) and support vector machine (SVM) are evaluated to predict transmission results.
+Gini index criteria is used for DTC splits.
+Class weights are balanced according to sample distributions for both methods.
+Data set size is directly proportional to simulation duration.
+Thus, increasing simulation duration, improves the prediction accuracy.
+For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use their new SFs.
 
 \par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator.
-To come up with a solution for SF assignment issue, a machine learning prediction scheme is integrated to the simulator.
+To come up with a solution for SF assignment issue, DTC and SCM prediction schemes are integrated to the simulator.
 A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes.
-Simulator first runs a simulation with assigning random SF to nodes.
+Simulator first runs a simulation with assigning random SF to the nodes.
 Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
 Transmission result can be successful, interfered or under sensitivity.
-These data are fed to Python scikit-learn decision tree classifier.
+These data are fed to Python scikit-learn DTC or SVM classifiers.
 After the classifier model is generated, a second simulation is run with classifier.
-Simulator selects the lowest possible SF considering prediction of transmission.
-If a transmission of a node with lowest possible SF is predicted as interfered transmission by classifier, then simulator assigns a higher SF.
-If no SF is predicted as successful, then simulator again assigns lowest possible SF.
+Simulator selects the most suitable SF considering prediction of transmission.
+If a transmission of a node with lowest possible SF is predicted as interfered by the classifier, then simulator assigns a higher SF.
+If no SF is predicted as successful, then simulator again assigns the lowest possible SF.
+
+TODOOOOO
 
@@
 In the simulator, downlink transmissions are not considered.
-Downlink transmissions are rare in real world deployments since ISM band regulations dictates duty cycle transmission limit for all devices including GWs.
-This makes downlink transmissions to rare.
+Downlink transmissions are rare in real world deployments since ISM band regulations dictates duty cycle limit for all devices including GWs.
 
@@
 
-\begin{table}
-\centering
-\caption{EU863-870 ISM Band LoRa Default Channels \cite{lorawan.regional.parameters}.}
-\label{table:max_tx_power}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Channel} & \textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
-      1 &     868.1 &   125 &   14 \\ \hline
-      2 &     868.3 &   125 &   14 \\ \hline
-      3 &     868.5 &   125 &   14 \\
-
-\hline
-\end{tabular}
-\end{table}
-
-\par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
-In the simulator, it is assumed that nodes always select maximum allowed transmit power which is 14 dBm.
+\par In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm for European ISM band regulation.
 Different channel transmissions are independent from each other.
@@
 
-\begin{table}
-\centering
-\caption{Gateway Sensitivity for Different SFs \cite{SX1276}.}
-\label{table:gw_sf_sensitivity}
-\begin{tabular}{|c|c|c|c|c|c|c|c|}
-\hline
-\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
-\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
-\multirow{3}{*}{\textbf{BW (kHz)}}  & 125 & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
-                                    & 250 & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
-                                    & 500 & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
-\end{tabular}
-\end{table}
+\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.
 
-\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in Table \ref{table:gw_sf_sensitivity} \cite{SX1276}.
-In the simulator, 125 kHz bandwidth is used which is allowed bandwidth in EU863-870 regulation.
-
-\par Free space propagation loss can be calculated as \cite{TR136.942}:
-
-\begin{equation} \label{eq:propagation_loss}
-\begin{split}
-P^{dB}_{PATH} = 40(1 - 4 \times 10^{-3} \times h){\log_{10} R|_{km}} \\
-- 18 {\log_{10} h|_{m}} + 21 {\log_{10} f|_{MHz}} + 80
-\end{split}
-\end{equation}
-
-\par Where, $h$ is the gateway altitude and $f$ is the frequency of the signal.
-In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz.
-With these assumptions, propagation loss calculation become \cite{7996384}:
-
-\begin{equation} \label{eq:propagation_loss_simplified}
-P^{dB}_{PATH} = 120.5 + 37.6 {\log_{10} R|_{km}}
-\end{equation}
+\par Free space propagation loss is calculated as in \cite{TR136.942}. In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz.
 
@@
 \par In the simulator, interference model described in \cite{7996384} is adopted.
-They introduced SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
+They use SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
 
-\par In our simulator, it is assumed that there is no other technology interference is present in the network except LoRa interference.
+\par In our simulator, it is assumed that there is no other technology interference in the network except LoRa interference.
 To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other.
-In simulator, following signal to interference plus noise ratio (SINR) threshold matrix from \cite{goursaud:hal-01231221} is used:
-
-\begin{equation} \label{eq:sinr}
-T = \begin{bmatrix}
-       6 & -16 & -18 & -19 & -19 & -20 \\
-     -24 &   6 & -20 & -22 & -22 & -22 \\
-     -27 & -27 &   6 & -23 & -25 & -25 \\
-     -30 & -30 & -30 &   6 & -26 & -28 \\
-     -33 & -33 & -33 & -33 &   6 & -29 \\
-     -36 & -36 & -36 & -36 & -36 &   6
-     \end{bmatrix}
-\end{equation}
+In simulator, following signal to interference plus noise ratio (SINR) threshold matrix from \cite{goursaud:hal-01231221} is used.
 
@@
 If there are more than one interferer, referenced signal must satisfy the margin for cumulative sum of all interfering signal received powers for each SF \cite{7996384}.
-SINR threshold can be calculated as:
-
-\begin{equation} \label{eq:sinr_db}
-SINR_{i,j} = \dfrac{P_{rc,0}}{\sum_{l \in I_j} P_{rc,l}}
-\end{equation}
-
-\par Where $P_{rc,0}$ is received signal power of referenced signal and $P_{rc,l}$ is received signal power of interfering signal at SF = j.
-If a packet with SF = i satisfies the following condition for every SF = j, then packet is survived from all interference.
-
-\begin{equation} \label{eq:sinr_t}
-SINR_{i,j}^{dB} > T_{i,j}
-\end{equation}
-
-\par To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power \cite{7996384}:
-
-\begin{equation} \label{eq:p_interference}
-P_{rc,y}^{interf} = \dfrac{P_{rc,y}(t_{interf})}{t_{x}}
-\end{equation}
-
-\par Where $t_{x}$ is transmission duration of referenced signal.
-$t_{interf}$ is intersection duration between referenced signal and interfering signal.
 Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$.
 Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.
-Duration of a LoRa transmission can be calculated as:
-
-\begin{equation} \label{eq:transmission_duration}
-t_{x} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}  \ s
-\end{equation}
 
 \section{Simulation Results}  \label{Simulation Results}
-\par For simulation results in this paper, global simulation parameters are set as follows. Packet size is 60 bytes including header and payload. Simulation duration is 3600 seconds.
+\par For simulation results in this paper, global simulation parameters are set as follows.
+Packet size is 60 bytes including header and payload, simulation duration is 3600 seconds, packet generation rate is 0.01 packet per second.
 
 \par In Figure \ref{fig:sf_pdr}, PDR plot of different SF assignment is shown.
-Randomly generated network topology radius is set to 3000 meters, number of GW is set to 1 and packet generation rate is set to 0.01 packet per second.
+Randomly generated network topology radius is set to 3000 meters and number of GW is set to 1.
 Increasing SFs increases time on air.
@@
 \par In Figure \ref{fig:gw_pdr}, PDR plot of different number of GW is shown.
-Randomly generated topology radius is set to 3000 meters, packet generation rate is set to 0.01 packet per second and lowest possible SF assignment scheme is used.
+Randomly generated topology radius is set to 3000 meters and lowest possible SF assignment scheme is used.
 Increasing number of GWs, decreases the SFs of nodes and decreases time on air.
@@
 \par In Figure \ref{fig:r_pdr}, PDR plot of different network topology radius is shown.
-Number of GW is set to 1, packet generation rate is set to 0.01 packet per second and lowest possible SF assignment scheme is used.
+Number of GW is set to 1 and lowest possible SF assignment scheme is used.
 Increasing network topology radius, increases the number of under sensitivity transmissions thus decreases the PDR of the network.
@@
 
-\par In Figure \ref{fig:pr_pdr}, PDR plot of different packet generation rate is shown.
-Randomly generated topology radius is set to 3000 meters, number of GW is set to 1 and lowest possible SF assignment scheme is used.
-Increasing packet generation rate, increases the number of collisions thus decreases the PDR of the network.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{{pr_pdr_r3000_g1_s3600}.png}
-\caption{PDR of different packet rates.}
-\label{fig:pr_pdr}
-\end{figure}
+\par For space constraints, we omit figures for network throughput output, total transmit energy consumption output and different packet generation rate input.
+We invite readers to experiment the simulation tool with different simulation inputs and outputs.
+\cite{simlorasf}.
 
 \par In Figure \ref{fig:prediction_pdr}, PDR plot of machine learning prediction scheme is shown.
-Randomly generated network topology radius is set to 5000 meters, number of GW is set to 3 and packet generation rate is set to 0.01 packet per second.
+Randomly generated network topology radius is set to 5000 meters and number of GW is set to 3.
 Prediction model needs nodes location so three GWs are used for prediction SF scheme since three GWs are enough to roughly locate nodes positions by triangulation.
@@
 
+\par In Table \ref{table:prediction_pdr}, TODO
+
+\par In Table \ref{table:prediction_accuracy}, TODO
+
 \begin{table}
 \centering
-\caption{PDR of lowest SF selection and different SF classifier schemes.}
+\caption{PDR of lowest possible SF and smart SF assignment schemes.}
 \label{table:prediction_pdr}
@@
 \centering
-\caption{Prediction accuracy (\%) of different classifiers.}
+\caption{Prediction accuracy (\%) of SVM and DTC classifiers.}
 \label{table:prediction_accuracy}
@@
 Reinforcement learning is a good candidate for this network optimization issue.
-TODO, node id without location TODO
+Other transmission parameters such as node id and transmission time can be included to proposed scheme to improve performance.
 

### 2019-01-13 paper.tex
prediction accuracy par

@@
 \begin{IEEEkeywords}
-LoRa, Spreading Factor, IoT, LPWAN, Machine Learning
+LoRa, LoRaWAN, Spreading Factor, LPWAN, Machine Learning
 \end{IEEEkeywords}
@@
 \par For space constraints, we omit figures for network throughput output, total transmit energy consumption output and different packet generation rate input.
-We invite readers to experiment the simulation tool with different simulation inputs and outputs.
+We invite readers to experiment the simulation tool with different inputs and outputs.
 \cite{simlorasf}.
@@
 If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.
+\ref{table:prediction_pdr} TODO
 
@@
 \end{figure}
-
-\par In Table \ref{table:prediction_pdr}, TODO
-
-\par In Table \ref{table:prediction_accuracy}, TODO
 
@@
 \end{table}
+
+\par In Table \ref{table:prediction_accuracy}, prediction accuracy of SVM and DTC prediction schemes are shown for different network topology radius and number of nodes. Number of GW is set to 3. TODO
 

### 2019-01-14 paper.tex
pdr table par

@@
 
-\par In this work, decision tree classifier (DTC) and support vector machine (SVM) are evaluated to predict transmission results.
-Gini index criteria is used for DTC splits.
+\par In this work, decision tree classifier (DTC) and support vector machine (SVM) classifier are evaluated to predict transmission results.
 Class weights are balanced according to sample distributions for both methods.
+For DTC, Gini impurity criteria is used to measure the quality of a splits.
+For SVM, penalty parameter is set to 1, degree is set to 3 and RBF kernel is used.
+
+\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator.
 Data set size is directly proportional to simulation duration.
 Thus, increasing simulation duration, improves the prediction accuracy.
-For real world cases, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use their new SFs.
+In real world deployments, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use their new SFs.
 
-\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator.
-To come up with a solution for SF assignment issue, DTC and SCM prediction schemes are integrated to the simulator.
-A classifier is trained from transmission logs and this classifier is used for selecting best possible SF for the nodes.
+\par To come up with a solution for SF assignment issue, DTC and SCM prediction schemes are integrated to the simulator.
+A classifier is trained from transmission logs and this classifier is used for selecting optimum SF for the nodes.
 Simulator first runs a simulation with assigning random SF to the nodes.
 Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
-Transmission result can be successful, interfered or under sensitivity.
+Possible transmission results are successful, interfered or under sensitivity.
 These data are fed to Python scikit-learn DTC or SVM classifiers.
 After the classifier model is generated, a second simulation is run with classifier.
-Simulator selects the most suitable SF considering prediction of transmission.
-If a transmission of a node with lowest possible SF is predicted as interfered by the classifier, then simulator assigns a higher SF.
+Simulator selects optimum SF considering prediction of transmission.
+If a transmission of a node with lowest possible SF is predicted as interfered by the classifier, then simulator assigns higher SF and calculates new prediction.
 If no SF is predicted as successful, then simulator again assigns the lowest possible SF.
-
-TODOOOOO
 
@@
 Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
-With these inputs, the simulator provides outputs of total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
+With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
+Simulator also produces  prediction accuracy and confusion matrix for machine learning schemes.
 
@@
 \par In Figure \ref{fig:sf_pdr}, PDR plot of different SF assignment is shown.
-Randomly generated network topology radius is set to 3000 meters and number of GW is set to 1.
+Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1.
 Increasing SFs increases time on air.
@@
 
-\par In Figure \ref{fig:gw_pdr}, PDR plot of different number of GW is shown.
+\par In Figure \ref{fig:gw_pdr}, PDR plot of different number of GWs is shown.
 Randomly generated topology radius is set to 3000 meters and lowest possible SF assignment scheme is used.
@@
 
-\par In Figure \ref{fig:prediction_pdr}, PDR plot of machine learning prediction scheme is shown.
-Randomly generated network topology radius is set to 5000 meters and number of GW is set to 3.
-Prediction model needs nodes location so three GWs are used for prediction SF scheme since three GWs are enough to roughly locate nodes positions by triangulation.
-Prediction scheme gives better PDR when number of nodes increases since number of interferences also increases.
-Prediction scheme can only improve network performance when LoRa interference is present.
-Also, prediction scheme gives better results when nodes are deployed close to the gateway, since nodes have margin to increase their SFs.
-If a node is far away from the GW, then prediction scheme cannot increase the SF to avoid interference since the SF is already high.
-\ref{table:prediction_pdr} TODO
+\par In Figure \ref{fig:prediction_pdr}, result of proposed technique is presented.
+PDR plot of lowest SF scheme and machine learning prediction schemes are shown in same figure.
+Randomly generated network topology radius is set to 5000 meters and number of GWs is set to 3.
+Prediction model needs nodes location and three GWs are enough to roughly locate nodes positions by triangulation.
+In Table \ref{table:prediction_pdr}, PDR values for different topology radius are presented.
+Both prediction SF schemes give better PDR than lowest SF schemes when number of nodes increases since number of interference also increases.
+Prediction schemes can only improve network performance when LoRa interference is present.
+Also, prediction schemes give better results when nodes are deployed close to the gateway, since nodes have margin to increase their SFs.
+If a node is far away from the GW, then prediction schemes cannot increase the SF to avoid interference since the SF is already high.
+
+In Table \ref{table:prediction_accuracy}, prediction accuracy of SVM and DTC prediction schemes are presented for different network topology radius and number of nodes. Number of GW is set to 3.
 
@@
 
-\par In Table \ref{table:prediction_accuracy}, prediction accuracy of SVM and DTC prediction schemes are shown for different network topology radius and number of nodes. Number of GW is set to 3. TODO
-
 \begin{table}
 \centering
-\caption{Prediction accuracy (\%) of SVM and DTC classifiers.}
+\caption{Prediction accuracy of SVM and DTC classifiers.}
 \label{table:prediction_accuracy}

### 2019-01-14 paper.tex
shortened

@@
 \begin{abstract}
-Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over very large geographical areas.
+Low power wide area network (LPWAN) technologies offer affordable connectivity to massive number of low-power devices distributed over large geographical areas.
 Focus of this work is one of the most promising LPWAN technologies: LoRa.
@@
 This enables tradeoff between receive sensitivity and transmission time on air \cite{7803607}.
-Simultaneous same spreading factor (SF) transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
+Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
 In this work, a LoRa discrete event simulator is developed to study different LoRa SF assignment strategies.
@@
 \section{LoRa}
-\par LoRa is the name of the proprietary physical layer radio/chipset technology that provides wireless link solution for low power wide area networks.
+\par LoRa is the name of the proprietary physical layer radio/chipset technology that provides wireless link solution for LPWANs.
 LoRa is designed and patented by Semtech Corporation.
@@
 \par Simultaneous different SF transmissions are orthogonal to each other up to some extent.
-Which means, a LoRa gateway can simultaneously receive multiple transmissions with different SFs.
-However, simultaneous transmissions with same SF may not be received by gateway due to collision.
-For this reason, spreading factor assignment of nodes is crucial for network performance.
+Which means, a LoRa GW can simultaneously receive multiple transmissions with different SFs.
+However, simultaneous transmissions with same SF may not be received by GW due to collision.
+For this reason, SF assignment of nodes is crucial for network performance.
 
 \par In a LoRaWAN network, initially, a node is not aware of how far it is from a GW.
-However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission.
+However, a node can guess the distance from a GW by observing received signal power of a downlink transmission.
 If a downlink transmission received signal power is too high, then node can decrease its next transmission SF to decrease power consumption.
 This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper.
-Also, a GW can request from a node to decrease node's SF or transmit power.
+Lowest possible SF assignment scheme is commonly used in LoRaWAN deployments.
+GW can request from a node to decrease its SF or transmit power.
 
@@
 \section{LoRaWAN} \label{LoRaWAN}
-\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN.
-LoRaWAN is a medium access control (MAC) layer protocol which designed for large scale LoRa networks.
+\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
 LoRaWAN is developed and maintained by LoRa Alliance.
 LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
-LoRaWAN  provides inter-operability between different LoRa networks.
-LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmitting, accepting the possibility of the collision.
-
-\par A typical LoRaWAN network consist of three network entities.
+LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision.
+A typical LoRaWAN network consist of following three network entities.
 
 \subsection{End node}
-\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways.
+\par LoRaWAN end node (EN) is a low power embedded device that only communicates to GWs.
 LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
-Different classes provides LPWAN solutions to different deployments and topologies.
+Different classes provides LPWAN solutions to different applications and deployments.
 Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission.
 Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission.
-Receive window is synchronized using a beacon packet transmitted by gateways.
+Receive window is synchronized using a beacon packet transmitted by GWs.
 Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
@@
 \subsection{Gateway}
-\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes.
-A typical gateway can receive from multiple channels at the same time.
-Gateways are typically connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.
+\par LoRaWAN gateway (GW) is a device that communicates to end nodes.
+A typical GW can receive from multiple channels at the same time.
+GWs are typically connected to power grid, so power consumption of a GW is insignificant in most of the deployments.
 
@@
 They implemented a LoRa simulator in ns-3 to study scalability and performance of LoRaWAN networks.
-Their results show that LoRaWAN networks scale well as the number of nodes and gateways increases.
-They also show that SF assignment has great effect on LoRaWAN network performance.
+Their results show that SF assignment has great effect on LoRaWAN network performance.
 
@@
 Authors introduced an error model for determining range as well as interference between multiple simultaneous LoRa transmissions.
-Their simulator supports LoRaWAN Class A end devices, multiple gateways, both upstream and downstream confirmed messages.
+Their simulator supports LoRaWAN Class A end devices, multiple GWs, both upstream and downstream confirmed messages.
 Their results show that allocating network parameters to end devices is hugely important for the performance of LoRaWAN networks.
@@
 \par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions with same and different SFs.
-They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the gateway for modeling interference in multiple gateway scenarios.
+They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the GW for modeling interference in multiple GWs scenarios.
 Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.
@@
 Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}.
-However, how network should assign which SF to which node becomes another issue.
-Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other higher SF transmissions.
+However, how network should assign which SF to which node becomes an important question.
+Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other high SF transmissions.
 In multiple GW scenarios, this approach may increase the collisions with nodes in other GW's range.
@@
 \par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules.
-For this reason, we proposed a machine learning based SF assignment scheme to decrease same SF transmission collisions.
-This technique works by learning the behavior of the nodes in a network.
+For this reason, we proposed a machine learning based SF assignment approach to decrease same SF transmission collisions.
+This technique works by learning the transmission behavior of the nodes in a network.
 A NS can keep track of successful uplink transmissions and their SFs.
@@
 
-\par In this work, decision tree classifier (DTC) and support vector machine (SVM) classifier are evaluated to predict transmission results.
+\par In this work, decision tree classifier (DTC) and support vector machine (SVM) schemes are evaluated to predict transmission results.
 Class weights are balanced according to sample distributions for both methods.
@@
 
-\par We have been able to generate mass amount of LoRaWAN transmission logs for different topologies using the simulator.
+\par We have been able to generate mass amount of LoRaWAN transmission records for different topologies using the simulator.
 Data set size is directly proportional to simulation duration.
 Thus, increasing simulation duration, improves the prediction accuracy.
-In real world deployments, NS can keep track of transmission and it can create a classifier daily basis, then GWs can request from nodes to use their new SFs.
+In real world deployments, NS can keep track of all the transmission and it can create a classifier daily basis, then GWs can request from nodes to use their new SFs.
 
@@
 \par A discrete event simulator is implemented in Python to study effects of different SF strategies in LoRaWANs.
-The LoRaWAN SF simulation tool source code is available at \cite{simlorasf}.
+LoRaWAN SF simulation tool source code is available at \cite{simlorasf}.
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
-Simulator can generate uniformly distributed circular shape network topology with radius (m), number of nodes and number of gateways input parameters.
-Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and spreading factor assignment method.
+Simulator can generate uniformly distributed circular shape network topology with input parameters; radius (m), number of nodes and number of GWs.
+Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and SF assignment method.
 With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
@@
 Transmissions are always initiated by end nodes in pure ALOHA manner.
-Nodes generate a new packet according to Poisson interval for given packet rate input parameter.
+Nodes generate a new packet randomly according to Poisson interval for given packet rate input parameter.
 In the simulator, downlink transmissions are not considered.
@@
 
-\par In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm for European ISM band regulation.
+\par In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm for European ISM band regulation \cite{lorawan.regional.parameters}.
 Different channel transmissions are independent from each other.
@@
 
-\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.
+\par Receive sensitivity of a LoRa GW for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.
 
@@
 
-\par If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission at the duration of transmission.
+\par If the received signal power is higher than the GW sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission.
 
@@
 \section{Simulation Results}  \label{Simulation Results}
-\par For simulation results in this paper, global simulation parameters are set as follows.
-Packet size is 60 bytes including header and payload, simulation duration is 3600 seconds, packet generation rate is 0.01 packet per second.
+\par For simulation results in this paper, global simulation parameters are set as follows; packet size is 60 bytes including header and payload, simulation duration is 3600 seconds, packet generation rate is 0.01 packet per second.
 
@@
 Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1.
-Increasing SFs increases time on air.
-This increases the number of collisions thus decreases the PDR of the network.
-PDR of higher SFs are decreased to almost 0 while number of nodes is increasing.
-Since network topology radius is quite small almost all SFs can reach to the GW.
+Increasing SFs, increases time on air.
+This increases the number of collisions thus decreases the PDR.
+Higher SFs effected exponentially while number of nodes increases.
+Since network topology radius is quite short, all SFs can reach to the GW.
 
@@
 \includegraphics[width=\linewidth]{{gw_pdr_r3000_p0.01_s3600}.png}
-\caption{PDR of different number of gateways.}
+\caption{PDR of different number of GWs.}
 \label{fig:gw_pdr}
@@
 \par For space constraints, we omit figures for network throughput output, total transmit energy consumption output and different packet generation rate input.
-We invite readers to experiment the simulation tool with different inputs and outputs.
+We invite readers to experiment the simulation tool with different input parameters and their results.
 \cite{simlorasf}.
@@
 In Table \ref{table:prediction_pdr}, PDR values for different topology radius are presented.
-Both prediction SF schemes give better PDR than lowest SF schemes when number of nodes increases since number of interference also increases.
-Prediction schemes can only improve network performance when LoRa interference is present.
-Also, prediction schemes give better results when nodes are deployed close to the gateway, since nodes have margin to increase their SFs.
+Both prediction SVM and DTC SF schemes give better PDR than lowest SF schemes when number of nodes increases.
+Increasing number of nodes, increases number of interference.
+Prediction schemes  improve network performance when LoRa interference is high.
+Also, prediction schemes give better results when nodes are deployed close to the GW, since nodes have margin to increase their SFs when deployed close to the GW.
 If a node is far away from the GW, then prediction schemes cannot increase the SF to avoid interference since the SF is already high.
 
-In Table \ref{table:prediction_accuracy}, prediction accuracy of SVM and DTC prediction schemes are presented for different network topology radius and number of nodes. Number of GW is set to 3.
+\par In Table \ref{table:prediction_accuracy}, prediction accuracy of SVM and DTC schemes are presented for different network topology radius and number of nodes.
+Prediction accuracy is not directly proportional to network PDR.
+Correct prediction of a interfered transmission may not increase the PDR but increases the accuracy metric.
+DTC gives better network PDR results than SVM, even overall SVM prediction accuracy is high.
+LoRaWAN simulation usually produces imbalanced dataset, DTC gives better accuracy for interfered transmission labels.
+We invite readers to experiment the simulation tool to review confusion matrices for different machine learning schemes.
 
@@
 \section{Conclusion} \label{Conclusion}
-\par In this paper, after a brief introduction about LPWAN technologies, we provide background information about one of the most promising LPWAN technology, LoRa.
+\par In this paper, after a brief introduction about LPWAN technologies, we provide background information about LoRa and LoRaWAN.
 Then, we discuss about LoRa modulation basics and spreading factor assignment issue.
 We have implemented a discrete event simulator from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
-We have shown how same SF transmission collisions can be avoided and we proposed a machine learning based solution for collision with same SF transmissions.
+We have shown how same SF transmission collisions can be avoided and we proposed machine learning based solution for collision with same SF transmissions.
 We illustrate simulation results for proposed machine learning SF assignment technique.
-Our simulation results show that, proposed technique can increase network performance for LoRaWAN networks when nodes are deployed close to the GWs.
+Our simulation results show that, proposed technique can increase network performance for LoRaWAN networks especially nodes are deployed close to the GWs.
 

### 2019-01-17 paper.tex
grammer

@@
 \par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth.
-This enables tradeoff between receive sensitivity and transmission time on air \cite{7803607}.
+This enables tradeoff between receive sensitivity and time on air of transmission \cite{7803607}.
 Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
 In this work, a LoRa discrete event simulator is developed to study different LoRa SF assignment strategies.
-We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using the simulator.
+Also an SF assignment scheme is proposed to increase network performance with the help of machine learning methods and the proposal is verified using the simulator.
 
@@
 \par Simultaneous different SF transmissions are orthogonal to each other up to some extent.
-Which means, a LoRa GW can simultaneously receive multiple transmissions with different SFs.
-However, simultaneous transmissions with same SF may not be received by GW due to collision.
+Which means that, a LoRa GW can simultaneously receive multiple transmissions with different SFs.
+However, simultaneous transmissions with the same SF may not be received by the GW due to collision.
 For this reason, SF assignment of nodes is crucial for network performance.
 
-\par In a LoRaWAN network, initially, a node is not aware of how far it is from a GW.
+\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a GW.
 However, a node can guess the distance from a GW by observing received signal power of a downlink transmission.
-If a downlink transmission received signal power is too high, then node can decrease its next transmission SF to decrease power consumption.
+If received signal power of a downlink transmission is too high, then node can decrease its next transmission SF to decrease power consumption.
 This SF assignment method is called lowest possible SF assignment scheme for the rest of the paper.
@@
 Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes.
-Most of the end devices which close to the GW will fall into lowest SF (SF7) area section.
-End devices which close to the GW will probably select lowest SF all the time.
-This results a lot of collisions between same SF transmissions.
+Most of the end devices close to the GW will fall into lowest SF (SF7) area section.
+End devices close to the GW will probably select lowest SF all the time.
+This causes a lot of collisions between same SF transmissions.
 Number of collisions increases while number of end devices close to the GW increases.
@@
 \section{LoRaWAN} \label{LoRaWAN}
-\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
+\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN is which designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
 LoRaWAN is developed and maintained by LoRa Alliance.
 LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
-LoRaWAN is based on pure ALOHA medium access which means end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision.
-A typical LoRaWAN network consist of following three network entities.
+LoRaWAN is based on pure ALOHA medium access which means that end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision.
+A typical LoRaWAN network consists of following three network entities.
 
@@
 LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
-Different classes provides LPWAN solutions to different applications and deployments.
-Class A end nodes generate uplink transmission at any time and only receives a period of time after uplink transmission.
+Different classes provide LPWAN solutions to different applications and deployments.
+Class A end nodes generate uplink transmission at any time and only receive a period of time after uplink transmission.
 Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission.
@@
 Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
-This provides Class C end nodes to low latency downlink communication, but it requires more power consumption.
-In this paper, we focused on Class A end devices since Class A behavior provides the lowest power consumption.
+This provides Class C end nodes with low latency downlink communication, but it requires more power consumption.
+In this paper, only Class A end devices are considered since Class A behavior provides the lowest power consumption.
 
@@
 A typical GW can receive from multiple channels at the same time.
-GWs are typically connected to power grid, so power consumption of a GW is insignificant in most of the deployments.
+GWs are usually connected to power grid, so power consumption of a GW is insignificant in most of the deployments.
 
@@
 LPWAN technologies and especially LoRa attracted researchers’ attentions lately.
-We listed some of these works which studies LoRaWAN and LoRa SF.
+Some of these works which studies LoRa/LoRaWAN SF are listed.
 
@@
 Their simulator supports LoRaWAN Class A end devices, multiple GWs, both upstream and downstream confirmed messages.
-Their results show that allocating network parameters to end devices is hugely important for the performance of LoRaWAN networks.
+Their results show that allocating network parameters to end devices is highly important for the performance of LoRaWAN networks.
 
@@
 The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal.
-Their experimental results show that this power difference is around 16 dB and this power difference can be seen when an interferer is close to a receiver or sum of multiple interferer energy can create this power difference.
+Their experimental results show that this power difference is around 16 dB and this power difference can be seen when an interfering signal is close to a receiver or sum of interfering signals' energy can create this power difference.
 
-\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions with same and different SFs.
+\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions with the same and different SFs.
 They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the GW for modeling interference in multiple GWs scenarios.
@@
 \section{Proposed Technique} \label{Proposed Technique}
-\par The collision issue illustrated in Figure \ref{fig:collision} can be solved by forcing some of the close nodes to select higher SF even they can able to communicate with lower SF.
-This may prevent collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}.
+\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SF even they are able to communicate with lower SF.
+This has potential to prevent collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}.
 Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}.
-However, how network should assign which SF to which node becomes an important question.
-Increasing a node's SF should be done carefully since higher SF means higher time on air and higher time on air is more prone to collision with other high SF transmissions.
-In multiple GW scenarios, this approach may increase the collisions with nodes in other GW's range.
-Thus, extra care should be taken for nodes in intersection area of GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
+However, distribution of SF among nodes becomes an important question.
+Increasing a node's SF should be done carefully since higher SF means longer time on air and longer time on air is more prone to collision with other high SF transmissions.
+In multiple GWs scenarios, this approach may increase the collisions with nodes in other GW's range.
+Thus, extra care should be taken for nodes in intersection area of the GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
 
 \par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules.
-For this reason, we proposed a machine learning based SF assignment approach to decrease same SF transmission collisions.
+For this reason, machine learning based SF assignment approach is proposed to decrease same SF transmission collisions.
 This technique works by learning the transmission behavior of the nodes in a network.
@@
 NS can also keep track of some of the collided transmission if header part of the packet is not interfered at location of the GW.
-However, NS cannot keep track of transmissions with receive power lower than the GW sensitivity.
+However, NS cannot keep track of transmissions with receive power lower than sensitivity of the GW.
 Using these obtained information, NS can train a classifier for predicting future transmission result for specific node and specific SF.
@@
 
-\par We have been able to generate mass amount of LoRaWAN transmission records for different topologies using the simulator.
+\par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using the simulator.
 Data set size is directly proportional to simulation duration.
@@
 In the simulator, downlink transmissions are not considered.
-Downlink transmissions are rare in real world deployments since ISM band regulations dictates duty cycle limit for all devices including GWs.
+Downlink transmissions are rare in real world deployments since ISM band regulations dictate duty cycle limit for all devices including GWs.
 
@@
 But, in this work we focus on SF orthogonality.
-Thus, we only utilized single channel transmissions in the simulator.
+Thus, only single channel transmissions are utilized in the simulator.
 
@@
 To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other.
-In simulator, following signal to interference plus noise ratio (SINR) threshold matrix from \cite{goursaud:hal-01231221} is used.
+In simulator, signal to interference plus noise ratio (SINR) threshold matrix from \cite{goursaud:hal-01231221} is used.
 
-\par To decide if a referenced signal is interfered at receiver by a interferer signal, SINR threshold matrix is used.
-$T_{i,j}$ is the needed SINR margin in dB unit between referenced signal with SF = i and interferer signal with SF = j to correctly decode the referenced signal.
-If there are more than one interferer, referenced signal must satisfy the margin for cumulative sum of all interfering signal received powers for each SF \cite{7996384}.
+\par To decide if a referenced signal is interfered at receiver by an interfering signal, SINR threshold matrix is used.
+$T_{i,j}$ is the needed SINR margin in dB unit between referenced signal with SF = i and interfering signal with SF = j to correctly decode the referenced signal.
+If there are more than one interfering signal, referenced signal must satisfy the margin for cumulative sum of all interfering signal received power for each SF \cite{7996384}.
 Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$.
@@
 This increases the number of collisions thus decreases the PDR.
-Higher SFs effected exponentially while number of nodes increases.
+High SF schemes gives poor PDR results while number of nodes increases.
 Since network topology radius is quite short, all SFs can reach to the GW.
@@
 \par In Figure \ref{fig:prediction_pdr}, result of proposed technique is presented.
-PDR plot of lowest SF scheme and machine learning prediction schemes are shown in same figure.
+PDR plot of lowest SF scheme and machine learning prediction schemes are shown in the same figure.
 Randomly generated network topology radius is set to 5000 meters and number of GWs is set to 3.
-Prediction model needs nodes location and three GWs are enough to roughly locate nodes positions by triangulation.
-In Table \ref{table:prediction_pdr}, PDR values for different topology radius are presented.
+Prediction model needs nodes location and three GWs are enough to locate nodes positions by triangulation.
+In Table \ref{table:prediction_pdr}, PDR values for various topologies with different radii are presented.
 Both prediction SVM and DTC SF schemes give better PDR than lowest SF schemes when number of nodes increases.
@@
 Prediction schemes  improve network performance when LoRa interference is high.
-Also, prediction schemes give better results when nodes are deployed close to the GW, since nodes have margin to increase their SFs when deployed close to the GW.
+Also, prediction schemes give better results when nodes are deployed close to the GW, since nodes have margin to increase their SFs when they are deployed close to the GW.
 If a node is far away from the GW, then prediction schemes cannot increase the SF to avoid interference since the SF is already high.
@@
 LoRaWAN simulation usually produces imbalanced dataset, DTC gives better accuracy for interfered transmission labels.
-We invite readers to experiment the simulation tool to review confusion matrices for different machine learning schemes.
+We invite readers to experiment the simulation tool to review confusion matrices for SVM and DTC schemes.
 
@@
 \section{Conclusion} \label{Conclusion}
-\par In this paper, after a brief introduction about LPWAN technologies, we provide background information about LoRa and LoRaWAN.
-Then, we discuss about LoRa modulation basics and spreading factor assignment issue.
-We have implemented a discrete event simulator from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
-We have shown how same SF transmission collisions can be avoided and we proposed machine learning based solution for collision with same SF transmissions.
-We illustrate simulation results for proposed machine learning SF assignment technique.
-Our simulation results show that, proposed technique can increase network performance for LoRaWAN networks especially nodes are deployed close to the GWs.
+\par In this work, after a brief introduction about LPWAN technologies, background information about LoRa and LoRaWAN is provided.
+Then, LoRa modulation basics and spreading factor assignment issue is  discussed.
+A discrete event simulator is implemented from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
+Same SF transmission collisions are shown and machine learning based solution is proposed for collision with same SF transmissions.
+Simulation results for proposed machine learning SF assignment technique is illustrated.
+Simulation results show that, proposed technique can increase network performance for LoRaWAN networks especially nodes are deployed close to the GWs.
 
-\par As for future works, transmit power optimization can be included to proposed machine learning scheme.
-In this paper, we assume that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
-This will make transmissions to more vulnerable to interference thus requires extra care.
+\par As for future works, transmit power optimization can be included in proposed machine learning scheme.
+In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
+This will make transmissions more vulnerable to interference thus requires extra care.
 Also, other machine learning methods can be investigated for SF optimization.

### 2019-01-18 paper.tex
advisor advives

@@
 Collisions in LoRaWAN networks are strongly related with spreading factor (SF) assignment of nodes which indeed effects network performance.
-In this work, a simulation environment to study different SF assignment schemes is implemented.
-Furthermore, a novel SF assignment strategy is proposed which utilizes a machine learning technique for optimization of SF assignment.
-Finally, promising simulation results for proposed machine learning SF assignment technique is presented.
+In this work, a simulation environment to evaluate the performance of SF assignment schemes is implemented.
+Furthermore, a novel smart SF assignment strategy which utilizes Support Vector Machine (SVM) and Decision Tree Classifier (DTC) machine learning techniques for optimization of SF assignment is proposed.
+It is observed and presented that the proposed smart SF assignment techniques give promising simulation results.
 \end{abstract}
@@
 \par There are several emerging LPWAN technologies.
-LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies.
+LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies \cite{7815384}.
 LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}.
@@
 \par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth.
-This enables tradeoff between receive sensitivity and time on air of transmission \cite{7803607}.
-Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance.
-In this work, a LoRa discrete event simulator is developed to study different LoRa SF assignment strategies.
-Also an SF assignment scheme is proposed to increase network performance with the help of machine learning methods and the proposal is verified using the simulator.
+This enables tradeoff between receive sensitivity and air time of transmission \cite{7803607}.
+Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other.
+Thus SF assignment is crucial for overall network performance.
+In this work, a LoRa discrete event simulator is developed to study the performance of LoRa SF assignment strategies.
+Also a machine learning based SF assignment approach is proposed.
+Support Vector Machine (SVM) and Decision Tree Classifier (DTC) techniques are employed and the introduced schemes are called as smart SF schemes.
+The performance of the smart schemes  is compared with the performance of the lowest SF assignment scheme.
+It is shown that the proposed smart SF assignment schemes give promising results.
 
-\par The following two sections provides a background information about LoRa and LoRaWAN.
-Section \ref{Other Related Works} describes other related works.
-Section \ref{Proposed Technique} describes proposed technique for SF assignment.
+\par This paper is organized as follows: Section \ref{LoRa} and Section \ref{LoRaWAN} provides some background information about LoRa and LoRaWAN.
+Section \ref{Related Works} summarize  the other related works.
+Section \ref{Proposed Technique} describes the smart SF assignment technique proposed.
 Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}.
-Finally, Section \ref{Conclusion} concludes the paper.
+Finally, Section \ref{Conclusion} concludes the paper by givin future directions.
 
-\section{LoRa}
-\par LoRa is the name of the proprietary physical layer radio/chipset technology that provides wireless link solution for LPWANs.
-LoRa is designed and patented by Semtech Corporation.
-LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS).
-A chirp is a sinusoidal signal which its frequency increases over time.
-Chirp frequency increases linearly and sweeps the entire bandwidth.
-We invite readers to refer to \cite{AN1200.22} for further details.
+\section{LoRa} \label{LoRa}
+\par LoRa is a proprietary physical layer radio/chipset technology that provides wireless link solution for LPWANs.
+It uses proprietary spread spectrum modulation technique that is the derivative of Chirp Spread Spectrum (CSS).
+A chirp is a sinusoidal signal of which frequency increases over time.
+Chirp frequency increases linearly and sweeps the entire bandwidth \cite{AN1200.22}.
 
 \subsection{Spreading Factor}
-\par Spreading factor (SF) is the ratio between symbol rate and chirp rate.
-The ratio between symbol and chirp rate is equal to $2$\textsuperscript{SF}.
+\par The ratio between symbol and chirp rate is equal to $2$\textsuperscript{SF}.
 SF can take values between 7 to 12.
@@
 Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes.
-Most of the end devices close to the GW will fall into lowest SF (SF7) area section.
-End devices close to the GW will probably select lowest SF all the time.
-This causes a lot of collisions between same SF transmissions.
-Number of collisions increases while number of end devices close to the GW increases.
+The end devices close to the GW will fall into lowest SF (SF7) area section.
+The end devices close to the GW will probably select the lowest SF mot the time.Hence the number of collisions will increase as the number of end devices close to the GW increases.
 
@@
 \section{LoRaWAN} \label{LoRaWAN}
-\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN is which designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
+\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN is which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
 LoRaWAN is developed and maintained by LoRa Alliance.
@@
 Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
-This provides Class C end nodes with low latency downlink communication, but it requires more power consumption.
-In this paper, only Class A end devices are considered since Class A behavior provides the lowest power consumption.
+This provides Class C end nodes with low latency downlink communication, which it requires more power consumption.
+In this paper, only Class A end devices are considered since Class A behavior leads to the lowest power consumption.
 
 \subsection{Gateway}
-\par LoRaWAN gateway (GW) is a device that communicates to end nodes.
+\par LoRaWAN gateway (GW) is a device that communicates with the end nodes.
 A typical GW can receive from multiple channels at the same time.
@@
 
-\section{Other Related Works} \label{Other Related Works}
-\par The literature related to the work presented in this paper has started growing recently.
+\section{Related Works} \label{Related Works}
+\par The literature related to the work presented in this paper has started to grow recently.
 LPWAN technologies and especially LoRa attracted researchers’ attentions lately.
-Some of these works which studies LoRa/LoRaWAN SF are listed.
+Some of these works which studies LoRa/LoRaWAN SF are summarized.
 
@@
 Their simulator supports LoRaWAN Class A end devices, multiple GWs, both upstream and downstream confirmed messages.
-Their results show that allocating network parameters to end devices is highly important for the performance of LoRaWAN networks.
+Their results show that allocating network parameters such as SF is highly important for the performance of LoRaWAN networks.
 
-\par In \cite{8267219}, the authors studied imperfect orthogonality between different LoRa SF transmissions.
-The authors state that a LoRa transmission can be interfered even between different SF transmissions when power of the interfering signal significantly overcomes the reference signal.
-Their experimental results show that this power difference is around 16 dB and this power difference can be seen when an interfering signal is close to a receiver or sum of interfering signals' energy can create this power difference.
+\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
+The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal significantly is greater than the reference signal.
+Their experimental results show that this power difference is around 16 dB such a this power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.
 
-\par In \cite{8430542}, the authors investigated the impact of interference caused by simultaneous LoRa transmissions with the same and different SFs.
+\par In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs.
 They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the GW for modeling interference in multiple GWs scenarios.
@@
 \section{Proposed Technique} \label{Proposed Technique}
-\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SF even they are able to communicate with lower SF.
-This has potential to prevent collisions due to the orthogonality of different SF as shown in Figure \ref{fig:collision_solution_single_gw}.
+\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even they are able to communicate with lower SFs.
+This has potential to prevent collisions due to the orthogonality of the SFs as shown in Figure \ref{fig:collision_solution_single_gw}.
 Higher SF assigned nodes are drawn with bold circle border in Figure \ref{fig:collision_solution_single_gw}.
-However, distribution of SF among nodes becomes an important question.
-Increasing a node's SF should be done carefully since higher SF means longer time on air and longer time on air is more prone to collision with other high SF transmissions.
-In multiple GWs scenarios, this approach may increase the collisions with nodes in other GW's range.
-Thus, extra care should be taken for nodes in intersection area of the GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
+However, distribution of SF among nodes becomes an important problem.
+Increasing a node's SF should be done carefully since higher SF means longer air time and longer air time means increasing the probability of collisions with other high SF transmissions.
+In multiple GWs scenarios, this approach may increase the collisions with the nodes in other GW's range.
+Thus, extra care should be taken for nodes in the intersection area of the GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
 
 \par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules.
-For this reason, machine learning based SF assignment approach is proposed to decrease same SF transmission collisions.
-This technique works by learning the transmission behavior of the nodes in a network.
-A NS can keep track of successful uplink transmissions and their SFs.
-NS can also keep track of some of the collided transmission if header part of the packet is not interfered at location of the GW.
-However, NS cannot keep track of transmissions with receive power lower than sensitivity of the GW.
-Using these obtained information, NS can train a classifier for predicting future transmission result for specific node and specific SF.
-Using this prediction model NS can assign the SF to nodes considering the collision probability.
+For this reason, machine learning based SF assignment approach is proposed to decrease the colisioins for the same SF transmissions.
+This technique starts by learning the transmission behavior of the nodes in a network.
+An NS can keep track of successful uplink transmissions and their SFs.
+This NS can also keep track of some of the collided transmissions if header part of the packet is not interfered at the GW.
+However, NS cannot keep track of transmissions with lower receive power than sensitivity of the GW.
+Using this information, NS can train a classifier to predict future transmission result for a specific node and a specific SF.
+Using this prediction model NS can assign SFs to nodes considering the collision probability.
 
-\par In this work, decision tree classifier (DTC) and support vector machine (SVM) schemes are evaluated to predict transmission results.
+\par In this work, decision tree classifier (DTC) and support vector machine (SVM) \cite{Alpaydin} schemes are employed to predict the transmission results.
 Class weights are balanced according to sample distributions for both methods.
-For DTC, Gini impurity criteria is used to measure the quality of a splits.
+For DTC, Gini impurity criteria is used to measure the quality of splits.
 For SVM, penalty parameter is set to 1, degree is set to 3 and RBF kernel is used.
 
-\par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using the simulator.
-Data set size is directly proportional to simulation duration.
-Thus, increasing simulation duration, improves the prediction accuracy.
-In real world deployments, NS can keep track of all the transmission and it can create a classifier daily basis, then GWs can request from nodes to use their new SFs.
+\par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using our simulator.
+Definetely, data set size is directly proportional to simulation duration.
+Thus, increasing the simulation duration, improves the prediction accuracy.
+In real world deployments, NS can keep track of all transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.
 
-\par To come up with a solution for SF assignment issue, DTC and SCM prediction schemes are integrated to the simulator.
+\par To come up with a solution for SF assignment, DTC and SVM prediction schemes are integrated to the simulator.
 A classifier is trained from transmission logs and this classifier is used for selecting optimum SF for the nodes.
-Simulator first runs a simulation with assigning random SF to the nodes.
-Simulation results are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and result of the transmission.
+Simulator first runs a simulation by assigning random SFs to the nodes.
+Simulation logs are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and the result of the transmission.
 Possible transmission results are successful, interfered or under sensitivity.
@@
 After the classifier model is generated, a second simulation is run with classifier.
-Simulator selects optimum SF considering prediction of transmission.
-If a transmission of a node with lowest possible SF is predicted as interfered by the classifier, then simulator assigns higher SF and calculates new prediction.
+Simulator selects optimum SF considering the prediction of transmission.
+If a transmission of a node with the lowest possible SF is predicted as interfered by the classifier, then simulator assigns higher SF and evaluates the new prediction.
 If no SF is predicted as successful, then simulator again assigns the lowest possible SF.
@@
 \section{Simulation Environment} \label{Simulation Environment}
-\par A discrete event simulator is implemented in Python to study effects of different SF strategies in LoRaWANs.
+\par A discrete event simulator is implemented in Python to study the effects of different SF strategies in LoRaWANs.
 LoRaWAN SF simulation tool source code is available at \cite{simlorasf}.
@@
 
-\par In the simulator, only LoRaWAN Class A devices are covered.
+\par The simulator only covers the LoRaWAN Class A devices.
 Transmissions are always initiated by end nodes in pure ALOHA manner.
-Nodes generate a new packet randomly according to Poisson interval for given packet rate input parameter.
-In the simulator, downlink transmissions are not considered.
+Nodes generate a new packet randomly according to a Poisson interval for given packet rate parameter.
+Downlink transmissions are not considered.
 Downlink transmissions are rare in real world deployments since ISM band regulations dictate duty cycle limit for all devices including GWs.
 
-\subsection{Link Model}
+\subsection{Link Model Employed}
 \par Link quality of a wireless system can be expressed by the metric of link budget.
@@
 
-\subsection{Interference Model}
+\subsection{Interference Model Employed}
 \par In the simulator, interference model described in \cite{7996384} is adopted.
@@
 
-\par In our simulator, it is assumed that there is no other technology interference in the network except LoRa interference.
+\par It is assumed that there is no other technology interference in the network except LoRa interference.
 To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other.
@@
 \par To decide if a referenced signal is interfered at receiver by an interfering signal, SINR threshold matrix is used.
-$T_{i,j}$ is the needed SINR margin in dB unit between referenced signal with SF = i and interfering signal with SF = j to correctly decode the referenced signal.
+$T_{i,j}$ is SINR margin in dB unit between referenced signal with SF = i and interfering signal with SF = j to correctly decode the referenced signal.
 If there are more than one interfering signal, referenced signal must satisfy the margin for cumulative sum of all interfering signal received power for each SF \cite{7996384}.
@@
 \section{Simulation Results}  \label{Simulation Results}
-\par For simulation results in this paper, global simulation parameters are set as follows; packet size is 60 bytes including header and payload, simulation duration is 3600 seconds, packet generation rate is 0.01 packet per second.
+\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 sec, packet generation rate = 0.01 pac/sec.
 
-\par In Figure \ref{fig:sf_pdr}, PDR plot of different SF assignment is shown.
+\par In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown.
 Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1.
-Increasing SFs, increases time on air.
-This increases the number of collisions thus decreases the PDR.
-High SF schemes gives poor PDR results while number of nodes increases.
-Since network topology radius is quite short, all SFs can reach to the GW.
+Increasing SFs, increases air time.
+It increases the number of collisions thus decreases the PDR.
+High SF schemes gives poor PDR results as the number of nodes increases.
+Since network topology radius is quite small, all SFs can reach to GW.
 
@@
 \includegraphics[width=\linewidth]{{sf_pdr_r3000_g1_p0.01_s3600}.png}
-\caption{PDR of different SFs.}
+\caption{PDR for various SFs.}
 \label{fig:sf_pdr}
@@
 
-\par In Figure \ref{fig:gw_pdr}, PDR plot of different number of GWs is shown.
-Randomly generated topology radius is set to 3000 meters and lowest possible SF assignment scheme is used.
-Increasing number of GWs, decreases the SFs of nodes and decreases time on air.
-This decreases number of collisions thus increases the PDR of the network when network topology radius is constant.
+\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of GWs is shown.
+Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used.
+Increasing number of GWs, decreases the SFs of nodes and decreases air time.
+This decreases number of collisions, henc, increases the PDR of network when network radius is constant.
 
@@
 \includegraphics[width=\linewidth]{{gw_pdr_r3000_p0.01_s3600}.png}
-\caption{PDR of different number of GWs.}
+\caption{PDR for various number of GWs.}
 \label{fig:gw_pdr}
@@
 
-\par In Figure \ref{fig:r_pdr}, PDR plot of different network topology radius is shown.
-Number of GW is set to 1 and lowest possible SF assignment scheme is used.
-Increasing network topology radius, increases the number of under sensitivity transmissions thus decreases the PDR of the network.
+\par In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown.
+Number of GW is set to 1 and lowest SF assignment scheme is used.
+Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.
 
@@
 \includegraphics[width=\linewidth]{{r_pdr_g1_p0.01_s3600}.png}
-\caption{PDR of different topology radius.}
+\caption{PDR for various network radii.}
 \label{fig:r_pdr}
@@
 
-\par For space constraints, we omit figures for network throughput output, total transmit energy consumption output and different packet generation rate input.
-We invite readers to experiment the simulation tool with different input parameters and their results.
-\cite{simlorasf}.
+\par PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}.
+Randomly generated network radius is set to 5000 meters and number of GWs is set to 3.
+Prediction model needs nodes' locations and three GWs are enough to locate nodes positions by triangulation.
+In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented.
+Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases.
+Increasing number of nodes, increases number of interference.
+Smart schemes  improve network performance when LoRa interference is high.
+Moreover, smart schemes give better results when nodes are deployed closer to the GW, since nodes have margin to increase their SFs when they are deployed closer to the GW.
+If a node is far away from GW, then smart schemes cannot increase the SF to avoid interference since the asigned SF is already high.
 
-\par In Figure \ref{fig:prediction_pdr}, result of proposed technique is presented.
-PDR plot of lowest SF scheme and machine learning prediction schemes are shown in the same figure.
-Randomly generated network topology radius is set to 5000 meters and number of GWs is set to 3.
-Prediction model needs nodes location and three GWs are enough to locate nodes positions by triangulation.
-In Table \ref{table:prediction_pdr}, PDR values for various topologies with different radii are presented.
-Both prediction SVM and DTC SF schemes give better PDR than lowest SF schemes when number of nodes increases.
-Increasing number of nodes, increases number of interference.
-Prediction schemes  improve network performance when LoRa interference is high.
-Also, prediction schemes give better results when nodes are deployed close to the GW, since nodes have margin to increase their SFs when they are deployed close to the GW.
-If a node is far away from the GW, then prediction schemes cannot increase the SF to avoid interference since the SF is already high.
-
-\par In Table \ref{table:prediction_accuracy}, prediction accuracy of SVM and DTC schemes are presented for different network topology radius and number of nodes.
+\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes.
 Prediction accuracy is not directly proportional to network PDR.
-Correct prediction of a interfered transmission may not increase the PDR but increases the accuracy metric.
-DTC gives better network PDR results than SVM, even overall SVM prediction accuracy is high.
-LoRaWAN simulation usually produces imbalanced dataset, DTC gives better accuracy for interfered transmission labels.
-We invite readers to experiment the simulation tool to review confusion matrices for SVM and DTC schemes.
+Correct prediction of a interfered transmission may not increase the PDR but increases the accuracy.
+Mart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.
 
@@
 \includegraphics[width=\linewidth]{{prediction_pdr_r5000_g3_p0.01_s3600}.png}
-\caption{PDR of SF prediction scheme.}
+\caption{PDR for lowest SF and smart SF schemes.}
 \label{fig:prediction_pdr}
@@
 \centering
-\caption{PDR of lowest possible SF and smart SF assignment schemes.}
+\caption{PDR for lowest SF and smart SF schemes.}
 \label{table:prediction_pdr}
@@
 
+\par For space constraints, we omit results like network throughput, total transmit energy consumption.
+We invite readers to experiment the simulation tool with the parameters they like.
+\cite{simlorasf}.
+
 \section{Conclusion} \label{Conclusion}
-\par In this work, after a brief introduction about LPWAN technologies, background information about LoRa and LoRaWAN is provided.
-Then, LoRa modulation basics and spreading factor assignment issue is  discussed.
-A discrete event simulator is implemented from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
-Same SF transmission collisions are shown and machine learning based solution is proposed for collision with same SF transmissions.
-Simulation results for proposed machine learning SF assignment technique is illustrated.
-Simulation results show that, proposed technique can increase network performance for LoRaWAN networks especially nodes are deployed close to the GWs.
+\par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
+Moreover, we show how SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme.
+We present simulation results for the lowest SF assignment scheme and the proposed schemes.
+Simulation results show that, the the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.
 
-\par As for future works, transmit power optimization can be included in proposed machine learning scheme.
-In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to the GW can decrease transmit power to save energy.
+\par As for future work, transmit power optimization can be included to the proposed smart schemes.
+In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to GWs can decrease transmit power to save energy.
 This will make transmissions more vulnerable to interference thus requires extra care.
-Also, other machine learning methods can be investigated for SF optimization.
-Reinforcement learning is a good candidate for this network optimization issue.
-Other transmission parameters such as node id and transmission time can be included to proposed scheme to improve performance.
+Moreover, other machine learning methods can be investigated for SF assignment enhancement.
+Reinforcement learning could be a good candidate.
+Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve performance.
 

### 2019-01-18 paper.tex
spell check

@@
 Simultaneous same SF transmissions are prone to collision, however, different SF transmissions in the same channel are orthogonal to each other.
-Thus SF assignment is crucial for overall network performance.
+Thus, SF assignment is crucial for overall network performance.
 In this work, a LoRa discrete event simulator is developed to study the performance of LoRa SF assignment strategies.
-Also a machine learning based SF assignment approach is proposed.
+Also, a machine learning based SF assignment approach is proposed.
 Support Vector Machine (SVM) and Decision Tree Classifier (DTC) techniques are employed and the introduced schemes are called as smart SF schemes.
@@
 Simulation environments and results are shown in Section \ref{Simulation Environment} and \ref{Simulation Results}.
-Finally, Section \ref{Conclusion} concludes the paper by givin future directions.
+Finally, Section \ref{Conclusion} concludes the paper by giving future directions.
 
@@
 The end devices close to the GW will fall into lowest SF (SF7) area section.
-The end devices close to the GW will probably select the lowest SF mot the time.Hence the number of collisions will increase as the number of end devices close to the GW increases.
+The end devices close to the GW will probably select the lowest SF mot the time.
+Hence the number of collisions will increase as the number of end devices close to the GW increases.
 
@@
 The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal significantly is greater than the reference signal.
-Their experimental results show that this power difference is around 16 dB such a this power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.
+Their experimental results show that this power difference is around 16 dB such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.
 
@@
 \par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules.
-For this reason, machine learning based SF assignment approach is proposed to decrease the colisioins for the same SF transmissions.
+For this reason, machine learning based SF assignment approach is proposed to decrease the collisions for the same SF transmissions.
 This technique starts by learning the transmission behavior of the nodes in a network.
@@
 \par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using our simulator.
-Definetely, data set size is directly proportional to simulation duration.
+Data set size is directly proportional to simulation duration.
 Thus, increasing the simulation duration, improves the prediction accuracy.
@@
 Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1.
-Increasing SFs, increases air time.
+Increasing SFs increases air time.
 It increases the number of collisions thus decreases the PDR.
@@
 Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases.
-Increasing number of nodes, increases number of interference.
+Increasing number of nodes, increases number of interferences.
 Smart schemes improve network performance when LoRa interference is high.
 Moreover, smart schemes give better results when nodes are deployed closer to the GW, since nodes have margin to increase their SFs when they are deployed closer to the GW.
-If a node is far away from GW, then smart schemes cannot increase the SF to avoid interference since the asigned SF is already high.
+If a node is far away from GW, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.
 
@@
 Prediction accuracy is not directly proportional to network PDR.
-Correct prediction of a interfered transmission may not increase the PDR but increases the accuracy.
+Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy.
 Mart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.
@@
 
-\par For space constraints, we omit results like network throughput, total transmit energy consumption.
+\par For space constraints, we omit results like network throughput and total transmit energy consumption.
 We invite readers to experiment the simulation tool with the parameters they like.
@@
 We present simulation results for the lowest SF assignment scheme and the proposed schemes.
-Simulation results show that, the the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.
+Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.
 

### 2019-01-23 paper.tex
hoca name
grammer

@@
 
-\author{\IEEEauthorblockN{Tugrul Yatagan and Sema F. Oktug}
+\author{\IEEEauthorblockN{Tugrul Yatagan and Sema Oktug}
 \IEEEauthorblockA{Department of Computer Engineering,
@@
 Lowest possible SF assignment scheme is commonly used in LoRaWAN deployments.
-GW can request from a node to decrease its SF or transmit power.
+Also, a GW can request from a node to decrease its SF or transmit power.
 
@@
 The end devices close to the GW will fall into lowest SF (SF7) area section.
-The end devices close to the GW will probably select the lowest SF mot the time.
+The end devices close to the GW will probably select the lowest SF most of the time.
 Hence the number of collisions will increase as the number of end devices close to the GW increases.
@@
 \section{LoRaWAN} \label{LoRaWAN}
-\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN is which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
+\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
 LoRaWAN is developed and maintained by LoRa Alliance.
@@
 \par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
-The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal significantly is greater than the reference signal.
-Their experimental results show that this power difference is around 16 dB such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.
+The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal.
+Their experimental results show that this power difference is around 16 dB.
+Such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.
 
@@
 
-\par It is possible to generate mass amount of LoRaWAN transmission records for different topologies using our simulator.
+\par It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator.
 Data set size is directly proportional to simulation duration.
 Thus, increasing the simulation duration, improves the prediction accuracy.
-In real world deployments, NS can keep track of all transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.
+In real world deployments, NS can keep track of  transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.
 
@@
 Increasing number of GWs, decreases the SFs of nodes and decreases air time.
-This decreases number of collisions, henc, increases the PDR of network when network radius is constant.
+This decreases number of collisions, hence, increases the PDR of network when network radius is constant.
 
@@
 Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy.
-Mart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.
+Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.
 
@@
 \centering
-\caption{Prediction accuracy of SVM and DTC classifiers.}
+\caption{Prediction accuracy for SVM and DTC.}
 \label{table:prediction_accuracy}
@@
 \par For space constraints, we omit results like network throughput and total transmit energy consumption.
-We invite readers to experiment the simulation tool with the parameters they like.
-\cite{simlorasf}.
+We invite readers to experiment the simulation tool with the parameters they like \cite{simlorasf}.
 
@@
 \par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
-Moreover, we show how SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme.
+Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme.
 We present simulation results for the lowest SF assignment scheme and the proposed schemes.
@@
 Reinforcement learning could be a good candidate.
-Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve performance.
+Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve prediction performance.
 

### 2019-01-23 paper.tex
add new related work

@@
 
-\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
+\par In \cite{8267219} and \cite{8430542}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
 The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal.
-Their experimental results show that this power difference is around 16 dB.
-Such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.
+Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.
 
-\par In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs.
-They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the GW for modeling interference in multiple GWs scenarios.
-Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.
+\par In \cite{8115779}, the authors introduced two new algorithms for SF assignment in LoRaWANs.
+First algorithm assigns SFs based on the total number of connected devices, second algorithm assigns SFs by balancing air time of the packets in each SF.
+Their simulation results show that their proposed algorithms give promising results.
 

### 2019-02-25 paper.tex
related hoca
radius
reviewer grammer fix

@@
 \par Simultaneous different SF transmissions are orthogonal to each other up to some extent.
-Which means that, a LoRa GW can simultaneously receive multiple transmissions with different SFs.
+Which means that, a LoRa gateway (GW) can simultaneously receive multiple transmissions with different SFs.
 However, simultaneous transmissions with the same SF may not be received by the GW due to collision.
@@
 Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
-This provides Class C end nodes with low latency downlink communication, which it requires more power consumption.
+This provides Class C end nodes with low latency downlink communication, which requires more power consumption.
 In this paper, only Class A end devices are considered since Class A behavior leads to the lowest power consumption.
@@
 First algorithm assigns SFs based on the total number of connected devices, second algorithm assigns SFs by balancing air time of the packets in each SF.
-Their simulation results show that their proposed algorithms give promising results.
+Based on the simulation results, it is shown that the proposed algorithms give promising results.
 
 \section{Proposed Technique} \label{Proposed Technique}
-\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even they are able to communicate with lower SFs.
+\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even though they are able to communicate with lower SFs.
 This has potential to prevent collisions due to the orthogonality of the SFs as shown in Figure \ref{fig:collision_solution_single_gw}.
@@
 \multicolumn{3}{|c|}{}                                             & 100           & 500           & 1000          \\ \hline
-\multirow{12}{*}{\textbf{R (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
+\multirow{12}{*}{\textbf{r (m)}} & \multirow{3}{*}{3000}  & Lowest & 97.8          & 86.0          & 72.3          \\ \cline{3-6}
                                  &                        & SVM    & 98.0          & 88.2          & 75.2          \\ \cline{3-6}
@@
 \multicolumn{3}{|c|}{}                                         & 100           & 500           & 1000          \\ \hline
-\multirow{8}{*}{\textbf{R (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
+\multirow{8}{*}{\textbf{r (m)}} & \multirow{2}{*}{3000}  & SVM & 82.4          & 70.4          & 71.7          \\ \cline{3-6}
                                 &                        & DTC & 86.0          & 67.3          & 70.4          \\ \cline{2-6}

### 2019-02-26 paper.tex
extend abbreviations

@@
 \section{LoRa} \label{LoRa}
-\par LoRa is a proprietary physical layer radio/chipset technology that provides wireless link solution for LPWANs.
+\par LoRa is a proprietary physical layer radio/chipset technology that provides wireless link solution for low power wide area networks.
 It uses proprietary spread spectrum modulation technique that is the derivative of Chirp Spread Spectrum (CSS).
@@
 
-Where, $R_{b}$ is data rate in bps, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.
+Where, $R_{b}$ is data rate in bit/sec, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.
 
@@
 \par Simultaneous different SF transmissions are orthogonal to each other up to some extent.
-Which means that, a LoRa gateway (GW) can simultaneously receive multiple transmissions with different SFs.
-However, simultaneous transmissions with the same SF may not be received by the GW due to collision.
+Which means that, a LoRa gateway can simultaneously receive multiple transmissions with different SFs.
+However, simultaneous transmissions with the same SF may not be received by the gateway due to collision.
 For this reason, SF assignment of nodes is crucial for network performance.
 
-\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a GW.
-However, a node can guess the distance from a GW by observing received signal power of a downlink transmission.
+\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a gateway.
+However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission.
 If received signal power of a downlink transmission is too high, then node can decrease its next transmission SF to decrease power consumption.
@@
 Lowest possible SF assignment scheme is commonly used in LoRaWAN deployments.
-Also, a GW can request from a node to decrease its SF or transmit power.
+Also, a gateway can request from a node to decrease its SF or transmit power.
 
-\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single GW is illustrated.
-Different color rings represent achievable range of different SFs from the GW and different color circles represent selected SF of the nodes.
-The end devices close to the GW will fall into lowest SF (SF7) area section.
-The end devices close to the GW will probably select the lowest SF most of the time.
-Hence the number of collisions will increase as the number of end devices close to the GW increases.
+\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single gateway is illustrated.
+Different color rings represent achievable range of different SFs from the gateway and different color circles represent selected SF of the nodes.
+The end devices close to the gateway will fall into lowest SF (SF7) area section.
+The end devices close to the gateway will probably select the lowest SF most of the time.
+This causes a lot of collisions between same SF transmissions.
+Hence the number of collisions will increase as the number of end devices close to the gateway increases.
 
@@
 \includegraphics[width=\linewidth]{collision}
-\caption{Collision between nodes close to the GW.}
+\caption{Collision between nodes close to the gateway.}
 \label{fig:collision}
@@
 LoRa Alliance is an open, non-profit organization dedicated to standardization of LoRaWAN.
+LoRaWAN provides inter-operability between different LoRa networks.
+LoRa can be used as a wireless link technology without complying LoRaWAN, however this would break inter-operability between different LoRa networks.
 LoRaWAN is based on pure ALOHA medium access which means that end nodes do not check whether the channel is free or not before transmission, accepting the possibility of a collision.
@@
 \subsection{End node}
-\par LoRaWAN end node (EN) is a low power embedded device that only communicates to GWs.
+\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways.
 LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
@@
 Class B end nodes extend Class A behavior by adding scheduled receive windows for downlink transmission.
-Receive window is synchronized using a beacon packet transmitted by GWs.
+Receive window is synchronized using a beacon packet transmitted by gateways.
 Class C end nodes extend Class A behavior by keeping receive window open all the time except uplink transmission.
@@
 \subsection{Gateway}
-\par LoRaWAN gateway (GW) is a device that communicates with the end nodes.
-A typical GW can receive from multiple channels at the same time.
-GWs are usually connected to power grid, so power consumption of a GW is insignificant in most of the deployments.
+\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes.
+A typical gateway can receive from multiple channels at the same time.
+gateways are usually connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.
 
@@
 Authors introduced an error model for determining range as well as interference between multiple simultaneous LoRa transmissions.
-Their simulator supports LoRaWAN Class A end devices, multiple GWs, both upstream and downstream confirmed messages.
+Their simulator supports LoRaWAN Class A end devices, multiple gateways, both upstream and downstream confirmed messages.
 Their results show that allocating network parameters such as SF is highly important for the performance of LoRaWAN networks.
 
-\par In \cite{8267219} and \cite{8430542}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
+\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
 The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal.
+Their experimental results show that this power difference is around 16 dB.
+Such a power difference can be seen when an interfering signal is close to a receiver or the sum of interfering signals' energy can create this power difference.
+
+\par In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs.
+They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the gateway for modeling interference in multiple gateways scenarios.
 Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.
@@
 Increasing a node's SF should be done carefully since higher SF means longer air time and longer air time means increasing the probability of collisions with other high SF transmissions.
-In multiple GWs scenarios, this approach may increase the collisions with the nodes in other GW's range.
-Thus, extra care should be taken for nodes in the intersection area of the GWs illustrated in Figure \ref{fig:collision_solution_multi_gw}.
+In multiple gateways scenarios, this approach may increase the collisions with the nodes in other gateway's range.
+Thus, extra care should be taken for nodes in the intersection area of the gateways illustrated in Figure \ref{fig:collision_solution_multi_gw}.
 
@@
 An NS can keep track of successful uplink transmissions and their SFs.
-This NS can also keep track of some of the collided transmissions if header part of the packet is not interfered at the GW.
-However, NS cannot keep track of transmissions with lower receive power than sensitivity of the GW.
+This NS can also keep track of some of the collided transmissions if header part of the packet is not interfered at the gateway.
+However, NS cannot keep track of transmissions with lower receive power than sensitivity of the gateway.
 Using this information, NS can train a classifier to predict future transmission result for a specific node and a specific SF.
@@
 Thus, increasing the simulation duration, improves the prediction accuracy.
-In real world deployments, NS can keep track of  transmissions and it can create a classifier daily basis, then GWs can request from nodes to use suggested SFs.
+In real world deployments, NS can keep track of  transmissions and it can create a classifier daily basis, then gateways can request from nodes to use suggested SFs.
 
@@
 Simulator first runs a simulation by assigning random SFs to the nodes.
-Simulation logs are combined into four data columns which are; X and Y coordinates of the transmission source, SF of the transmission and the result of the transmission.
+Simulation logs are combined into four data columns which are: X and Y coordinates of the transmission source, SF of the transmission and the result of the transmission.
 Possible transmission results are successful, interfered or under sensitivity.
@@
 \includegraphics[width=\linewidth]{collision_solution_single_gw}
-\caption{Collision avoidance by using higher SF for nodes close to the GW.}
+\caption{Collision avoidance by using higher SF for nodes close to the gateway.}
 \label{fig:collision_solution_single_gw}
@@
 \includegraphics[width=\linewidth]{collision_solution_multi_gw}
-\caption{Collision avoidance for intersecting GWs.}
+\caption{Collision avoidance for intersecting gateways.}
 \label{fig:collision_solution_multi_gw}
@@
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
-Simulator can generate uniformly distributed circular shape network topology with input parameters; radius (m), number of nodes and number of GWs.
-Global simulation input parameters are simulation duration (s), packet size (B), packet generation rate (p/s) and SF assignment method.
-With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput (bps) and total transmit energy consumption (J).
+Simulator can generate uniformly distributed circular shape network topology with following input parameters: radius in meter (m), number of nodes and number of gateways.
+Global simulation input parameters are simulation duration in second (s), packet size in Byte (B), packet generation rate in packets per second (pac/sec) and SF assignment method.
+With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput in bits per second (bit/sec) and total transmit energy consumption in Joule (J).
 Simulator also produces prediction accuracy and confusion matrix for machine learning schemes.
@@
 Downlink transmissions are not considered.
-Downlink transmissions are rare in real world deployments since ISM band regulations dictate duty cycle limit for all devices including GWs.
+Downlink transmissions are rare in real world deployments since ISM band regulations dictate duty cycle transmission limit for all devices including gateways.
 
@@
 
-\par Receive sensitivity of a LoRa GW for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.
+\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.
 
@@
 
-\par If the received signal power is higher than the GW sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission.
+\par If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission.
 
@@
 \section{Simulation Results} \label{Simulation Results}
-\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 sec, packet generation rate = 0.01 pac/sec.
+\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds, packet generation rate = 0.01 pac/sec.
 
 \par In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown.
-Randomly generated network topology radius is set to 3000 meters and number of GWs is set to 1.
+Randomly generated network topology radius is set to 3000 meters and number of gateways is set to 1.
 Increasing SFs increases air time.
@@
 High SF schemes gives poor PDR results as the number of nodes increases.
-Since network topology radius is quite small, all SFs can reach to GW.
+Since network topology radius is quite small, all SFs can reach to gateway.
 
@@
 \includegraphics[width=\linewidth]{{sf_pdr_r3000_g1_p0.01_s3600}.png}
-\caption{PDR for various SFs.}
+\caption{PDR for various SFs. (r = 3000 m, GW = 1)}
 \label{fig:sf_pdr}
@@
 
-\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of GWs is shown.
+\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of gateways is shown.
 Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used.
-Increasing number of GWs, decreases the SFs of nodes and decreases air time.
+Increasing number of gateways, decreases the SFs of nodes and decreases air time.
 This decreases number of collisions, hence, increases the PDR of network when network radius is constant.
@@
 \includegraphics[width=\linewidth]{{gw_pdr_r3000_p0.01_s3600}.png}
-\caption{PDR for various number of GWs.}
+\caption{PDR for various number of GWs. (r = 3000 m, SF = SF{\_}Lowest)}
 \label{fig:gw_pdr}
@@
 \par In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown.
-Number of GW is set to 1 and lowest SF assignment scheme is used.
+Number of gateway is set to 1 and lowest SF assignment scheme is used.
 Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.
@@
 \includegraphics[width=\linewidth]{{r_pdr_g1_p0.01_s3600}.png}
-\caption{PDR for various network radii.}
+\caption{PDR for various network radii. (GW = 1, SF = SF{\_}Lowest)}
 \label{fig:r_pdr}
@@
 \par PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}.
-Randomly generated network radius is set to 5000 meters and number of GWs is set to 3.
-Prediction model needs nodes' locations and three GWs are enough to locate nodes positions by triangulation.
+Randomly generated network radius is set to 5000 meters and number of gateways is set to 3.
+Prediction model needs nodes' locations and three gateways are enough to locate position of nodes by triangulation.
 In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented.
@@
 Smart schemes improve network performance when LoRa interference is high.
-Moreover, smart schemes give better results when nodes are deployed closer to the GW, since nodes have margin to increase their SFs when they are deployed closer to the GW.
-If a node is far away from GW, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.
+Moreover, smart schemes give better results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway.
+If a node is far away from gateway, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.
 
@@
 Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.
+LoRaWAN simulation usually produces imbalanced dataset, DTC gives better accuracy for interfered transmission labels.
 
@@
 \includegraphics[width=\linewidth]{{prediction_pdr_r5000_g3_p0.01_s3600}.png}
-\caption{PDR for lowest SF and smart SF schemes.}
+\caption{PDR for lowest and smart SF schemes. (r = 5000 m, GW = 3)}
 \label{fig:prediction_pdr}
@@
 \centering
-\caption{PDR for lowest SF and smart SF schemes.}
+\caption{PDR for lowest and smart SF schemes. (GW = 3)}
 \label{table:prediction_pdr}
@@
 \centering
-\caption{Prediction accuracy for SVM and DTC.}
+\caption{Prediction accuracy for SVM and DTC. (GW = 3)}
 \label{table:prediction_accuracy}
@@
 
-\par For space constraints, we omit results like network throughput and total transmit energy consumption.
+\par For space constraints, we omit simulation results like network throughput, total transmit energy consumption and confusion matrices.
 We invite readers to experiment the simulation tool with the parameters they like \cite{simlorasf}.
@@
 \section{Conclusion} \label{Conclusion}
-\par In this paper, we present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
+\par In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed.
+We present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
 Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme.
 We present simulation results for the lowest SF assignment scheme and the proposed schemes.
-Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to GWs.
+Simulation results show that, the proposed smart schemes can increase network performance for LoRaWAN networks, especially, when the nodes are deployed close to gateways.
 
 \par As for future work, transmit power optimization can be included to the proposed smart schemes.
-In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to GWs can decrease transmit power to save energy.
+In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to gateways can decrease transmit power to save energy.
 This will make transmissions more vulnerable to interference thus requires extra care.

### 2019-02-26 paper.tex
revert omitted lora details

@@
 
-Where, $R_{b}$ is data rate in bit/sec, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.
+Where, $R_{b}$ is data rate in bps, SF is spreading factor $SF \in \{7,..,12\}$, CR is error correction code rate $CR \in \{1,..,4\}$ and $BW$ is bandwidth in Hertz \cite{AN1200.22}.
 
@@
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
-Simulator can generate uniformly distributed circular shape network topology with following input parameters: radius in meter (m), number of nodes and number of gateways.
-Global simulation input parameters are simulation duration in second (s), packet size in Byte (B), packet generation rate in packets per second (pac/sec) and SF assignment method.
-With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput in bits per second (bit/sec) and total transmit energy consumption in Joule (J).
+Simulator can generate uniformly distributed circular shape network topology with following input parameters: radius (r) in meter (m), number of nodes and number of gateways.
+Global simulation input parameters are simulation duration in second (s), packet size in Byte (B), packet generation rate in packets per second (pps) and SF assignment method.
+With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput in bits per second (bps) and total transmit energy consumption in Joule (J).
 Simulator also produces prediction accuracy and confusion matrix for machine learning schemes.
@@
 $L^{dB}_{PATH}$ is the propagation path loss between transmitter and receiver antennas in open space.
-In the simulator, it is assumed that total of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dB.
+In the simulator, it is assumed that sum of system gains $G^{dB}_{SYS}$ and system losses $L^{dB}_{SYS}$ is +7 dB.
 
-\par In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm for European ISM band regulation \cite{lorawan.regional.parameters}.
+\begin{table}
+\centering
+\caption{EU863-870 ISM band LoRa default channels \cite{lorawan.regional.parameters}.}
+\label{table:max_tx_power}
+\begin{tabular}{|c|c|c|}
+\hline
+\textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
+      868.1 &   125 &   14 \\ \hline
+      868.3 &   125 &   14 \\ \hline
+      868.5 &   125 &   14 \\ \hline
+\end{tabular}
+\end{table}
+
+\par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
+In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm.
 Different channel transmissions are independent from each other.
@@
 
-\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in \cite{SX1276}. In the simulator, 125 kHz bandwidth receive sensitivities are used.
+\begin{table}
+\centering
+\caption{Gateway sensitivity for different SFs \cite{SX1276}.}
+\label{table:gw_sf_sensitivity}
+\begin{tabular}{|c|c|c|c|c|c|c|c|}
+\hline
+\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
+\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
+\multirow{3}{*}{\textbf{BW (kHz)}}  & 125 & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
+                                    & 250 & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
+                                    & 500 & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
+\end{tabular}
+\end{table}
 
-\par Free space propagation loss is calculated as in \cite{TR136.942}. In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz.
+\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in Table \ref{table:gw_sf_sensitivity}.
+In the simulator, 125 kHz bandwidth receive sensitivities are used.
+
+\par Free space propagation loss is calculated as \cite{TR136.942}:
+
+\begin{equation} \label{eq:propagation_loss}
+\begin{split}
+P^{dB}_{PATH} = 40(1 - 4 \times 10^{-3} \times h){\log_{10} R|_{km}} \\
+- 18 {\log_{10} h|_{m}} + 21 {\log_{10} f|_{MHz}} + 80
+\end{split}
+\end{equation}
+
+\par Where, $h$ is the gateway altitude and $f$ is the frequency of the signal.
+In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz.
+With these assumptions, propagation loss calculation become \cite{7996384}:
+
+\begin{equation} \label{eq:propagation_loss_simplified}
+P^{dB}_{PATH} = 120.5 + 37.6 {\log_{10} R|_{km}}
+\end{equation}
 
@@
 
-\par It is assumed that there is no other technology interference in the network except LoRa interference.
+\par In this work, it is assumed that there is no other technology interference in the network except LoRa interference.
 To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other.
-In simulator, signal to interference plus noise ratio (SINR) threshold matrix from \cite{goursaud:hal-01231221} is used.
+In simulator, signal to interference plus noise ratio (SINR) threshold matrix $T_{i,j}$ from \cite{goursaud:hal-01231221} is used.
 
@@
 If there are more than one interfering signal, referenced signal must satisfy the margin for cumulative sum of all interfering signal received power for each SF \cite{7996384}.
+SINR threshold can be calculated as:
+
+\begin{equation} \label{eq:sinr_db}
+SINR_{i,j} = \dfrac{P_{rc,0}}{\sum_{l \in I_j} P_{rc,l}}
+\end{equation}
+
+\par Where $P_{rc,0}$ is received signal power of referenced signal and $P_{rc,l}$ is received signal power of interfering signal for SF = j.
+If a packet with SF = i satisfies the following condition for every SF = j, then packet is survived from all interferences.
+
+\begin{equation} \label{eq:sinr_t}
+SINR_{i,j}^{dB} > T_{i,j}
+\end{equation}
+
+\par To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power at receiver \cite{7996384}:
+
+\begin{equation} \label{eq:p_interference}
+P_{rc,y}^{interf} = \dfrac{P_{rc,y}(t_{interf})}{t_{x}}
+\end{equation}
+
+\par Where $t_{x}$ is transmission duration of referenced signal.
+$t_{interf}$ is overlapping duration between referenced signal and interfering signal.
 Transmission duration of a packet can be calculated by data rate $R_{b}$ and packet size $PS$.
 Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.
+Duration of a LoRa transmission can be calculated as:
+
+\begin{equation} \label{eq:transmission_duration}
+t_{x} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}  \ s
+\end{equation}
 
 \section{Simulation Results} \label{Simulation Results}
-\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds, packet generation rate = 0.01 pac/sec.
+\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds, packet generation rate = 0.01 pps.
 

### 2019-02-27 paper.tex
training details

@@
 Furthermore, a novel smart SF assignment strategy which utilizes Support Vector Machine (SVM) and Decision Tree Classifier (DTC) machine learning techniques for optimization of SF assignment is proposed.
-It is observed and presented that the proposed smart SF assignment techniques give promising simulation results.
+It is observed and presented that the proposed smart SF assignment techniques give promising simulation results in terms of packet delivery ratio (PDR).
 \end{abstract}
@@
 \section{Introduction}
-\par In the last few years, number of Internet of Things (IoT) applications increased exponentially \cite{7721743}.
+\IEEEPARstart{N}{umber} of Internet of Things (IoT) applications increased exponentially in last few years \cite{7721743}.
 Recent developments on LPWAN technologies has great impact on growth of number of IoT applications.
@@
 
-\par There are several emerging LPWAN technologies.
+There are several emerging LPWAN technologies.
 LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well-known LPWAN technologies \cite{7815384}.
@@
 
-\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth.
+LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth.
 This enables tradeoff between receive sensitivity and air time of transmission \cite{7803607}.
@@
 
-\par This paper is organized as follows: Section \ref{LoRa} and Section \ref{LoRaWAN} provides some background information about LoRa and LoRaWAN.
+This paper is organized as follows: Section \ref{LoRa} and Section \ref{LoRaWAN} provides some background information about LoRa and LoRaWAN.
 Section \ref{Related Works} summarize  the other related works.
@@
 \section{LoRa} \label{LoRa}
-\par LoRa is a proprietary physical layer radio/chipset technology that provides wireless link solution for low power wide area networks.
+LoRa is a proprietary physical layer radio/chipset technology that provides wireless link solution for low power wide area networks.
 It uses proprietary spread spectrum modulation technique that is the derivative of Chirp Spread Spectrum (CSS).
@@
 \subsection{Spreading Factor}
-\par The ratio between symbol and chirp rate is equal to $2$\textsuperscript{SF}.
+The ratio between symbol and chirp rate is equal to $2$\textsuperscript{SF}.
 SF can take values between 7 to 12.
@@
 
-\par When BW and CR are constant, as the SF increases, the data rate decreases.
+When BW and CR are constant, as the SF increases, the data rate decreases.
 Increasing the SF makes the signal more resilient to noise thus increases the transmission range.
@@
 \subsection{Spreading Factor Assignment Issue}
-\par Simultaneous different SF transmissions are orthogonal to each other up to some extent.
+Simultaneous different SF transmissions are orthogonal to each other up to some extent.
 Which means that, a LoRa gateway can simultaneously receive multiple transmissions with different SFs.
@@
 
-\par In a LoRaWAN network, initially, a node is not aware of how far away it is from a gateway.
+In a LoRaWAN network, initially, a node is not aware of how far away it is from a gateway.
 However, a node can guess the distance from a gateway by observing received signal power of a downlink transmission.
@@
 
-\par In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single gateway is illustrated.
+In Figure \ref{fig:collision}, a LoRaWAN network deployed with a single gateway is illustrated.
 Different color rings represent achievable range of different SFs from the gateway and different color circles represent selected SF of the nodes.
@@
 \section{LoRaWAN} \label{LoRaWAN}
-\par LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
+LoRa has an open standard medium access control (MAC) layer protocol called LoRaWAN which is designed for large scale LoRa networks considering well known LPWAN challenges and their best practice solutions.
 LoRaWAN is developed and maintained by LoRa Alliance.
@@
 \subsection{End node}
-\par LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways.
+LoRaWAN end node (EN) is a low power embedded device that only communicates to gateways.
 LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C.
@@
 \subsection{Gateway}
-\par LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes.
+LoRaWAN gateway (GW) is a device that receive/transmit packets coming from/to end nodes.
 A typical gateway can receive from multiple channels at the same time.
@@
 \subsection{Network server}
-\par LoRaWAN network server (NS) is a server that provides MAC layer processing.
+LoRaWAN network server (NS) is a server that provides MAC layer processing.
 Network server routes messages from application to end nodes and vice versa.
@@
 \section{Related Works} \label{Related Works}
-\par The literature related to the work presented in this paper has started to grow recently.
+The literature related to the work presented in this paper has started to grow recently.
 LPWAN technologies and especially LoRa attracted researchers’ attentions lately.
@@
 
-\par In \cite{7996384}, the authors evaluated the performance of LoRaWAN networks in a smart city scenario.
+In \cite{7996384}, the authors evaluated the performance of LoRaWAN networks in a smart city scenario.
 The authors proposed a link measurement and a link performance model for LoRa.
@@
 
-\par In \cite{8090518}, another LoRaWAN ns-3 simulator is presented.
+In \cite{8090518}, another LoRaWAN ns-3 simulator is presented.
 Authors introduced an error model for determining range as well as interference between multiple simultaneous LoRa transmissions.
@@
 
-\par In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
+In \cite{8267219}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
 The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal.
@@
 
-\par In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs.
+In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs.
 They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the gateway for modeling interference in multiple gateways scenarios.
@@
 
-\par In \cite{8115779}, the authors introduced two new algorithms for SF assignment in LoRaWANs.
+In \cite{8115779}, the authors introduced two new algorithms for SF assignment in LoRaWANs.
 First algorithm assigns SFs based on the total number of connected devices, second algorithm assigns SFs by balancing air time of the packets in each SF.
@@
 \section{Proposed Technique} \label{Proposed Technique}
-\par The collision issue illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even though they are able to communicate with lower SFs.
+The collision problem illustrated in Figure \ref{fig:collision} is solved by forcing some of the close nodes to select higher SFs even though they are able to communicate with lower SFs.
 This has potential to prevent collisions due to the orthogonality of the SFs as shown in Figure \ref{fig:collision_solution_single_gw}.
@@
 
-\par It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules.
+It is difficult to propose a single SF assignment rule for every possible LoRaWAN topology since every network is different and optimizing their nodes' SFs requires different rules.
 For this reason, machine learning based SF assignment approach is proposed to decrease the collisions for the same SF transmissions.
@@
 However, NS cannot keep track of transmissions with lower receive power than sensitivity of the gateway.
-Using this information, NS can train a classifier to predict future transmission result for a specific node and a specific SF.
+Using those obtained information, NS can train a classifier to predict future transmission result for a specific node and a specific SF.
 Using this prediction model NS can assign SFs to nodes considering the collision probability.
 
-\par In this work, decision tree classifier (DTC) and support vector machine (SVM) \cite{Alpaydin} schemes are employed to predict the transmission results.
+In this work, decision tree classifier (DTC) and support vector machine (SVM) \cite{Alpaydin} schemes are employed to predict the transmission results.
 Class weights are balanced according to sample distributions for both methods.
@@
 
-\par It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator.
-Data set size is directly proportional to simulation duration.
-Thus, increasing the simulation duration, improves the prediction accuracy.
+It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator.
+Dataset size is directly proportional to simulation duration.
+Thus, increasing the simulation duration, improves the prediction accuracy up to some extent.
 In real world deployments, NS can keep track of transmissions and it can create a classifier daily basis, then gateways can request from nodes to use suggested SFs.
 
-\par To come up with a solution for SF assignment, DTC and SVM prediction schemes are integrated to the simulator.
-A classifier is trained from transmission logs and this classifier is used for selecting optimum SF for the nodes.
-Simulator first runs a simulation by assigning random SFs to the nodes.
-Simulation logs are combined into four data columns which are: X and Y coordinates of the transmission source, SF of the transmission and the result of the transmission.
-Possible transmission results are successful, interfered or under sensitivity.
-These data are fed to Python scikit-learn DTC or SVM classifiers.
-After the classifier model is generated, a second simulation is run with classifier.
-Simulator selects optimum SF considering the prediction of transmission.
-If a transmission of a node with the lowest possible SF is predicted as interfered by the classifier, then simulator assigns higher SF and evaluates the new prediction.
-If no SF is predicted as successful, then simulator again assigns the lowest possible SF.
+There are three features in the dataset generated by the simulation tool.
+Dataset features are: X and Y coordinates of the transmission source, SF of the transmission.
+X and Y coordinate feature values are continuous numbers, SF feature values are integer numbers between 7 to 12.
+Dataset class label is the result of the transmission.
+Dataset class label values are successful, interfered and under sensitivity.
+DTC and SVM prediction schemes are integrated into the simulation tool to study smart SF assignment schemes.
+A classifier is trained from generated dataset and this classifier is used for selecting optimum SF for the nodes in the network.
+
+The tool first runs a simulation with random SF scheme.
+After random SF scheme simulation completed, transmission logs are combined into three feature columns and one class label column to create the dataset.
+Then the dataset is fed to Python scikit-learn DTC or SVM classifiers for training phase.
+After the classifier model is built, second simulation is run with the trained classifier.
+The tool selects optimum SF for transmissions considering prediction of the transmissions.
+For every transmission, the classifier calculates the transmission result for the lowest possible SF.
+If the transmission result is classified as interfered, then the tool increases the SF and predicts a new transmission result.
+If the new transmission result is classified as successful, then simulator continue to execute with selected SF.
+If no SF transmission result is predicted as successful, then the tool selects the lowest possible SF and continue to execute.
 
@@
 \section{Simulation Environment} \label{Simulation Environment}
-\par A discrete event simulator is implemented in Python to study the effects of different SF strategies in LoRaWANs.
-LoRaWAN SF simulation tool source code is available at \cite{simlorasf}.
+A discrete event simulator is developed in Python to study the effects of different SF strategies in LoRaWANs.
+LoRaWAN SF simulation tool source code is available at GitHub \cite{simlorasf}.
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
@@
 With these inputs, the simulator produces total number of generated packets, number of successfully received packets, number of interfered packets, number of under sensitivity packets, network packet delivery ratio percentage (PDR), network throughput in bits per second (bps) and total transmit energy consumption in Joule (J).
-Simulator also produces prediction accuracy and confusion matrix for machine learning schemes.
+Simulator also produces prediction accuracy percentage and confusion matrix for machine learning schemes.
 
-\par The simulator only covers the LoRaWAN Class A devices.
+The simulator only covers the LoRaWAN Class A devices.
 Transmissions are always initiated by end nodes in pure ALOHA manner.
@@
 \subsection{Link Model Employed}
-\par Link quality of a wireless system can be expressed by the metric of link budget.
+Link quality of a wireless system can be expressed by the metric of link budget.
 Link budget is a measure of all gains and losses from transmitter device to receiver device.
@@
 
-\par Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver.
+Where, $P^{dBm}_{RX}$ is the expected receive power at the receiver.
 $P^{dBm}_{TX}$ is the transmit power of the transmitter.
@@
 
-\par Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
+Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
 In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm.
@@
 
-\par Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in Table \ref{table:gw_sf_sensitivity}.
-In the simulator, 125 kHz bandwidth receive sensitivities are used.
+Receive sensitivity of a LoRa gateway for different SFs and BWs in dBm unit can be found in Table \ref{table:gw_sf_sensitivity}. In the simulator, 125 kHz bandwidth receive sensitivities are used.
 
-\par Free space propagation loss is calculated as \cite{TR136.942}:
+Free space propagation loss is calculated as \cite{TR136.942}:
 
@@
 
-\par Where, $h$ is the gateway altitude and $f$ is the frequency of the signal.
+Where, $h$ is the gateway altitude and $f$ is the frequency of the signal.
 In this work, it is assumed that $h$ = 15 m and $f$ = 868 MHz.
@@
 
-\par If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission.
+If the received signal power is higher than the gateway sensitivity, then signal can be decoded by the receiver successfully when there is no interfering transmission.
 
 \subsection{Interference Model Employed}
-\par In the simulator, interference model described in \cite{7996384} is adopted.
-They use SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
+In the simulator, interference model described in \cite{7996384} is adopted. They use SINR threshold matrix for modeling LoRa interference between simultaneous but different SF LoRa transmissions.
 
-\par In this work, it is assumed that there is no other technology interference in the network except LoRa interference.
+In this work, it is assumed that there is no other technology interference in the network except LoRa interference.
 To exploit imperfect orthogonality of different SF transmission, simulator should calculate the effect of different SF transmissions to each other.
@@
 
-\par To decide if a referenced signal is interfered at receiver by an interfering signal, SINR threshold matrix is used.
+To decide if a referenced signal is interfered at receiver by an interfering signal, SINR threshold matrix is used.
 $T_{i,j}$ is SINR margin in dB unit between referenced signal with SF = i and interfering signal with SF = j to correctly decode the referenced signal.
@@
 
-\par Where $P_{rc,0}$ is received signal power of referenced signal and $P_{rc,l}$ is received signal power of interfering signal for SF = j.
+Where $P_{rc,0}$ is received signal power of referenced signal and $P_{rc,l}$ is received signal power of interfering signal for SF = j.
 If a packet with SF = i satisfies the following condition for every SF = j, then packet is survived from all interferences.
@@
 
-\par To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power at receiver \cite{7996384}:
+To calculate interfering power at receiver, we should consider the case where two transmissions are not perfectly overlapping. To equalize the interfering power at receiver \cite{7996384}:
 
@@
 
-\par Where $t_{x}$ is transmission duration of referenced signal.
+Where $t_{x}$ is transmission duration of referenced signal.
 $t_{interf}$ is overlapping duration between referenced signal and interfering signal.
@@
 \section{Simulation Results} \label{Simulation Results}
-\par For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds, packet generation rate = 0.01 pps.
+For simulation results in this paper, global simulation parameters are set as follows: packet size = 60 Bytes, simulation duration = 3600 seconds and packet generation rate = 0.01 pps.
+First, single gateway and multiple gateway LoRaWAN network simulation results are presented to show correctness of the simulator, then, smart SF scheme simulation results are presented.
 
-\par In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown.
+\subsection{Single Gateway}
+In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown.
 Randomly generated network topology radius is set to 3000 meters and number of gateways is set to 1.
@@
 
-\par In Figure \ref{fig:gw_pdr}, PDR plot for various number of gateways is shown.
+In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown.
+Number of gateway is set to 1 and lowest SF assignment scheme is used.
+Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{{r_pdr_g1_p0.01_s3600}.png}
+\caption{PDR for various network radii. (GW = 1, SF = SF{\_}Lowest)}
+\label{fig:r_pdr}
+\end{figure}
+
+\subsection{Multiple Gateway}
+In Figure \ref{fig:gw_pdr}, PDR plot for various number of gateways is shown.
 Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used.
@@
 
-\par In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown.
-Number of gateway is set to 1 and lowest SF assignment scheme is used.
-Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]{{r_pdr_g1_p0.01_s3600}.png}
-\caption{PDR for various network radii. (GW = 1, SF = SF{\_}Lowest)}
-\label{fig:r_pdr}
-\end{figure}
-
-\par PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}.
+\subsection{Smart Spreading Factor Schemes}
+PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}.
 Randomly generated network radius is set to 5000 meters and number of gateways is set to 3.
@@
 
-\par In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes.
+In Table \ref{table:prediction_accuracy}, prediction accuracy of smart SVM and smart DTC schemes are presented for various network radii and number of nodes.
 Prediction accuracy is not directly proportional to network PDR.
@@
 Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.
-LoRaWAN simulation usually produces imbalanced dataset, DTC gives better accuracy for interfered transmission labels.
+LoRaWAN simulation usually produces imbalanced dataset.
+Under sensitivity transmission labeled data points are less than successful transmission labeled data points.
+Interfered transmission labeled data points are even less then Under sensitivity transmission labeled data points.
+DTC gives better accuracy for interfered transmission label.
+Correct classification of interfered transmissions yields better network PDR results for smart SF scheme.
 
@@
 
-\par For space constraints, we omit simulation results like network throughput, total transmit energy consumption and confusion matrices.
+For space constraints, we omit simulation results like network throughput, total transmit energy consumption and confusion matrices.
 We invite readers to experiment the simulation tool with the parameters they like \cite{simlorasf}.
@@
 \section{Conclusion} \label{Conclusion}
-\par In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed.
+In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed.
 We present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
@@
 
-\par As for future work, transmit power optimization can be included to the proposed smart schemes.
+As for future work, transmit power optimization can be included to the proposed smart schemes.
 In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to gateways can decrease transmit power to save energy.
@@
 \section*{Acknowledgment}
-\par This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1.
+This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1.
 

### 2019-02-27 paper.tex
prediction results explanation

@@
 It is possible to generate mass amount of LoRaWAN transmission logs for different topologies using our simulator.
-Dataset size is directly proportional to simulation duration.
+Training dataset size is directly proportional to simulation duration.
 Thus, increasing the simulation duration, improves the prediction accuracy up to some extent.
@@
 There are three features in the dataset generated by the simulation tool.
-Dataset features are: X and Y coordinates of the transmission source, SF of the transmission.
-X and Y coordinate feature values are continuous numbers, SF feature values are integer numbers between 7 to 12.
-Dataset class label is the result of the transmission.
-Dataset class label values are successful, interfered and under sensitivity.
+Features of the dataset are: X coordinate of the transmission source, Y coordinate of the transmission source and SF of the transmission.
+X and Y coordinate feature values are continuous numbers.
+SF feature values are integer numbers between 7 to 12.
+Class label of the dataset is the result of the transmission.
+Class label values are successful, interfered and under sensitivity.
 DTC and SVM prediction schemes are integrated into the simulation tool to study smart SF assignment schemes.
@@
 The tool first runs a simulation with random SF scheme.
-After random SF scheme simulation completed, transmission logs are combined into three feature columns and one class label column to create the dataset.
+After random SF scheme simulation completed, transmission logs are combined into three feature columns and one class label column to create the training dataset.
 Then the dataset is fed to Python scikit-learn DTC or SVM classifiers for training phase.
@@
 The tool selects optimum SF for transmissions considering prediction of the transmissions.
-For every transmission, the classifier calculates the transmission result for the lowest possible SF.
-If the transmission result is classified as interfered, then the tool increases the SF and predicts a new transmission result.
+For every transmission, the classifier predicts the transmission result for the lowest possible SF.
+If the transmission result is predicted as interfered, then the tool increases the SF and predicts a new transmission result.
 If the new transmission result is classified as successful, then simulator continue to execute with selected SF.
@@
 \subsection{Single Gateway}
-In Figure \ref{fig:sf_pdr}, PDR plot of various SF assignment is shown.
+In Figure \ref{fig:sf_pdr}, PDR plots of various SF assignment schemes are shown.
 Randomly generated network topology radius is set to 3000 meters and number of gateways is set to 1.
 Increasing SFs increases air time.
-It increases the number of collisions thus decreases the PDR.
+This increases the number of collisions thus decreases the PDR.
 High SF schemes gives poor PDR results as the number of nodes increases.
@@
 
-In Figure \ref{fig:r_pdr}, PDR plot of various network radius is shown.
+In Figure \ref{fig:r_pdr}, PDR plots of various network radii are shown.
 Number of gateway is set to 1 and lowest SF assignment scheme is used.
@@
 \subsection{Multiple Gateway}
-In Figure \ref{fig:gw_pdr}, PDR plot for various number of gateways is shown.
+In Figure \ref{fig:gw_pdr}, PDR plots for various number of gateways are shown.
 Randomly generated topology radius is set to 3000 meters and the lowest SF assignment scheme is used.
@@
 
-\subsection{Smart Spreading Factor Schemes}
-PDR plot of the lowest SF scheme and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}.
+\subsection{Smart SF Schemes}
+PDR plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}.
 Randomly generated network radius is set to 5000 meters and number of gateways is set to 3.
@@
 Prediction accuracy is not directly proportional to network PDR.
-Correct prediction of an interfered transmission may not increase the PDR but increases the accuracy.
-Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is high.
-LoRaWAN simulation usually produces imbalanced dataset.
-Under sensitivity transmission labeled data points are less than successful transmission labeled data points.
-Interfered transmission labeled data points are even less then Under sensitivity transmission labeled data points.
-DTC gives better accuracy for interfered transmission label.
-Correct classification of interfered transmissions yields better network PDR results for smart SF scheme.
+Correct prediction of an interfered transmission may not increase the PDR but increases the prediction accuracy.
+Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is higher.
+LoRaWAN simulation usually produces imbalanced training dataset.
+Number of under sensitivity transmission labeled data points are less than number of successful transmission labeled data points.
+Number of interfered transmission labeled data points are even less then number of under sensitivity transmission labeled data points.
+In our case, DTC gives better prediction accuracy for interfered transmission labeled data points.
+Correct classification of interfered transmissions yields better network PDR results, thus smart DTC scheme gives better network PDR results.
 
@@
 In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed.
-We present a discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
+We present a open source discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
 Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme.

### 2019-02-27 paper.tex
simlorafs doi reference

@@
 A discrete event simulator is developed in Python to study the effects of different SF strategies in LoRaWANs.
-LoRaWAN SF simulation tool source code is available at GitHub \cite{simlorasf}.
+LoRaWAN SF simulation tool source code is available at GitHub \cite{tugrul_yatagan_2019_2579366}.
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
@@
 For space constraints, we omit simulation results like network throughput, total transmit energy consumption and confusion matrices.
-We invite readers to experiment the simulation tool with the parameters they like \cite{simlorasf}.
+We invite readers to experiment the simulation tool with the parameters they like \cite{tugrul_yatagan_2019_2579366}.
 

### 2019-02-28 paper.tex
prediction paraphrase

@@
 Different color rings represent achievable range of different SFs from the gateway and different color circles represent selected SF of the nodes.
-The end devices close to the gateway will fall into lowest SF (SF7) area section.
+The end devices close to the gateway will fall into the lowest SF (SF7) area section.
 The end devices close to the gateway will probably select the lowest SF most of the time.
@@
 A typical gateway can receive from multiple channels at the same time.
-gateways are usually connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.
+Gateways are usually connected to power grid, so power consumption of a gateway is insignificant in most of the deployments.
 
@@
 In Figure \ref{fig:r_pdr}, PDR plots of various network radii are shown.
-Number of gateway is set to 1 and lowest SF assignment scheme is used.
+Number of gateways is set to 1 and lowest SF assignment scheme is used.
 Increasing network radius, increases the number of under sensitivity transmissions thus decreases the PDR of network.
@@
 Correct prediction of an interfered transmission may not increase the PDR but increases the prediction accuracy.
-Smart DTC gives better PDR results than smart SVM, even overall SVM prediction accuracy is higher.
-LoRaWAN simulation usually produces imbalanced training dataset.
+Smart DTC gives better network PDR results than smart SVM, even overall smart SVM prediction accuracy is higher.
+The distribution of different labeled data points in the dataset is strongly related to simulation parameters.
+If simulation is run with small topology radius and high number of nodes, then number of interfered transmission labeled data points increases.
+On the other hand, using a large topology radius in the simulation causes the number of under sensitivity transmission labeled data points to increase.
+We choose moderately dense network parameters to bring it closer to real world deployments.
+With the simulation parameters utilized in this section, simulation usually produces imbalanced dataset.
 Number of under sensitivity transmission labeled data points are less than number of successful transmission labeled data points.
-Number of interfered transmission labeled data points are even less then number of under sensitivity transmission labeled data points.
-In our case, DTC gives better prediction accuracy for interfered transmission labeled data points.
+Besides, number of interfered transmission labeled data points are even less then number of under sensitivity transmission labeled data points.
+In this case, smart DTC predicts interfered transmission labeled data points more accurately.
 Correct classification of interfered transmissions yields better network PDR results, thus smart DTC scheme gives better network PDR results.
@@
 In this paper, after a brief introduction about LPWAN technologies, LoRa modulation basics and spreading factor assignment issue is discussed.
-We present a open source discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
+We present an open source discrete event simulator which is developed from scratch to study network performance of LoRaWAN and evaluate different SF assignment schemes.
 Moreover, we show how same SF collisions can be avoided, hence, we propose a machine learning based solutions called smart DTC scheme and smart SVM scheme.

### 2019-04-28 paper.tex
decrease figure sizes
Add Smetech ADR as realted work

@@
 Support Vector Machine (SVM) and Decision Tree Classifier (DTC) techniques are employed and the introduced schemes are called as smart SF schemes.
-The performance of the smart schemes is compared with the performance of the lowest SF assignment scheme.
+The performance of the smart schemes are compared with the performance of the lowest SF assignment scheme.
 It is shown that the proposed smart SF assignment schemes give promising results.
@@
 \centering
-\includegraphics[width=\linewidth]{collision}
+\includegraphics[width=0.85\linewidth]{collision.png}
 \caption{Collision between nodes close to the gateway.}
@@
 
-In \cite{8430542}, the authors investigated the impact of the interference caused by simultaneous LoRa transmissions with the same and different SFs.
-They derived aggregated co-SF and inter-SF interference power SIR distributions to capture the coverage distance from the gateway for modeling interference in multiple gateways scenarios.
+\par In \cite{8267219} and \cite{8430542}, the authors studied the effects of imperfect orthogonality between different LoRa SF transmissions.
+The authors state that a LoRa transmission can be interfered by a different SF transmission when power of the interfering signal is significantly greater than the reference signal.
 Their results show that transmission among different SFs can cause a significant impact in high-density LoRaWAN networks.
 
-In \cite{8115779}, the authors introduced two new algorithms for SF assignment in LoRaWANs.
-First algorithm assigns SFs based on the total number of connected devices, second algorithm assigns SFs by balancing air time of the packets in each SF.
-Based on the simulation results, it is shown that the proposed algorithms give promising results.
+The adaptive data rate (ADR) algorithm recommended by Semtech Corporation utilizes signal to noise ratio (SNR) of the last 20 transmissions to minimize the transmit power and SF while ensuring successful transmissions \cite{lorawan_adr}.
+The recommended algorithm uses a predefined SF and corresponding required SNR table.
+The algorithm considers maximum SNR of the last 20 transmissions.
+The algorithm increases transmit power or decrease data rate in case of low SNR and the opposite in case of high SNR.
 
@@
 \centering
-\includegraphics[width=\linewidth]{collision_solution_single_gw}
+\includegraphics[width=0.85\linewidth]{collision_solution_single_gw.png}
 \caption{Collision avoidance by using higher SF for nodes close to the gateway.}
@@
 \centering
-\includegraphics[width=\linewidth]{collision_solution_multi_gw}
+\includegraphics[width=\linewidth]{collision_solution_multi_gw.png}
 \caption{Collision avoidance for intersecting gateways.}
@@
 
-As for future work, transmit power optimization can be included to the proposed smart schemes.
+As for future work, transmit power optimization can be included to the proposed smart scheme.
 In this paper, it is assumed that nodes always use maximum transmit power for uplink transmission, however nodes close to gateways can decrease transmit power to save energy.
@@
 Reinforcement learning could be a good candidate.
-Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme to the improve prediction performance.
+Also, other transmission parameters such as node id and transmission time can be included to the proposed scheme in order to improve prediction performance.
 

### 2019-05-01 paper.tex
prediction energy figure is added

@@
 \centering
-\includegraphics[width=0.85\linewidth]{collision.png}
+\includegraphics[width=0.83\linewidth]{collision.png}
 \caption{Collision between nodes close to the gateway.}
@@
 The recommended algorithm uses a predefined SF and corresponding required SNR table.
-The algorithm considers maximum SNR of the last 20 transmissions.
 The algorithm increases transmit power or decrease data rate in case of low SNR and the opposite in case of high SNR.
@@
 \centering
-\includegraphics[width=0.85\linewidth]{collision_solution_single_gw.png}
+\includegraphics[width=0.83\linewidth]{collision_solution_single_gw.png}
 \caption{Collision avoidance by using higher SF for nodes close to the gateway.}
@@
 
-\begin{table}
-\centering
-\caption{EU863-870 ISM band LoRa default channels \cite{lorawan.regional.parameters}.}
-\label{table:max_tx_power}
-\begin{tabular}{|c|c|c|}
-\hline
-\textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
-      868.1 &   125 &   14 \\ \hline
-      868.3 &   125 &   14 \\ \hline
-      868.5 &   125 &   14 \\ \hline
-\end{tabular}
-\end{table}
-
-Maximum transmit power for European ISM band LoRa default channels can be found in Table \ref{table:max_tx_power}.
-In the simulator, it is assumed that nodes always select maximum allowed transmit power, which is 14 dBm.
+In the simulator, it is assumed that nodes always select maximum allowed transmit power for European ISM band, which is 14 dBm \cite{lorawan.regional.parameters}.
 Different channel transmissions are independent from each other.
@@
 Data rate of a LoRa transmission is already expressed in Equation \ref{eq:bit_rate_sf}.
-Duration of a LoRa transmission can be calculated as:
-
-\begin{equation} \label{eq:transmission_duration}
-t_{x} = \dfrac{8 \times PS|_{B}}{R_{b}|_{bps}}  \ s
-\end{equation}
 
@@
 \subsection{Smart SF Schemes}
-PDR plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr}.
+PDR and transmit energy consumption plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr} and \ref{fig:prediction_energy}.
 Randomly generated network radius is set to 5000 meters and number of gateways is set to 3.
@@
 In Table \ref{table:prediction_pdr}, PDR values for various network radii are presented.
-Both smart SVM and smart DTC schemes give better PDR than lowest SF schemes when number of nodes increases.
+Random SF scheme consumes far more energy than others, therefore it is not feasible for real world deployments.
+As number of nodes increases, both smart SVM and smart DTC schemes consume slightly more energy than lowest SF scheme, however they give better PDR results in return.
 Increasing number of nodes, increases number of interferences.
 Smart schemes improve network performance when LoRa interference is high.
-Moreover, smart schemes give better results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway.
+Moreover, smart schemes give better PDR results when nodes are deployed closer to the gateway, since nodes have margin to increase their SFs when they are deployed closer to the gateway.
 If a node is far away from gateway, then smart schemes cannot increase the SF to avoid interference since the assigned SF is already high.
@@
 \label{fig:prediction_pdr}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{{prediction_energy_r5000_g3_p0.01_s3600}.png}
+\caption{TX energy for lowest and smart SF schemes. (r = 5000 m, GW = 3)}
+\label{fig:prediction_energy}
 \end{figure}

### 2019-05-01 paper.tex
ack grammer

@@
 \subsection{Smart SF Schemes}
-PDR and transmit energy consumption plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr} and \ref{fig:prediction_energy}.
+PDR and transmit energy consumption plots of the lowest SF, random SF and the smart prediction schemes are shown in Figure \ref{fig:prediction_pdr} and Figure \ref{fig:prediction_energy}.
 Randomly generated network radius is set to 5000 meters and number of gateways is set to 3.
@@
 \section*{Acknowledgment}
-This work is supported by Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1.
+This work is supported by the Turkish Ministry of Development and Istanbul Technical University researcher support program under the Grant No. ITU-AYP-2017-1.
 

### 2019-05-07 paper.tex
update simlorasf doi

@@
 A discrete event simulator is developed in Python to study the effects of different SF strategies in LoRaWANs.
-LoRaWAN SF simulation tool source code is available at GitHub \cite{tugrul_yatagan_2019_2579366}.
+LoRaWAN SF simulation tool source code is available at GitHub \cite{tugrul_yatagan_2019_2671495}.
 Simulation tool supports custom LoRaWAN topologies as well as randomly generated LoRaWAN topologies.
@@
 For space constraints, we omit simulation results like network throughput, total transmit energy consumption and confusion matrices.
-We invite readers to experiment the simulation tool with the parameters they like \cite{tugrul_yatagan_2019_2579366}.
+We invite readers to experiment the simulation tool with the parameters they like \cite{tugrul_yatagan_2019_2671495}.
 
